### 2020-05-06 LaTex/paper.tex
current version

\documentclass[BCOR=1mm, DIV=calc,10pt,
twoside=true,
twocolumn,
headings=normal]{scrartcl}
%\KOMAoptions{DIV=calc}% recalculate the page layout with a calculated DIV value

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

\usepackage{amsmath}
\usepackage{bm}
\usepackage{url}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}

\usepackage{authblk}

\newcommand{\fig}{Fig. }
\newcommand{\eqn}{Eqn. }
\newcommand{\tab}{Tab. }
\newcommand{\wrt}{w.r.t. }
\newcommand{\etal}{ {\em et al. }}

\begin{document}

%\title{Prediction and Evaluation of individual Probability Density Functions with Machine Learning}
\title{Demand Forecasting of individual Probability Density Functions with Machine Learning}

\author[1]{Felix Wick \thanks{felix.wick@blueyonder.com}}
\author[2]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
\author[1]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
\author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
\affil[1]{\small Blue Yonder GmbH (Karlsruhe, Germany)}
\affil[2]{\small IUBH Internationale Hochschule (Erfurt, Germany)}

\date{}

\maketitle

\begin{abstract}
... abstract ...
\end{abstract}

{Keywords: \textbf{machine learning, demand forecasting}}

\section{Introduction}

Demand forecasting is one of the main challenges for retailers and at the core of business
operations. Due to its stochastic nature, demand is difficult to forecast as it depends on
many influencing factors and the realized demand can be interpreted as a random variable
that is described by an appropriate probability density function (PDF). In order to make
operational decisions, an optimal point estimator has to be defined that  can be used to
derive ordering decisions used in the replenishment process of the retailer. Demand
estimation is further complicated by the fact that retailers typically only observed
realized sales and not the actual demand, in case the demand exceeds the current stock
level, the data become censored. This decision process is complicated by a range of
factors: Even in the case of accurate  demand forecasts, the decision maker has to balance
conflicting metrics to reach an optimal decision: Ordering to few items may result in
stockout situations resulting in unrealized demand and unsatisfied customers. Ordering too
many items results in excess inventory which increases transport and storage costs and, in
the case of perishable goods, excessive waste as spoilt items need to be disposed of at
additional cost in addition to  the environmental impact. This situation is particularly
noticeable in the so called `'ultra-fresh'' category which includes items such as bakery
products, ready-meals, fresh diary products or certain meat products such as ground meat.
These items typically have a shelf-life from less than a business day to a few business
days at most with a continuous spectrum  depending on the exact item. In many situations,
additional constraints have to be considered to reach an optimal ordering decision:
Delivery cycles of items may vary depending on the type of item and the wholesaler or
manufacturer from which  they are procured. Retailers also operate at a given service
level to guarantee that a certain level of demand can be fulfilled. The exact service
level typically depends on the overall business policy of the retailer and may also depend
on individual products, ranging from ``never-out-of-stock'' items to a service level
exceeding e.g. 90\%.

The availability of Big Data allows capturing, storing and processing a vast amount of
data associated with demand such as historic sales records,  information about promotional
events or advertisements, pricing information, local weather at retail locations, seasonal
information  as well as a wide range of further variables. Modern machine learning
algorithms can then be used to predict the per-item demand distribution, corrected for
censored data from which an optimal point estimator can be derived to be used in the
subsequent ordering decision. It is important to note that demand as a random variable is
not identically and independently distributed (i.i.d.): While the probability distribution
describing the demand can be attributed to a given family or parametrization, the exact
parameters vary: Seasonal effects, finite life cycles of products and the introduction of
new products influence the demand distribution, as well as the local weather at the retail
location or the retail location itself in terms of size, assortment range, customer
diversity and other factors. The retailers themselves also actively influence demand by
using advertisements to highlight products, offering rebates or discounts for specific
products as well as pursuing an active pricing strategy. This means that while we can
generally assume that demand follows a specific type of probabilty distribution, its
parameters are unique to the  instance for which an estimate is required. For example, the
probability distribution governing the demand of a particular item is specific to the
item, date and retail location for which the forecast is made and depends on a wide range
of further influencing factors.

The remainder of the paper is organized as follows:
We first review the relevant literature and existing work in sec. \ref{sec:LitRev}.
We then describe our method to predict individual negative binomial PDFs by means of a
parametric approach including two distinct machine learning models for mean of variance in
sec. \ref{sec:pdfEstimation}.
After that we describe methods for the qualitative and quantitative evaluation of PDF
predictions in sec. \ref{sec:pdfEvaluation}.
Finally, we present a demand forecasting example to show an application of our methods in
sec. \ref{sec:example}.

\section{Literature Review}
\label{sec:LitRev}

Inventory management offers a rich theory and the extensive body of research can be
broadly grouped into the following two categories where the inventory control problem is
either based on some knowledge of the underlying demand  distribution or an integrated
approach that seeks to map directly from the available data (historic sales records and 
further variables) to the ordering decision. The latter approach is taken e.g. in
\cite{beutel2012safety,ban2019big,bertsimas2020predictive} and aims to avoid estimating
the underlying probability distribution. Although this approach seems preferable since it
avoids determining the full demand distribution and results directly in the desired
operational decision (order quantity), it faces several drawbacks. First of all, the full
probability distribution for the demand of an specific item at a given sales location and
business day includes all available information including the uncertainty of the modelled
demand. This can be used to simulate the performance on a per-item level and e.g. optimize
the impact on business strategy decisions on conflicting metrics such as stock out- and
waste-rate. Additionally, having the full demand probability distribution available allows
for an accute assessment of the forecast quality at all quantiles of the distribution,
including the often extensive tails of the distribution, as well as the analysis of
long-term  effect as the demand predictions including all future planned advertisements or
price-changes can be included into long-term forecasts. Furthermore, deriving the ordering
decision  directly couples the demand process with the complex delivery cycles and
constraints, keeping the steps separate allows greater operational flexibility and reduces
the impact of changing manufacturers or wholesalers, as well as  allowing a quick response
to changes in the delivery schedule wihtout having to re-compute the implicit demand
underlying the ordering decision. Finally, long-term demand forecasts may be shared with
other business units or external vendors  and wholesalers to ease their planning for the
production and supply-chain processes upstream of the retailer.

In contrast, more traditional inventory control systems rely on the knowledge of the
demand distribution in one form or another.  see e.g. \cite{silver1998} for an overview.
In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored
at regular intervals and orders are dispatched once the inventory level reaches a minimal
value $s$. In case of linear holding and shortage costs, such policies are optimal
\cite{Scarf1959}, although perishable goods pose more challenges, see e.g.
\cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be
included in these kind of inventory control systems \cite{minner2010periodic}. Perishable
goods are well described by the `'newsvendor-problem'' \cite{Edgeworth} where in the
simplest case all stock perishes at the end of the selling period (e.g. a business day). 
For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming
linear underage and overage costs $b,h >0$, the optimal quantile
$q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated
exactly. The main objective in any of these approaches is to determine the underlying
demand distribution. The simplest approach is to just use the observed sales events and
forecast these as a time-series  (see e.g. \cite{alwan2016}) or via sample average
approximation (SAA). (see e.g. \cite{shapiro2014} for an overview over SAA). However,
these approaches do not make use of any data apart from the sales record themselves,
although we know that many variables such as price, advertisements, etc. are highly
correlated with demand. Additionally, it is critical that we indeed reconstruct a demand
distribution, hence a simple point-estimator as provided by the most common statistical
techniques and machine-learning approaches will not suffice. Additionally, demand is not
identically and independently distributed (i.i.d.) but depends not only on external
factors such as season, weather, product life-cycle, but is also actively changed by the
retailer by setting a specific price, offering rebates or running advertisements.
Additionally, the demand implicitly depend on the location of the retail outlet as well as
the specifics of that location such as product  assortment influencing the choice of
possible replacement articles and many more. These complications are the main reason we
cannot treat the replenishment process as $n$ independent newsvendor-type problems.
Instead, we need to determine the full demand distribution from data, conditional on the
relevant variables such as date, location and item, taking all auxiliary data such as
article charactersitics, pricing, advertisements, retail location details, etc. into
account. This can be done in several ways: First, we can use a neural network
\cite{Feindt2006190} to learn the distribution from data and return a full distribution
per item, store and day from which the relevant quantile can be estimated. Similarly,
using quantile regression \cite{koenker2001} this approach can be implemented in different
frameworks. Alternatively, one can assume a given demand model and fit the model
parameters instead of reconstructing the complete distribution. This approach is
computationally favourable, as fewer parameters need to be estimated compared to the case
of the full distribution. Empirically, one can determine the best fitting distribution
from data \cite{adan1995}. However, given the stochastic nature of the demand, such an
empirically determined distribution is not expected to be stable and prone to sudden
changes. Instead, the choice of the demand distribution should be motivated by theoretic
considerations. The discrete demand is typically modelled as a negative binomial
distribution (NBD), also known as Gamma-Poisson distribution
\cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This
distribution arises if the Poisson parameter $\mu$ is a random variable itself that
follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$ and
is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$.

Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be
determined for each item at each sales location and ordering time, depending on all
auxiliary data describing article details, retail location and influencing factors such as
pricing and advertisement information.

\subsection*{Summary of Contributions}
What we are going to show.

\section{Negative Binomial PDF Estimation}
\label{sec:pdfEstimation}

To predict an individual PDF by means of a parametric approach, one has to rely on a model
assumption about the underlying distribution of the random variable to be predicted. In
the following, we describe such a parametric approach with the model assumption of a
negative binomial probability distribution. For a PDF prediction under a negative binomial
model assumption, there is the need for the estimation of two parameters, for example its
mean and its variance.
In any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding
predicted mean from the mean model as feature in the variance model.

This can be done using two independent models one,  to estimate the mean and the other
for the variance. At least in principle, any method can be used. However, as argued above,
in case of demand forecasting, each prediction is highly specific to the circumstances
in which it is used (such as product ID, day and store location) and may
depend on  a multitude of describing variables (features).
Machine learning algorithms are ideally suited for this task and in the
following we will use the ``Cyclic Boosting'' algorithm \cite{Wick2019}.
The major benefits of this algorithm are that it is not only extremly performant in practical 
applications of demand forecasting at large scale, but also fully explainable. Traditional machine
learning algorithms are typically ``black box'' approaches where the individual decision cannot be 
explained. Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
The main idea of this algorithm is the following: When used for regression,
the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following
way:
\begin{equation} \label{eqn:cb}
\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}
The parameters $f^k_j$ are the model parameters that are determined from 
features $j$. Each feature is discretisized appropriately into $k$ bins to reflect the 
specific behaviour of the feature. The global mean $\mu$ is determined from all observed
target values $y$ observed in the data. The factors  $f^k_j$ are determined
iteratively until convergence is reached and regularization techniques are applied
to avoid overtraining and improve the generalization ability of the algorithm.
The relative strength of the factors  $f^k_j$ are then directly interpretable in
relation how important a specific feature is for each individual prediction where deviations
from  $f^k_j=1$ indicate high importance.

%One can achieve this with two subsequent, independent machine learning models, the first
%to estimate the mean and the second to estimate the variance. The features may or may not
%differ between the mean and variance estimation models.

The assigned mean and variance predictions can then be used to generate individual
probability density functions according to a functional negative binomial distribution
assumption for each sample.

%While any machine learning algorithm outputting a conditional mean as prediction may be
%used to get an approximate estimate for the mean of an underlying negative binomial
%distribution, it is preferable to choose an algorithm using a negative binomial assumption
%for its inner regularization methods. An example for such a machine learning algorithm is
%Cyclic Boosting in multiplicative regression mode .
%In the upcoming section, we describe a method to predict the variance of a negative
%binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.

\subsection{Variance Estimation by Cyclic Boosting in Negative Binomial Width Mode}

Cyclic Boosting in a modified form of its multiplicative regression mode, namely the
negative binomial width mode presented in the following, can be used for the second machine
learning model to predict the negative binomial variance associated with the mean
predicted by the first model.

For regression, the parametrization of the negative binomial mass function
can be specified as \cite{hilbe2011negative}:

\begin{equation} \label{eqn:nbinom}
\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
\end{equation}
with mean $\mu$ and dispersion parameter $r$. The target variable $y$ takes the values $y = 0, 1, 2, ...$.

The variance estimation is achieved by minimizing the loss function defined in \eqn
\eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative
binomial distribution, with respect to $r_i$ over all input samples $i$, where the mean
values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine
learning model outputting the conditional mean.

\begin{equation} \label{eqn:loss_likelihood}
L(r) = -\mathcal{L}(r) = -\ln \sum_i \mathrm{NB}(y_i; \hat{\mu_i}, r_i)
\end{equation}

As stated in \eqn \eqref{eqn:r} the values $r_i$ are defined by the Cyclic Boosting
model parameters $f^k_j$ for each feature $j$ and bin $k$. For any concrete observation
$i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the
subsequent look-up into which bin this observation falls. The dispersion parameter $r$ is
bound to the interval $[1, \infty]$.

\textcolor{red}{Hm, this part is a bit quick. We should add a sentence to the NB loss 
and how this is motivated and then how this is connected to \eqn \eqref{eqn:r} and how
\eqn \eqref {eqn:cb} from CB comes into play.}
\begin{equation} \label{eqn:r}
r_i = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
\end{equation}

\textcolor{red}{Hm, this part is not quite clear - need to add some details what is meant}
As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an
iterative method corresponding to a cyclic coordinate descent, processing one feature with
all its bins at a time until convergence. Unlike in the original multiplicative regression
mode of Cyclic Boosting, the minimization of the loss function in \eqn
\eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method,
for example a random search. All other advantages of Cyclic Boosting presented in
\cite{cyclicboosting}, like for example individual explainability of predictions, remain
valid for its negative binomial width mode.

Finally, the variance $\hat{\sigma}^2_i$ can be calculated from the dispersion parameter
$\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}. And so, together with the estimated
mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of
the negative binomial distribution NB for the prediction of an individual observation $i$
are given.

\begin{equation} \label{eqn:variance_r}
\sigma^2 = \mu + \frac{\mu^2}{r}
\end{equation}

\section{Evaluation of PDF Predictions}
\label{sec:pdfEvaluation}

Statistical or machine learning methods that predict full individual probability functions typically lack
quantitative, or at least qualitative, evaluation of the PDF model output to assess its
correctness. In the case of an estimation of the determining parameters of an assumed
functional form for the PDF, assessing the correctness of the PDF model output refers to
the evaluation of the accuracy of the prediction of the different determining parameters,
e.g. mean and variance of a negative binomial distribution, and the validation of the
model assumptions, i.e. the choice of the underlying PDF, by comparing the probability
density estimations to observed data. However, the two methods presented in this section,
namely cumulative distribution function (CDF) histograms and inverse quantile plots, are
neither restricted to the evaluation of parametric PDF estimations nor to a specific
functional form, like a negative binomial, but can be used generically.

\subsection{Qualitative Evaluation of PDF Predictions}

\subsubsection{CDF Histogram}
\label{sec:cdf_histo}

\textcolor{red}{need to explain intuition a bit more, a diagram would be helpful.}
The idea here is that the individual PDF predictions are first transformed into individual
CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
values are then filled to a histogram.

\textcolor{red}{need to explain a bit more what we see and why we expect a uniform distribution.}
Figure \eqref{fig:cdf_histos} illustrates five different CDF histograms. If the PDF
prediction is correct, a flat uniform distribution should occur for the CDF histogram,
like illustrated by the final far-right column of shaded circles.

In the example of a parametric negative binomial PDF estimation of section
\eqref{sec:pdfEstimation}, this would indicate that both the mean and the variance
estimations as well as the choice of a negative binomial distribution as underlying PDF
were correct. Conversely, the first two plots of \fig \ref{fig:cdf_histos} indicate a
bias in the variance estimation and the third and fourth plot indicate a bias in the mean
estimation.
\textcolor{red}{Need to explain why we see that. Would be good to show the respective distributions
(or at least one example) of wher it doesn't fit.}

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/cdf_histos}
\caption{\label{fig:cdf_histos} ...}
\end{center}
\end{figure}

\subsubsection{Inverse Quantile Plot}

\textcolor{red}{We need to explain a lot more here. In particular, outside HEP the profile
plots are not known so we need to explain that first and then explain why we want to use
them here.}
The idea here is that the individual PDF predictions are first transformed into individual
CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
values are then compared (in the sense of higher or lower) to expected quantile values and
accordingly filled to a collection of profile plots, each of which representing one
expected quantile value.

\textcolor{red}{since no reader of this paper will have seen a profile plot before or
know what that is, we need to guide the audience step-by-step to what they should see and why it's
correct and what each column means, how we come to the conclusion, etc.}
\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse
quantile profile plots (each collection comparing to 7 expected quantile values), each
with only one bin for the sake of visualization, for separate sets of exemplary
probability density estimation and observed data combinations. Each shown line represents
the percentage of probability density estimation and observed data combinations for which
the observed data point should be above and below the quantile of the predicted PDF
indicated by that line (\textcolor{red}{median?}for example, the 0.50 line indicates that 50 percent of all
probability density estimation and observed data combinations in a given set of data
should fall above the line, and 50 percent should fall below the line), respectively. The
observation of the number of samples, indicated with shaded circles, that do in fact fall
above and below a particular line, then allows the evaluation of the accuracy of
probability density estimations. For a set of accurate PDF predictions, one expects a
uniform shape, like illustrated by the final far-right column of shaded circles.

\textcolor{red}{need to explain how one can see this}
In the example of a parametric negative binomial PDF estimation of section
\ref{sec:pdfEstimation}, this would indicate that both the mean and the variance
estimations as well as the choice of a negative binomial distribution as underlying PDF
were correct. Conversely, the leftmost two columns of shaded circles illustrate a case for
which the variance estimation has a bias and the center and center-right columns of shaded
circles illustrate a case for which the mean estimation has a bias.

\begin{figure}
\begin{center}
\includegraphics[width=8cm]{../figures/invquant_profiles}
\caption{\label{fig:invquant_profiles} ...}
\end{center}
\end{figure}

The two advantages of this method, as compared to the simpler method presented in section
\ref{sec:cdf_histo}, are that an inverse quantile plot supports the qualitative evaluation
of the predicted individual PDFs not only globally but (1) for different specified
quantiles (potentially hinting to deviations in e.g. the tails of the distributions) and
(2) in dependence of arbitrary variables of the data set (potentially hinting to
deviations in e.g. specific categories of a feature). Two examples for this can be found
in figures \ref{fig:invquant_mean} and \ref{fig:invquant_pg} in the next section.

\subsection{Quantitative Evaluation of PDF Predictions}

The methods described so far allow a detailed qualitative evaluation of PDF predictions.
However, in order to also quantify the quality of the PDF predictions, a measure of the
deviation of the PDF predictions from the optimal outcome given the observed target data
is needed. To achieve this, we compare the CDF histogram of the predicted PDFs with the
uniform distribution and define a prediction accuracy interval between 0 and 1.

Several different methods can be used to compute the deviance between two probability
distributions. Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a
distance function defined between two probability distributions on a given metric space
(also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence}
\cite{kullback1951}, a measure of how one probability distribution diverges from a second
expected probability distribution. The first has the advantage that, unlike the second, it
takes an underlying metric space into account, meaning that it depends on the distance of
potential deviations.

\section{Example: Demand Forecasting}
\label{sec:example}

... predict an individual demand volume for different product-location-date combinations ...
... example showing full prediction and evaluation chain ...

... plot showing individual PDF example ...

%\begin{figure}
%\begin{center}
%\includegraphics[width=8cm]{../figures/pdf_example}
%\caption{\label{fig:pdf_example} ...}
%\end{center}
%\end{figure}

... plot showing cdf histo ...

%\begin{figure}
%\begin{center}
%\includegraphics[width=8cm]{../figures/cdf_demand}
%\caption{\label{fig:cdf_demand} ...}
%\end{center}
%\end{figure}

... plot showing invquant profile for mean prediction on X-axis ...

%\begin{figure}
%\begin{center}
%\includegraphics[width=8cm]{../figures/invquant_mean}
%\caption{\label{fig:invquant_mean} ...}
%\end{center}
%\end{figure}

... plot showing invquant profile for product groups on X-axis ...

%\begin{figure}
%\begin{center}
%\includegraphics[width=8cm]{../figures/invquant_pg}
%\caption{\label{fig:invquant_pg} ...}
%\end{center}
%\end{figure}

...

\bibliography{paper}
\bibliographystyle{ieeetr}

%%
%% Appendices
%%
\appendix

\section{Profile Histograms}
In many cases, scatter plots are used to study the behaviour of two distributions or sets of data points visually. 
However, even for moderate amount of data, this approach becomes quickly difficult.
To illustrate this, a sample of $(x,y)$ data points was obtained in the following way:
The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian
distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$.
\fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{../figures/marginal} \includegraphics[scale=0.5]{../figures/scatter} 
\caption{\label{fig:scatter} Marginal distribution and scatter plot of variables $X$ and $Y$.}
\end{center}
\end{figure}

Although the simple linear correlation between $X$ and $Y$ is apparent in the scatter plot, finer details are not
visible and it is easy to imagine that a more complex relationship is difficult to discern.
Profile histograms are specifically designed to address this shortcomming. Intuitively, profile
histograms are a one-dimensional representation of the two-dimensional scatter plot and are obtained 
in the following way: The variable on the $x$ axis is discretizised into a suitable range of bins.
The exact choice of binning depends on the problem at hand. One can for example choose equidistant bins
in the range of the $x$ axis or non-equidistant bins such that each bin contains the same number of observations.
Then within each bin of the variable $X$, the a location and dispersion metric is calculated for the variable $Y$.
This means that the bin-borders on the $X$ axis are used as constraints on the variable $Y$ and with these
conditions applied, for example the sample mean of the selected $y$ values as well as the standard deviation
are calculated.
These location and dispersion metrics in each bin of $X$ are used to illustrate the behaviour of the variable $Y$
as the values of the variable $X$ change from bin to bin.
The resulting profile histogram is shown in \fig \ref{fig:profile}. This one-dimensional representation
allows to understand even a complex relationship between two variables visually. Note
that due to few data points at the edges of the distributions the profile histogram is expected
to show visual artifacts in the corresponding regions.
 \begin{figure}
\begin{center}
\includegraphics[scale=0.5]{../figures/profile}
\caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}
\end{center}
\end{figure}

\end{document}

### 2020-05-07 code/california_housing.py
added code for example

@@ -1,0 +1,375 @@
+"""
+Cond-dist predictions for the California Housing dataset.
+
+Source and problem description:
+
+* http://www.dcc.fc.up.pt/~ltorgo/Regression/cal_housing.html
+* Hastie, The Elements of Statistical Learning, 10.14.1
+
+
+Why are feature bins filled with **unnecessarily heterogeneous data** harmful
+in prediction projects?
+
+* Many machine-learning algorithms such as Neurobayes without internal boost
+  can only consider axis-parallel cuts through the feature space.
+
+  Often target averages inside these subspaces are used for the prediction,
+  in particular averages over such heterogeneous feature bins.
+
+* Such averages may not be meaningful or even **misleading**. For example, if
+  you lumped together apples and peas, then the average size is neither a good
+  predictor for apples nor for peas.
+
+  Taking the median is not much better.
+
+* Heterogeneous feature bins often lead to the famous **XOR problem** in
+  machine learning which is notoriously hard to solve for many
+  machine learning algorithms.
+
+* In **power-law distributions**, taking averages is very misleading, e.g. if
+  you lump together Harry Potter books with specialist books about twelve tone
+  music, then taking the average sales leads to very bad predictions.
+"""
+from __future__ import print_function, division
+
+import numpy as np
+import pandas as pd
+import cPickle
+
+import matplotlib.pylab as plt
+import matplotlib.ticker as ticker
+
+import sklearn.datasets
+import sklearn.cross_validation
+import sklearn.neighbors
+import sklearn.ensemble
+
+from nbpy import cyclic_boosting, pipeline, x_transformer, binning
+from nbpy.cyclic_boosting.plots import plot_analysis
+from nbpy import flags, meta_estimator, on_cols
+from nbpy.smoothing.onedim import IsotonicRegressor
+
+from nbpy.ext import predict_as_transform, plots
+
+
+USE_PICKLE = False
+GRADBOOST = False
+
+
+def get_data(test_size):
+    """Get the data for training and test.
+
+    For the parameters, see :func:`sklearn.cross_validation.train_test_split`.
+
+    :return: ``X_train, X_test, y_train, y_test``, the X's being Pandas
[…]

### 2020-05-07 code/kaggle_demand.py

@@ -1,0 +1,228 @@
+import sys
+
+import pandas as pd
+import numpy as np
+import datetime
+
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+
+from nbpy import binning, pipeline, cyclic_boosting
+from nbpy import flags
+from nbpy.smoothing.onedim import SeasonalSmoother, IsotonicRegressor
+from nbpy.cyclic_boosting.plots import plot_analysis
+
+from IPython import embed
+
+
+def plot_CB(filename, plobs, binner):
+    for i, p in enumerate(plobs):
+        plot_analysis(
+            plot_observer=p,
+            file_obj="plots/" + filename + "_{}".format(i), use_tightlayout=False,
+            binners=[binner]
+        )
+
+
+def plot_factors(est, X):
+    factors = {}
+    for feature in est.ests[-1][1].features:
+        ft = "" if feature.feature_type is None else "_{}".format(
+            feature.feature_type)
+
+        factors[
+            "_".join(feature.feature_group) + ft
+            ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False)) - 1.
+    df_factors = pd.DataFrame(factors)
+
+    ax = df_factors.loc[[0,10000,50000]].plot(kind='barh', title="factors", legend=True, fontsize=15)
+    fig = ax.get_figure()
+    ax.axvline(0, color="gray")
+    ax.set_yticks([])
+
+    def setfactorformat(x, pos):
+        return '%1.2f' % (x + 1)
+    ax.xaxis.set_major_formatter(ticker.FuncFormatter(setfactorformat))
+
+    fig.savefig('plots/factors.pdf')
+
+
+def plot_timeseries(df, suffix, title=''):
+    plt.figure()
+    df.index = df['date']
+    df['y'].plot(style='r', label="sales")
+    df['yhat'].plot(style='b', label="prediction")
+    plt.legend()
+    plt.title(title)
+    plt.tight_layout()
+    plt.savefig('plots/ts_{}.png'.format(suffix))
+
+
+def plotting(df, suffix=''):
+    df = df[['y', 'yhat', 'item', 'store', 'date']]
+
+    ts_data = df.groupby(['date'])[['y', 'yhat']].sum().reset_index()
+    plot_timeseries(ts_data, 'full' + suffix, 'all')
+
[…]

### 2020-06-09 LaTex/paper.tex
more literature and explanations everywhere

@@
 
-%\title{Prediction and Evaluation of individual Probability Density Functions with Machine Learning}
 \title{Demand Forecasting of individual Probability Density Functions with Machine Learning}
@@
 
+PDF
 \section{Introduction}
 
-Demand forecasting is one of the main challenges for retailers and at the core of business
-operations. Due to its stochastic nature, demand is difficult to forecast as it depends on
-many influencing factors and the realized demand can be interpreted as a random variable
-that is described by an appropriate probability density function (PDF). In order to make
-operational decisions, an optimal point estimator has to be defined that  can be used to
-derive ordering decisions used in the replenishment process of the retailer. Demand
-estimation is further complicated by the fact that retailers typically only observed
-realized sales and not the actual demand, in case the demand exceeds the current stock
-level, the data become censored. This decision process is complicated by a range of
-factors: Even in the case of accurate  demand forecasts, the decision maker has to balance
-conflicting metrics to reach an optimal decision: Ordering to few items may result in
-stockout situations resulting in unrealized demand and unsatisfied customers. Ordering too
-many items results in excess inventory which increases transport and storage costs and, in
-the case of perishable goods, excessive waste as spoilt items need to be disposed of at
-additional cost in addition to  the environmental impact. This situation is particularly
-noticeable in the so called `'ultra-fresh'' category which includes items such as bakery
-products, ready-meals, fresh diary products or certain meat products such as ground meat.
-These items typically have a shelf-life from less than a business day to a few business
-days at most with a continuous spectrum  depending on the exact item. In many situations,
-additional constraints have to be considered to reach an optimal ordering decision:
-Delivery cycles of items may vary depending on the type of item and the wholesaler or
-manufacturer from which  they are procured. Retailers also operate at a given service
-level to guarantee that a certain level of demand can be fulfilled. The exact service
-level typically depends on the overall business policy of the retailer and may also depend
-on individual products, ranging from ``never-out-of-stock'' items to a service level
-exceeding e.g. 90\%.
+Demand forecasting is one of the main challenges for retailers and at the core of business operations.
+Due to its stochastic nature, demand is difficult to forecast as it depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
+In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions used in the replenishment process of the retailer.
+Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.
+The ordering decision process is complicated by a range of factors: Even in the case of accurate  demand forecasts, the decision maker has to balance conflicting metrics to reach an optimal decision: Ordering to few items may result in stockout situations leading to unrealized demand and unsatisfied customers.
+Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
+This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
+These items typically have a shelf-life of less than a business day to a few business days at most, with a continuous spectrum depending on the exact item.
+In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured.
+Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled.
+The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g.
+90\%.
 
-The availability of Big Data allows capturing, storing and processing a vast amount of
-data associated with demand such as historic sales records,  information about promotional
-events or advertisements, pricing information, local weather at retail locations, seasonal
-information  as well as a wide range of further variables. Modern machine learning
-algorithms can then be used to predict the per-item demand distribution, corrected for
-censored data from which an optimal point estimator can be derived to be used in the
-subsequent ordering decision. It is important to note that demand as a random variable is
-not identically and independently distributed (i.i.d.): While the probability distribution
-describing the demand can be attributed to a given family or parametrization, the exact
-parameters vary: Seasonal effects, finite life cycles of products and the introduction of
-new products influence the demand distribution, as well as the local weather at the retail
-location or the retail location itself in terms of size, assortment range, customer
-diversity and other factors. The retailers themselves also actively influence demand by
-using advertisements to highlight products, offering rebates or discounts for specific
-products as well as pursuing an active pricing strategy. This means that while we can
-generally assume that demand follows a specific type of probabilty distribution, its
-parameters are unique to the  instance for which an estimate is required. For example, the
-probability distribution governing the demand of a particular item is specific to the
-item, date and retail location for which the forecast is made and depends on a wide range
-of further influencing factors.
+The availability of Big Data allows capturing, storing and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information as well as a wide range of further variables.
+Modern machine learning algorithms can then be used to predict the per-item demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision.
+It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.).
+While the probability distribution describing the demand can be attributed to a given family or parametrization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors.
+The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy.
+This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the  instance for which an estimate is required.
+For example, the probability distribution governing the demand of a particular item is specific to the item, date and retail location for which the forecast is made and depends on a wide range of further influencing factors.
 
-The remainder of the paper is organized as follows:
-We first review the relevant literature and existing work in sec. \ref{sec:LitRev}.
-We then describe our method to predict individual negative binomial PDFs by means of a
-parametric approach including two distinct machine learning models for mean of variance in
-sec. \ref{sec:pdfEstimation}.
-After that we describe methods for the qualitative and quantitative evaluation of PDF
-predictions in sec. \ref{sec:pdfEvaluation}.
-Finally, we present a demand forecasting example to show an application of our methods in
-sec. \ref{sec:example}.
+The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec.
+\ref{sec:LitRev}.
+We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec.
+\ref{sec:pdfEstimation}.
+After that we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec.
+\ref{sec:pdfEvaluation}.
+And finally, we present a demand forecasting example to show an application of our methods in sec.
+\ref{sec:example}.
 
@@
 
-Inventory management offers a rich theory and the extensive body of research can be
-broadly grouped into the following two categories where the inventory control problem is
-either based on some knowledge of the underlying demand  distribution or an integrated
-approach that seeks to map directly from the available data (historic sales records and 
-further variables) to the ordering decision. The latter approach is taken e.g. in
-\cite{beutel2012safety,ban2019big,bertsimas2020predictive} and aims to avoid estimating
-the underlying probability distribution. Although this approach seems preferable since it
-avoids determining the full demand distribution and results directly in the desired
-operational decision (order quantity), it faces several drawbacks. First of all, the full
-probability distribution for the demand of an specific item at a given sales location and
-business day includes all available information including the uncertainty of the modelled
-demand. This can be used to simulate the performance on a per-item level and e.g. optimize
-the impact on business strategy decisions on conflicting metrics such as stock out- and
-waste-rate. Additionally, having the full demand probability distribution available allows
-for an accute assessment of the forecast quality at all quantiles of the distribution,
-including the often extensive tails of the distribution, as well as the analysis of
-long-term  effect as the demand predictions including all future planned advertisements or
-price-changes can be included into long-term forecasts. Furthermore, deriving the ordering
-decision  directly couples the demand process with the complex delivery cycles and
-constraints, keeping the steps separate allows greater operational flexibility and reduces
-the impact of changing manufacturers or wholesalers, as well as  allowing a quick response
-to changes in the delivery schedule wihtout having to re-compute the implicit demand
-underlying the ordering decision. Finally, long-term demand forecasts may be shared with
-other business units or external vendors  and wholesalers to ease their planning for the
-production and supply-chain processes upstream of the retailer.
+Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision.
+This approach is often referred to as "data-driven newsvendor'' and discussed  e.g.
+in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
+It aims to avoid estimating the underlying probability distribution and use the available data to derive the operational decisions directly.
+An overview of a range approaches can also be found in \cite{huber2019data}.
 
-In contrast, more traditional inventory control systems rely on the knowledge of the
-demand distribution in one form or another.  see e.g. \cite{silver1998} for an overview.
-In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored
-at regular intervals and orders are dispatched once the inventory level reaches a minimal
-value $s$. In case of linear holding and shortage costs, such policies are optimal
-\cite{Scarf1959}, although perishable goods pose more challenges, see e.g.
-\cite{Nahmias1973,Nahmias1975,nahmias1978}. Additionally, service level constraints can be
-included in these kind of inventory control systems \cite{minner2010periodic}. Perishable
-goods are well described by the `'newsvendor-problem'' \cite{Edgeworth} where in the
-simplest case all stock perishes at the end of the selling period (e.g. a business day). 
-For a detailed review of the newsvendor problem see e.g. \cite{Khouja1999537}. Assuming
-linear underage and overage costs $b,h >0$, the optimal quantile
-$q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated
-exactly. The main objective in any of these approaches is to determine the underlying
-demand distribution. The simplest approach is to just use the observed sales events and
-forecast these as a time-series  (see e.g. \cite{alwan2016}) or via sample average
-approximation (SAA). (see e.g. \cite{shapiro2014} for an overview over SAA). However,
-these approaches do not make use of any data apart from the sales record themselves,
-although we know that many variables such as price, advertisements, etc. are highly
-correlated with demand. Additionally, it is critical that we indeed reconstruct a demand
-distribution, hence a simple point-estimator as provided by the most common statistical
-techniques and machine-learning approaches will not suffice. Additionally, demand is not
-identically and independently distributed (i.i.d.) but depends not only on external
-factors such as season, weather, product life-cycle, but is also actively changed by the
-retailer by setting a specific price, offering rebates or running advertisements.
-Additionally, the demand implicitly depend on the location of the retail outlet as well as
-the specifics of that location such as product  assortment influencing the choice of
-possible replacement articles and many more. These complications are the main reason we
-cannot treat the replenishment process as $n$ independent newsvendor-type problems.
-Instead, we need to determine the full demand distribution from data, conditional on the
-relevant variables such as date, location and item, taking all auxiliary data such as
-article charactersitics, pricing, advertisements, retail location details, etc. into
-account. This can be done in several ways: First, we can use a neural network
-\cite{Feindt2006190} to learn the distribution from data and return a full distribution
-per item, store and day from which the relevant quantile can be estimated. Similarly,
-using quantile regression \cite{koenker2001} this approach can be implemented in different
-frameworks. Alternatively, one can assume a given demand model and fit the model
-parameters instead of reconstructing the complete distribution. This approach is
-computationally favourable, as fewer parameters need to be estimated compared to the case
-of the full distribution. Empirically, one can determine the best fitting distribution
-from data \cite{adan1995}. However, given the stochastic nature of the demand, such an
-empirically determined distribution is not expected to be stable and prone to sudden
-changes. Instead, the choice of the demand distribution should be motivated by theoretic
-considerations. The discrete demand is typically modelled as a negative binomial
-distribution (NBD), also known as Gamma-Poisson distribution
-\cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}. This
-distribution arises if the Poisson parameter $\mu$ is a random variable itself that
-follows a Gamma distribution. The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$ and
-is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$.
+Although this approach seems preferable, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks.
+First of all, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand.
+This can be used to simulate the performance on a per-item level and e.g.
+optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate.
+Additionally, having the full demand probability distribution available allows for an accurate assessment of the forecast quality of all quantiles of the distribution, including the often extensive tails of the distribution.
+Separating the demand forecast from the operational decisions (i.e.
+calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning or counterfactual analysis to investigate how the demand would have been had the input variables had different values.
+This way, effects of promotions, advertisements or price-changes can be analyzed  with the same model that is used for operational decisions.
+Furthermore, deriving the ordering decision directly couples the demand process with the complex delivery cycles and constraints, while keeping the steps separate allows greater operational flexibility and reduces the impact of changing manufacturers or wholesalers, as well as  allowing a quick response to changes in the delivery schedule without having to re-compute the implicit demand underlying the ordering decision.
+Finally, long-term demand forecasts may be shared with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer.
+From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other.
+For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.
 
-Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be
-determined for each item at each sales location and ordering time, depending on all
-auxiliary data describing article details, retail location and influencing factors such as
-pricing and advertisement information.
+In contrast to the direct mapping the observed data to ordering decisions, more traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g.
+\cite{silver1998} for an overview.
+In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$.
+In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g.
+\cite{Nahmias1973,Nahmias1975,nahmias1978}.
+Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}.
+Perishable goods are well described by the "newsvendor-problem'' \cite{Edgeworth}, where in the simplest case all stock perishes at the end of the selling period (e.g.
+a business day).
+For a detailed review of the newsvendor problem see e.g.
+\cite{Khouja1999537}.
+Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.
+
+The main objective in any of these approaches is to determine the underlying demand distribution.
+The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g.
+\cite{alwan2016}) or via sample average approximation (SAA), see e.g.
+\cite{shapiro2014} for an overview.
+However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand.
+Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e.
+the  calculation of the optimal order quantity.
+
+However, in order to be able to fully optimize the operational decision, it is critical that one indeed reconstructs a full demand distribution.
+This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine-learning approaches, will not suffice.
+Additionally, we need to consider that demand is not i.i.d., but depends on external factors such as season, weather, product life-cycle, and is also actively changed by the retailer by setting a specific price, offering rebates or running advertisements.
+Additionally, the demand implicitly depends on the location of the retail outlet as well as the specifics of that location, such as product  assortment influencing the choice of possible replacement articles and many more.
+These complications are the main reason we cannot treat the replenishment process as $n$ independent newsvendor-type problems.
+
+Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc.
+into account.
+This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution from which an empirical probability distribution can be interpolated.
+Using a dedicated neural network \cite{Feindt2006190}, either the  full probability distribution or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
+Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}.
+This approach is computationally favourable, as fewer parameters need to be estimated compared to the case of the full distribution.
+Empirically, one can determine the best fitting distribution from data \cite{adan1995}.
+However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes.
+Instead, the choice of the demand distribution should be motivated by theoretic considerations.
+The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}.
+This distribution arises if the Poisson parameter $\mu$ is a random variable itself that follows a Gamma distribution.
+The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$.
+Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
 
 \subsection*{Summary of Contributions}
-What we are going to show.
+
+In the following, we will demonstrate how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer.
+Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
+In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.
+
+Additionally, we show how the cumulative distribution function (CDF) or inverse quantiles can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution.
+This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer.
 
@@
 
-To predict an individual PDF by means of a parametric approach, one has to rely on a model
-assumption about the underlying distribution of the random variable to be predicted. In
-the following, we describe such a parametric approach with the model assumption of a
-negative binomial probability distribution. For a PDF prediction under a negative binomial
-model assumption, there is the need for the estimation of two parameters, for example its
-mean and its variance.
-In any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding
-predicted mean from the mean model as feature in the variance model.
+To predict an individual PDF by means of a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted.
+In the following, we describe such a parametric approach with the model assumption of a negative binomial probability distribution.
+For a PDF prediction under a negative binomial model assumption, there is the need for the estimation of two parameters, for example its mean and its variance.
 
-This can be done using two independent models one,  to estimate the mean and the other
-for the variance. At least in principle, any method can be used. However, as argued above,
-in case of demand forecasting, each prediction is highly specific to the circumstances
-in which it is used (such as product ID, day and store location) and may
-depend on  a multitude of describing variables (features).
-Machine learning algorithms are ideally suited for this task and in the
-following we will use the ``Cyclic Boosting'' algorithm \cite{Wick2019}.
-The major benefits of this algorithm are that it is not only extremly performant in practical 
-applications of demand forecasting at large scale, but also fully explainable. Traditional machine
-learning algorithms are typically ``black box'' approaches where the individual decision cannot be 
-explained. Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
-The main idea of this algorithm is the following: When used for regression,
-the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following
-way:
+This can be done using two independent models, one to estimate the mean and the other for the variance.
+At least in principle, any method can be used.
+However, as argued above, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as product ID, date, and store location) and may depend on  a multitude of describing variables (features).
+Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm.
+The major benefits of this method are that it is not only extremely performant in practical applications of demand forecasting at large scale, but also fully explainable.
+Traditional machine learning algorithms are typically "black box'' approaches, where the individual decision cannot be explained.
+Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
+The main idea of this algorithm is the following: When used for regression, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:
+
 \begin{equation} \label{eqn:cb}
@@
 \end{equation}
-The parameters $f^k_j$ are the model parameters that are determined from 
-features $j$. Each feature is discretisized appropriately into $k$ bins to reflect the 
-specific behaviour of the feature. The global mean $\mu$ is determined from all observed
-target values $y$ observed in the data. The factors  $f^k_j$ are determined
-iteratively until convergence is reached and regularization techniques are applied
-to avoid overtraining and improve the generalization ability of the algorithm.
-The relative strength of the factors  $f^k_j$ are then directly interpretable in
-relation how important a specific feature is for each individual prediction where deviations
-from  $f^k_j=1$ indicate high importance.
 
-%One can achieve this with two subsequent, independent machine learning models, the first
-%to estimate the mean and the second to estimate the variance. The features may or may not
-%differ between the mean and variance estimation models.
+The parameters $f^k_j$ are the model parameters that are determined from features $j$.
+Each feature is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
+The global mean $\mu$ is determined from all target values $y$ observed in the data.
+The factors $f^k_j$ are determined iteratively until convergence is reached and regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm.
+The relative strength of the factors  $f^k_j$ are then directly interpretable in relation to how important a specific feature is for each individual prediction, where deviations from  $f^k_j=1$ indicate high importance.
 
-The assigned mean and variance predictions can then be used to generate individual
-probability density functions according to a functional negative binomial distribution
-assumption for each sample.
+\textcolor{red}{Need for some details on Cyclic Boosting training process}
 
-%While any machine learning algorithm outputting a conditional mean as prediction may be
-%used to get an approximate estimate for the mean of an underlying negative binomial
-%distribution, it is preferable to choose an algorithm using a negative binomial assumption
-%for its inner regularization methods. An example for such a machine learning algorithm is
-%Cyclic Boosting in multiplicative regression mode .
-%In the upcoming section, we describe a method to predict the variance of a negative
-%binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.
+For the PDF estimation, two subsequent, independent Cyclic Boosting models are used, the first to estimate the mean and the second to estimate the variance.
+The features may or may not differ between the mean and variance estimation models.
+But in any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
+The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.
+
+While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.
+
+In the upcoming section, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.
 
@@
 
-Cyclic Boosting in a modified form of its multiplicative regression mode, namely the
-negative binomial width mode presented in the following, can be used for the second machine
-learning model to predict the negative binomial variance associated with the mean
-predicted by the first model.
+Cyclic Boosting in a modified form of its multiplicative regression mode, namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
 
-For regression, the parametrization of the negative binomial mass function
-can be specified as \cite{hilbe2011negative}:
+For regression, the parametrization of the negative binomial mass function can be specified as \cite{hilbe2011negative}:
 
@@
 \end{equation}
+
 with mean $\mu$ and dispersion parameter $r$. The target variable $y$ takes the values $y = 0, 1, 2, ...$.
 
-The variance estimation is achieved by minimizing the loss function defined in \eqn
-\eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative
-binomial distribution, with respect to $r_i$ over all input samples $i$, where the mean
-values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine
-learning model outputting the conditional mean.
+The variance estimation is achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to $r_i$ over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
 
@@
 
-As stated in \eqn \eqref{eqn:r} the values $r_i$ are defined by the Cyclic Boosting
-model parameters $f^k_j$ for each feature $j$ and bin $k$. For any concrete observation
-$i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the
-subsequent look-up into which bin this observation falls. The dispersion parameter $r$ is
-bound to the interval $[1, \infty]$.
+As stated in \eqn \eqref{eqn:r}, the values $r_i$ are defined by the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$.
+For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls.
+The dispersion parameter $r$ is bound to the interval $[1, \infty]$.
 
-\textcolor{red}{Hm, this part is a bit quick. We should add a sentence to the NB loss 
-and how this is motivated and then how this is connected to \eqn \eqref{eqn:r} and how
-\eqn \eqref {eqn:cb} from CB comes into play.}
 \begin{equation} \label{eqn:r}
@@
 
-\textcolor{red}{Hm, this part is not quite clear - need to add some details what is meant}
-As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an
-iterative method corresponding to a cyclic coordinate descent, processing one feature with
-all its bins at a time until convergence. Unlike in the original multiplicative regression
-mode of Cyclic Boosting, the minimization of the loss function in \eqn
-\eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method,
-for example a random search. All other advantages of Cyclic Boosting presented in
-\cite{cyclicboosting}, like for example individual explainability of predictions, remain
-valid for its negative binomial width mode.
+\textcolor{red}{Need for more details.}
 
-Finally, the variance $\hat{\sigma}^2_i$ can be calculated from the dispersion parameter
-$\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}. And so, together with the estimated
-mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of
-the negative binomial distribution NB for the prediction of an individual observation $i$
-are given.
+As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence.
+Unlike in the original multiplicative regression mode of Cyclic Boosting, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search.
+All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
+
+Finally, the variance $\hat{\sigma}^2_i$ can be calculated from the dispersion parameter $\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}.
+And so, together with the estimated mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of the negative binomial distribution NB for the prediction of an individual observation $i$ are given.
 
@@
 
-Statistical or machine learning methods that predict full individual probability functions typically lack
-quantitative, or at least qualitative, evaluation of the PDF model output to assess its
-correctness. In the case of an estimation of the determining parameters of an assumed
-functional form for the PDF, assessing the correctness of the PDF model output refers to
-the evaluation of the accuracy of the prediction of the different determining parameters,
-e.g. mean and variance of a negative binomial distribution, and the validation of the
-model assumptions, i.e. the choice of the underlying PDF, by comparing the probability
-density estimations to observed data. However, the two methods presented in this section,
-namely cumulative distribution function (CDF) histograms and inverse quantile plots, are
-neither restricted to the evaluation of parametric PDF estimations nor to a specific
-functional form, like a negative binomial, but can be used generically.
+Many statistical and most machine learning methods do not provide a full probability density distribution as result.
+Instead, these methods typically predict a single numerical quantity that is then compared to the observed concrete realization of the random variable using metrics such as the mean squared error (MSE),  mean absolute deviation (MAD) or others.
+In the setting of a retailer, the observed quantity is the sales of individual products (denoted by $y$) and most machine learning approaches would then predict a single number $\hat{y}$ as a direct estimate of the sales.
+However, reducing the prediction to a single number does not allow to account for the uncertainty of the prediction or the dynamics of the system.
+Instead, it is imperative to predict the full PDF for each prediction to be able to optimize the subsequent operational decision.
+Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or at least qualitative evaluation methods to assess whether the full distribution has been forecast correctly, in particular in the tails of the distribution.
+
+For an estimation of the determining parameters of an assumed functional form for the PDF, assessing the correctness of the PDF model output refers to the evaluation of the accuracy of the prediction of the different determining parameters.
+In the case of the negative binomial distribution used in this work, we have to verify that mean and variance are determined accurately, as well as checking that the choice of the underlying model can describe the observed data.
+
+In the following, we will show how histograms of the CDF and inverse quantiles can be used to determine the accuracy of the predicted PDFs.
+Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.
 
@@
 
-\subsubsection{CDF Histogram}
+As a first step, we use the inverse quantile function or probability integral transform (PIT), see e.g.
+\cite{Angus1994,casella2002statistical}.
+This offers a first qualitative assessment by looking at the distribution of the quantiles of the predicted density via the CDF.
+In case of a correctly calibrated predicted PDF, this distribution is expected to be uniform and any deviation can be interpreted as a hint that the predicted probability function is not correct \cite{diebold1998vevaluating}.
+
+\subsubsection{Histogram of the CDF}
 \label{sec:cdf_histo}
 
-\textcolor{red}{need to explain intuition a bit more, a diagram would be helpful.}
-The idea here is that the individual PDF predictions are first transformed into individual
-CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
-values are then filled to a histogram.
+The CDF of a PDF $f(x)$ is defined as:
 
-\textcolor{red}{need to explain a bit more what we see and why we expect a uniform distribution.}
-Figure \eqref{fig:cdf_histos} illustrates five different CDF histograms. If the PDF
-prediction is correct, a flat uniform distribution should occur for the CDF histogram,
-like illustrated by the final far-right column of shaded circles.
+\begin{equation}
+\label{eqn:CDF}
+F_X(x) = P(X \le x) = \int_{-\infty}^{x} f_X(x^\prime) dx^\prime
+\end{equation}
 
-In the example of a parametric negative binomial PDF estimation of section
-\eqref{sec:pdfEstimation}, this would indicate that both the mean and the variance
-estimations as well as the choice of a negative binomial distribution as underlying PDF
-were correct. Conversely, the first two plots of \fig \ref{fig:cdf_histos} indicate a
-bias in the variance estimation and the third and fourth plot indicate a bias in the mean
+Here, $F_X(x)$ is the CDF with $\lim_{x \to -\infty}F_X(x) = 0$ and $\lim_{x \to \infty}F_x(x) = 1$.
+The cumulative distribution describes the probability that a the variable has a value smaller than $x$ and intuitively represents the area under $f(x^\prime)$ up to a point $x$.
+
+If the CDF is continuous and strictly increasing, then the inverse of the CDF $F^{-1}(y)$ exists and is a unique real-valued number $x$ for each $y \in [0,1]$ so that we can write $F(x) = y$.
+The latter defines the inverse quantile function because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
+
+\begin{equation}
+Q_\tau = F^{-1}(\tau)
+\end{equation}
+
+Using the example of the normal distribution with $\mathcal{N}(0,1)$ as shown in \fig \ref{fig:PdfCdf}, we can identify the median ($\tau = 0.5$) by first looking at the CDF in the lower part of the figure, look at $y=0.5$ on the $y$-axis and then identify the point on the $x$ axis for both the PDF $f(x)$ and the CDF $F(x)$ that correspond to the quantile $\tau$.
+In the case of the normal distribution, this is of course the central value at zero.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=8cm]{../figures/PdfCdf}
+\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution funtion of a normal distrubution.}
+\end{center}
+\end{figure}
+
+We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e.
+$F:t \to s$.
+Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF.
+This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s\in [0,1]$ in which $s$ is defined and $s$ can be interpreted as the cumulative distribution of its own probability distribution:
+
+\begin{equation}
+s = G(s) = \int_{-\infty}^{s} g(s^\prime) ds^\prime
+\end{equation}
+
+In case of discrete probability functions such as the negative binomial function, the same argument still holds but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
+\cite[p.
+54]{casella2002statistical}.
+In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the inverse quantiles or values of the CDF is filled using random numbers according to the intervals of the CDF.
+For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram is randomly chosen in the interval $[0,0.75]$.
+Proceeding similarly for all other observed values, the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.
+
+A histogram of the quantiles (or PIT values) of each prediction is therefore expected to be uniformly distributed in $[0,1]$ if the predicted probability distribution $f(x)$ is correctly calibrated.
+
+This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of quantiles (or PIT values) for five different cases.
+If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution as shown in the last part of the figure.
+The first two plots of \fig \ref{fig:cdf_histos} indicate a bias in the variance estimation and the third and fourth plot indicate a bias in the mean
 estimation.
-\textcolor{red}{Need to explain why we see that. Would be good to show the respective distributions
-(or at least one example) of wher it doesn't fit.}
 
@@
 \subsubsection{Inverse Quantile Plot}
+Evaluating the full probability function in the simplest case, where we have only one model that can be used to predict all observations, is straight-forward: We could fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
+Then, we could directly compare the model curve with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
+In practical applications however, we have a large number of effective prediction models, since although we always use the same model parametrization, such as the negative binomial distribution, its parameters have to be determined at the required level of granularity.
+For example, for daily orders, we need to predict the parameters of the negative binomial distribution for each location, sales day, and product.
+Unlike the simple case discussed initially, where we had many observations to compare the prediction model to, we now have just one observation per prediction, meaning that we cannot use statistical tests directly.
+Instead we use the fact that a correctly calibrated, predicted PDF should return a uniform cumulative density distribution {\em regardless} of the shape of the predicted distribution.
 
-\textcolor{red}{We need to explain a lot more here. In particular, outside HEP the profile
-plots are not known so we need to explain that first and then explain why we want to use
-them here.}
-The idea here is that the individual PDF predictions are first transformed into individual
-CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
-values are then compared (in the sense of higher or lower) to expected quantile values and
-accordingly filled to a collection of profile plots, each of which representing one
-expected quantile value.
+We therefore proceed as follows: First, each individual predicted probability density distribution is transformed to its cumulative density distribution.
+In the example of the granularity from above, we would then have a CDF for each predicted PDF for each product, sales day, and sales location.
+The CDF values, also known as quantiles, of the corresponding truth values are then compared (in the sense of higher or lower) to expected quantile values and accordingly filled to a collection of profile plots, each of which representing one expected quantile value.
 
-\textcolor{red}{since no reader of this paper will have seen a profile plot before or
-know what that is, we need to guide the audience step-by-step to what they should see and why it's
-correct and what each column means, how we come to the conclusion, etc.}
-\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse
-quantile profile plots (each collection comparing to 7 expected quantile values), each
-with only one bin for the sake of visualization, for separate sets of exemplary
-probability density estimation and observed data combinations. Each shown line represents
-the percentage of probability density estimation and observed data combinations for which
-the observed data point should be above and below the quantile of the predicted PDF
-indicated by that line (\textcolor{red}{median?}for example, the 0.50 line indicates that 50 percent of all
-probability density estimation and observed data combinations in a given set of data
-should fall above the line, and 50 percent should fall below the line), respectively. The
-observation of the number of samples, indicated with shaded circles, that do in fact fall
-above and below a particular line, then allows the evaluation of the accuracy of
-probability density estimations. For a set of accurate PDF predictions, one expects a
-uniform shape, like illustrated by the final far-right column of shaded circles.
+\textcolor{red}{Need for more explanations}
+
+\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse quantile profile plots (each collection comparing to 7 expected quantile values), each with only one bin for the sake of visualization, for separate sets of exemplary probability density estimation and observed data combinations.
+Each shown line represents the percentage of probability density estimation and observed data combinations for which the observed data point should be above and below the quantile of the predicted PDF indicated by that line (for example, the 0.50 line, corresponding to the median, indicates that 50 percent of all probability density estimation and observed data combinations in a given set of data should fall above the line, and 50 percent should fall below the line), respectively.
+The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of probability density estimations.
+For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles.
 
 \textcolor{red}{need to explain how one can see this}
-In the example of a parametric negative binomial PDF estimation of section
-\ref{sec:pdfEstimation}, this would indicate that both the mean and the variance
-estimations as well as the choice of a negative binomial distribution as underlying PDF
-were correct. Conversely, the leftmost two columns of shaded circles illustrate a case for
-which the variance estimation has a bias and the center and center-right columns of shaded
-circles illustrate a case for which the mean estimation has a bias.
+
+In the example of a parametric negative binomial PDF estimation of section \ref{sec:pdfEstimation}, this would indicate that both the mean and the variance estimations as well as the choice of a negative binomial distribution as underlying PDF were correct.
+Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias and the center and center-right columns of shaded circles illustrate a case for which the mean estimation has a bias.
 
@@
 
-The two advantages of this method, as compared to the simpler method presented in section
-\ref{sec:cdf_histo}, are that an inverse quantile plot supports the qualitative evaluation
-of the predicted individual PDFs not only globally but (1) for different specified
-quantiles (potentially hinting to deviations in e.g. the tails of the distributions) and
-(2) in dependence of arbitrary variables of the data set (potentially hinting to
-deviations in e.g. specific categories of a feature). Two examples for this can be found
-in figures \ref{fig:invquant_mean} and \ref{fig:invquant_pg} in the next section.
+The two advantages of this method, as compared to the simpler method presented in section \ref{sec:cdf_histo}, are that an inverse quantile plot supports the qualitative evaluation of the predicted individual PDFs not only globally but (1) for different specified quantiles (potentially hinting to deviations in e.g.
+the tails of the distributions) and (2) in dependence of arbitrary variables of the data set (potentially hinting to deviations in e.g.
+specific categories of a feature).
+Two examples for this can be found in figures \ref{fig:invquant_mean} and \ref{fig:invquant_pg} in the next section.
 
@@
 The methods described so far allow a detailed qualitative evaluation of PDF predictions.
-However, in order to also quantify the quality of the PDF predictions, a measure of the
-deviation of the PDF predictions from the optimal outcome given the observed target data
-is needed. To achieve this, we compare the CDF histogram of the predicted PDFs with the
-uniform distribution and define a prediction accuracy interval between 0 and 1.
+However, in order to also quantify the quality of the PDF predictions, a measure of the deviation of the PDF predictions from the optimal outcome given the observed target data is needed.
+To achieve this, we compare the CDF histogram of the predicted PDFs with the uniform distribution and define a prediction accuracy interval between 0 and 1.
 
-Several different methods can be used to compute the deviance between two probability
-distributions. Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a
-distance function defined between two probability distributions on a given metric space
-(also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence}
-\cite{kullback1951}, a measure of how one probability distribution diverges from a second
-expected probability distribution. The first has the advantage that, unlike the second, it
-takes an underlying metric space into account, meaning that it depends on the distance of
-potential deviations.
+Several different methods can be used to compute the deviance between two probability distributions.
+Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a distance function defined between two probability distributions on a given metric space (also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, a measure of how one probability distribution diverges from a second expected probability distribution.
+The first has the advantage that, unlike the second, it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
+
+\textcolor{red}{Need to add some formulas.}
 
@@
 
-... predict an individual demand volume for different product-location-date combinations ...
-... example showing full prediction and evaluation chain ...
+\textcolor{red}{... predict an individual demand volume for different product-location-date combinations ...}
 
-... plot showing individual PDF example ...
+\textcolor{red}{... example showing full prediction and evaluation chain ...}
+
+\textcolor{red}{... plot showing individual PDF example ...}
 
@@
 
-... plot showing cdf histo ...
+\textcolor{red}{... plot showing cdf histo ...}
 
@@
 
-... plot showing invquant profile for mean prediction on X-axis ...
+\textcolor{red}{... plot showing invquant profile for mean prediction on X-axis ...}
 
@@
 
-... plot showing invquant profile for product groups on X-axis ...
+\textcolor{red}{... plot showing invquant profile for product groups on X-axis ...}
 
@@
 %\end{figure}
-
-...
 
@@
 \section{Profile Histograms}
+
 In many cases, scatter plots are used to study the behaviour of two distributions or sets of data points visually.
 However, even for moderate amount of data, this approach becomes quickly difficult.
-To illustrate this, a sample of $(x,y)$ data points was obtained in the following way:
-The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian
-distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$.
-\fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs. $y$.
+To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$.
+\fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs.
+$y$.
 
@@
 
-Although the simple linear correlation between $X$ and $Y$ is apparent in the scatter plot, finer details are not
-visible and it is easy to imagine that a more complex relationship is difficult to discern.
-Profile histograms are specifically designed to address this shortcomming. Intuitively, profile
-histograms are a one-dimensional representation of the two-dimensional scatter plot and are obtained 
-in the following way: The variable on the $x$ axis is discretizised into a suitable range of bins.
-The exact choice of binning depends on the problem at hand. One can for example choose equidistant bins
-in the range of the $x$ axis or non-equidistant bins such that each bin contains the same number of observations.
+Although the simple linear correlation between $X$ and $Y$ is apparent in the scatter plot, finer details are not visible and it is easy to imagine that a more complex relationship is difficult to discern.
+Profile histograms are specifically designed to address this shortcoming.
+Intuitively, profile histograms are a one-dimensional representation of the two-dimensional scatter plot and are obtained  in the following way: The variable on the $x$ axis is discretizised into a suitable range of bins.
+The exact choice of binning depends on the problem at hand.
+One can for example choose equidistant bins in the range of the $x$ axis or non-equidistant bins such that each bin contains the same number of observations.
 Then within each bin of the variable $X$, the a location and dispersion metric is calculated for the variable $Y$.
-This means that the bin-borders on the $X$ axis are used as constraints on the variable $Y$ and with these
-conditions applied, for example the sample mean of the selected $y$ values as well as the standard deviation
-are calculated.
-These location and dispersion metrics in each bin of $X$ are used to illustrate the behaviour of the variable $Y$
-as the values of the variable $X$ change from bin to bin.
-The resulting profile histogram is shown in \fig \ref{fig:profile}. This one-dimensional representation
-allows to understand even a complex relationship between two variables visually. Note
-that due to few data points at the edges of the distributions the profile histogram is expected
-to show visual artifacts in the corresponding regions.
+This means that the bin-borders on the $X$ axis are used as constraints on the variable $Y$ and with these conditions applied, for example the sample mean of the selected $y$ values as well as the standard deviation are calculated.
+These location and dispersion metrics in each bin of $X$ are used to illustrate the behaviour of the variable $Y$ as the values of the variable $X$ change from bin to bin.
+The resulting profile histogram is shown in \fig \ref{fig:profile}.
+This one-dimensional representation allows to understand even a complex relationship between two variables visually.
+Note that due to few data points at the edges of the distributions the profile histogram is expected to show visual artifacts in the corresponding regions.
+
  \begin{figure}

### 2020-06-10 LaTex/paper.tex
added some changes in 4th section

@@
 
-\textcolor{red}{Need for more explanations}
+Cyclic Boosting addresses the challenge of prediction explainability on a fundamental level. Rather than relying on black box approaches, individual predictions $\hat{y_i}$ for single
+observations should not only be accurate but also explainable. Strength and significance of
+each feature variable contributing to the prediction can be explained using profile plots.
+
+\begin{figure}[htp]
+    \centering
+    \includegraphics[width=8cm]{Profile_plots.png}
+    \caption{}
+     \label{fig:Profile Plots}
+\end{figure}
+
+Figure 3 shows analysis of the profile plots for feature variable ”item” after the
+final iteration.
+The shown data points indicate mean values of prediction $\hat{y}$ and ground truth y for each bin divided by the global mean (prediction hardly visible due to good agreement).
+Note that the smoothed factors shown here are influenced by correlations to all other features in the model as well.
+This visualization directly
+indicates possible deviations from the optimal fit results
+in the different bins. Here, no significant deviations are
+present across the whole range of values. Furthermore,
+the smoothed values of the factors, i.e. the actual fitted
+parameters of the model, are shown. These differ from the
+mean values of the target and prediction in the different
+bins divided by the global mean due to correlations with
+other features. Right hand side plot in Figure 3 indicate values of prediction $\hat{y}$ and ground truth y where maximum points should fall on diagonal.
+The idea here is that the individual PDF predictions are first transformed into individual
+CDF predictions and the CDF values, also known as quantiles, of the corresponding truth
+values are then compared (in the sense of higher or lower) to expected quantile values and
+accordingly filled to a collection of profile plots, each of which representing one
+expected quantile value.
 
@@
 For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles.
-
-\textcolor{red}{need to explain how one can see this}
-
-In the example of a parametric negative binomial PDF estimation of section \ref{sec:pdfEstimation}, this would indicate that both the mean and the variance estimations as well as the choice of a negative binomial distribution as underlying PDF were correct.
+Here all shaded circles are exactly falling on the line which indicates that both the mean and the variance estimations as well as the choice of a negative binomial distribution as underlying PDF were correct.
 Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias and the center and center-right columns of shaded circles illustrate a case for which the mean estimation has a bias.
@@
 Several different methods can be used to compute the deviance between two probability distributions.
-Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a distance function defined between two probability distributions on a given metric space (also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, a measure of how one probability distribution diverges from a second expected probability distribution.
+Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a distance function defined between two probability distributions on a given metric space (also known as earth mover’s distance), which is defined by
+
+\begin{equation}
+{\displaystyle W_{p}(\mu ,\nu)=\left(\inf \operatorname {E} {\big [}d(X,Y)^{p}{\big ]}\right)^{1/p},}
+\end{equation}
+
+(where ${{E}[Z]}$ denotes the expected value of a random variable Z and the infimum is taken over all joint distributions of the random variables X and Y with marginals ${\mu}$ and ${\nu}$  respectively) and the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, a measure of how one probability distribution diverges from a second expected probability distribution.
+For  probability distributions P and Q defined on the same probability space, ${\displaystyle {\mathcal {X}}}$, the Kullback–Leibler divergence from Q to P is defined as
+
+\begin{equation}
+{\displaystyle D_{\text{KL}}(P\parallel Q)=\sum _{x\in {\mathcal {X}}}P(x)\log \left({\frac {P(x)}{Q(x)}}\right)}
+\end{equation}
+
 The first has the advantage that, unlike the second, it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
-
-\textcolor{red}{Need to add some formulas.}
 

### 2020-06-12 LaTex/paper.tex
adapted formulas for distance measures

--- 2020-06-09 LaTex/paper.tex
@@
 
-\textcolor{red}{Need for more explanations}
-
 \fig \ref{fig:invquant_profiles} illustrates five different collections of inverse quantile profile plots (each collection comparing to 7 expected quantile values), each with only one bin for the sake of visualization, for separate sets of exemplary probability density estimation and observed data combinations.
@@
 For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles.
-
-\textcolor{red}{need to explain how one can see this}
-
-In the example of a parametric negative binomial PDF estimation of section \ref{sec:pdfEstimation}, this would indicate that both the mean and the variance estimations as well as the choice of a negative binomial distribution as underlying PDF were correct.
-Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias and the center and center-right columns of shaded circles illustrate a case for which the mean estimation has a bias.
+Here, all shaded circles exactly fall on the line, indicating that both mean and variance estimations, as well as the choice of a negative binomial distribution as underlying PDF, were correct.
+Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias, and the center and center-right columns of shaded circles illustrate cases for which the mean estimation has a bias.
 
@@
 To achieve this, we compare the CDF histogram of the predicted PDFs with the uniform distribution and define a prediction accuracy interval between 0 and 1.
+Several different methods can be used to compute the deviance between two probability distributions.
+Good choices are the \textbf{1st Wasserstein distance} \cite{olkin1982}, the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, and the \textbf{Jensen-Shannon divergence}  \cite{dagan1997}.
 
-Several different methods can be used to compute the deviance between two probability distributions.
-Good choices are the \textbf{Wasserstein metric} \cite{olkin1982}, a distance function defined between two probability distributions on a given metric space (also known as earth mover’s distance), and the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, a measure of how one probability distribution diverges from a second expected probability distribution.
-The first has the advantage that, unlike the second, it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
+The 1st Wasserstein distance, also known as earth mover'€™s distance (EMD), represents a distance function defined between two probability distributions on a given metric space, and for our purposes here can defined by
 
-\textcolor{red}{Need to add some formulas.}
+\begin{equation}
+\text{EMD}(P, Q) = 2 \cdot \frac{\sum_{k=1}^N |F_P(x_k) - F_Q(x_k)|}{N},
+\end{equation}
+
+where $F_P(X)$ and $F_Q(X)$ are the CDFs of the two PDFs $P(X)$ and $Q(X)$, respectively, and $x_k$ denotes the average value of $X$ in bin $k$, with $X$ being divided in $N$ bins.
+
+The Kullback-Leibler divergence is a measure of how one probability distribution diverges from a second expected probability distribution.
+For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-“Leibler divergence from Q to P is defined as
+
+\begin{equation}
+D_{\text{KL}}(P \parallel Q) = \sum _{k=1}^N P(x_k) \log \left({\frac{P(x_k)}{Q(x_k)}}\right),
+\end{equation}
+
+where $\log$ can be either base-2 or natural logarithm.
+
+The Jensen-Shannon divergence (JSD) can be seen as a symmetrized and smoothed version of the Kullback-Leibler divergence:
+
+\begin{equation}
+D_{\text{JSD}}(P \parallel Q) = 0.5  \cdot (D_{\text{KL}}(P \parallel M) + D_{\text{KL}}(P \parallel M)),
+\end{equation}
+
+where $M = 0.5  \cdot (P + Q)$.
+
+The EMD has the advantage over the mentioned divergences that it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
 

### 2020-07-20 LaTex/paper.tex
added abstract, conclusion, CB recap

@@
 \usepackage{graphicx}
-\usepackage{subfigure}
+%\usepackage{subfigure}
 \usepackage{booktabs} % for professional tables
@@
 \usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
+%\usepackage{algorithmic}
 \usepackage{graphicx}
@@
 \begin{abstract}
-... abstract ...
+Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes.
+While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying random distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
+We present an extension to our earlier work on a supervised machine learning method called Cyclic Boosting to predict full individual probability density functions in form of negative binomial distributions, that are fully explainable on the individual prediction level.
+Also, we describe new techniques for both qualitative and quantitative evaluation of predicted probability density functions by comparison with observed data.
 \end{abstract}
@@
 
-Although this approach seems preferable, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks.
+Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks.
 First of all, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand.
@@
 optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate.
-Additionally, having the full demand probability distribution available allows for an accurate assessment of the forecast quality of all quantiles of the distribution, including the often extensive tails of the distribution.
-Separating the demand forecast from the operational decisions (i.e.
-calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning or counterfactual analysis to investigate how the demand would have been had the input variables had different values.
-This way, effects of promotions, advertisements or price-changes can be analyzed  with the same model that is used for operational decisions.
-Furthermore, deriving the ordering decision directly couples the demand process with the complex delivery cycles and constraints, while keeping the steps separate allows greater operational flexibility and reduces the impact of changing manufacturers or wholesalers, as well as  allowing a quick response to changes in the delivery schedule without having to re-compute the implicit demand underlying the ordering decision.
-Finally, long-term demand forecasts may be shared with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer.
+By forecasting the full demand distribution as opposed to point estimators such as the expected demand or the median of the distribution, the forecast quality can be evaluated for quantiles of the distribution, including the often extensive tails of the distribution.
+Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e.
+the mechanism behind the stochastic behavior of the customer demand.
+However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
+Using an operational quantity such as the order quantity will in most cases be an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
+From a practical perspective, separating the demand forecast from the operational decisions (i.e.
+calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning and reduces the complexity as it avoids coupling the complex delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
+This also allows to share long-term demand predictions with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer.
 From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other.
@@
 Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm.
-The major benefits of this method are that it is not only extremely performant in practical applications of demand forecasting at large scale, but also fully explainable.
+
+This means we use two subsequent, independent Cyclic Boosting models in order to estimate individual PDFs, the first to estimate the mean and the second to estimate the variance.
+The features may or may not differ between the mean and variance estimation models.
+But in any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
+The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.
+While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.
+
+In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.
+
+\subsection{Cyclic Boosting Recap}
+\label{sec:CB}
+
+In our earlier work \cite{Wick2019}, we have presented a new machine learning algorithm, Cyclic Boosting, which can be categorized as a generalized additive model with a cyclic coordinate descent optimization featuring a boosting-like update of parameters.
+The major benefits of Cyclic Boosting are that it is not only extremely performant in practical applications, like demand forecasting, at large scale, but also fully explainable.
 Traditional machine learning algorithms are typically "black box'' approaches, where the individual decision cannot be explained.
 Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
-The main idea of this algorithm is the following: When used for regression, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:
+The main idea of this algorithm is the following: When used for regression, denoted as multiplicative regression mode, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:
 
@@
 
-The parameters $f^k_j$ are the model parameters that are determined from features $j$.
+The values $f^k_j$ are the model parameters that are determined from features $j$.
 Each feature is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
@@
 The factors $f^k_j$ are determined iteratively until convergence is reached and regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm.
-The relative strength of the factors  $f^k_j$ are then directly interpretable in relation to how important a specific feature is for each individual prediction, where deviations from  $f^k_j=1$ indicate high importance.
+The deviations of the different factors from $f^k_j=1$ are then directly interpretable in terms of how much a specific feature contributes to each individual prediction.
 
-\textcolor{red}{Need for some details on Cyclic Boosting training process}
-
-For the PDF estimation, two subsequent, independent Cyclic Boosting models are used, the first to estimate the mean and the second to estimate the variance.
-The features may or may not differ between the mean and variance estimation models.
-But in any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
-The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.
-
-While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.
-
-In the upcoming section, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.
+The following meta-algorithm is used to determine the model parameters $f^k_j$ from the training data:
+\begin{enumerate}
+\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and
+features $j$.}
+\item{Initialize the factors $f^k_j \leftarrow 1$}
+\item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
+\begin{equation} \label{factors}
+g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}
+\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
+\end{equation}
+This means $g$ is the factor that is multiplied to $f_{t-1}$ in each iteration. Here, $\hat{y}_\tau$ is calculated according to eqn. \ref{eqn:cb} with the current values of the aggregated factors $f$:
+\begin{equation} \label{factors3}
+\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
+\end{equation}
+To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ employs $f^k_{j,t-1}$ in the calculation of $\hat{y}$.
+For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).
+}
+\item{Quit when stopping criteria, e.g.
+the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
+\end{enumerate}
 
@@
 
-Cyclic Boosting in a modified form of its multiplicative regression mode, namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
+Cyclic Boosting in a modified form of its multiplicative regression mode (briefly described in sec.
+\ref{sec:CB}), namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
 
@@
 
-The variance estimation is achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to $r_i$ over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
+The variance estimation is achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $r_i$ according to  \eqn \eqref{eqn:r}, over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
 
@@
 \end{equation}
-
-\textcolor{red}{Need for more details.}
 
@@
 
-The 1st Wasserstein distance, also known as earth mover'€™s distance (EMD), represents a distance function defined between two probability distributions on a given metric space, and for our purposes here can defined by
+The 1st Wasserstein distance, also known as earth mover distance (EMD), represents a distance function defined between two probability distributions on a given metric space, and for our purposes here can defined by
 
@@
 The Kullback-Leibler divergence is a measure of how one probability distribution diverges from a second expected probability distribution.
-For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-“Leibler divergence from Q to P is defined as
+For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-Leibler divergence from Q to P is defined as
 
@@
 
-\textcolor{red}{... example showing full prediction and evaluation chain ...}
+\subsection{Mean Estimation}
 
-\textcolor{red}{... plot showing individual PDF example ...}
+...
 
-%\begin{figure}
-%\begin{center}
+\begin{figure}
+\begin{center}
+%\includegraphics[width=8cm]{../figures/cb_mean}
+\caption{\label{fig:cb_mean} 2 CB analysis plots for mean model}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+%\includegraphics[width=8cm]{../figures/mean_prediction}
+\caption{\label{fig:mean_prediction} time series of mean prediction and actuals for 2 product groups}
+\end{center}
+\end{figure}
+
+\subsection{Variance Estimation and PDF Predictions}
+
+...
+
+\begin{figure}
+\begin{center}
+%\includegraphics[width=8cm]{../figures/cb_var}
+\caption{\label{fig:cb_var} 2 CB analysis plots for variance model}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 %\includegraphics[width=8cm]{../figures/pdf_example}
-%\caption{\label{fig:pdf_example} ...}
-%\end{center}
-%\end{figure}
+\caption{\label{fig:pdf_example} PDF example}
+\end{center}
+\end{figure}
 
-\textcolor{red}{... plot showing cdf histo ...}
+\subsection{Evaluation of PDF Predictions}
 
-%\begin{figure}
-%\begin{center}
+...
+
+\begin{figure}
+\begin{center}
 %\includegraphics[width=8cm]{../figures/cdf_demand}
-%\caption{\label{fig:cdf_demand} ...}
-%\end{center}
-%\end{figure}
+\caption{\label{fig:cdf_demand} CDF histogram}
+\end{center}
+\end{figure}
 
-\textcolor{red}{... plot showing invquant profile for mean prediction on X-axis ...}
+\begin{figure}
+\begin{center}
+%\includegraphics[width=8cm]{../figures/invquant_mean}
+\caption{\label{fig:invquant_mean} invquant profile for mean prediction on X-axis}
+\end{center}
+\end{figure}
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=8cm]{../figures/invquant_mean}
-%\caption{\label{fig:invquant_mean} ...}
-%\end{center}
-%\end{figure}
+\begin{figure}
+\begin{center}
+%\includegraphics[width=8cm]{../figures/invquant_pg}
+\caption{\label{fig:invquant_pg} invquant profile for product groups on X-axis}
+\end{center}
+\end{figure}
 
-\textcolor{red}{... plot showing invquant profile for product groups on X-axis ...}
+... quantitative CDF accuracy ...
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=8cm]{../figures/invquant_pg}
-%\caption{\label{fig:invquant_pg} ...}
-%\end{center}
-%\end{figure}
+\section{Conclusion}
+
+We have shown how to use two subsequent machine learning models for mean and variance estimation together with a negative binomial model assumption to come up with individual PDF predictions.
+This setup is especially useful for retail demand forecasting, where the true demand follows a negative binomial distribution quite closely.
+Compared to a model-free approach like quantile regression, the distributional assumption therefore drastically reduces the uncertainty of the resulting predictions, which are, by using Cyclic Boosting as underlying machine learning algorithm, fully explainable on the individual level.
+
+Furthermore, we have presented new qualitative and quantitative methods for evaluating predictions in form of full PDFs.
+For qualitative evaluation, it is important to check the full PDF, especially its tails that are often crucial for subsequent decision making, as well as investigate aggregations of individual PDFs (to reduce the uncertainty of the evaluation) dependent on specific variables, and we have proposed a novel profile approach called inverse quantile plot for this.
+For quantitative evaluation, it is always desirable to have a single number to compare different models in terms of accuracy, and we have suggested to use the deviance between the CDF histogram of the predicted PDFs and the uniform distribution to compare PDF accuracies.
 

### 2020-07-21 LaTex/paper.tex
added example code and figures (need for refinement and other data set)

@@
 specific categories of a feature).
-Two examples for this can be found in figures \ref{fig:invquant_mean} and \ref{fig:invquant_pg} in the next section.
+Two examples for this can be found in figures \ref{fig:invquant_mean} and \ref{fig:invquant_dayofweek} in the next section.
 
@@
 
-\subsection{Mean Estimation}
+\subsection{Mean and Variance Estimations}
 
-...
+\textcolor{red}{...}
 
@@
 \begin{center}
-%\includegraphics[width=8cm]{../figures/cb_mean}
-\caption{\label{fig:cb_mean} 2 CB analysis plots for mean model}
+\includegraphics[width=8cm]{../figures/mean_prediction}
+\caption{\label{fig:mean_prediction} time series of mean prediction and actuals for one item in all stores}
 \end{center}
@@
 \begin{center}
-%\includegraphics[width=8cm]{../figures/mean_prediction}
-\caption{\label{fig:mean_prediction} time series of mean prediction and actuals for 2 product groups}
-\end{center}
-\end{figure}
-
-\subsection{Variance Estimation and PDF Predictions}
-
-...
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=8cm]{../figures/cb_var}
-\caption{\label{fig:cb_var} 2 CB analysis plots for variance model}
-\end{center}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=8cm]{../figures/pdf_example}
-\caption{\label{fig:pdf_example} PDF example}
+\includegraphics[width=8cm]{../figures/cdf}
+\caption{\label{fig:cdf_example} CDF example}
 \end{center}
@@
 
-...
+\textcolor{red}{...}
 
@@
 \begin{center}
-%\includegraphics[width=8cm]{../figures/cdf_demand}
+\includegraphics[width=8cm]{../figures/cdf_truth}
 \caption{\label{fig:cdf_demand} CDF histogram}
@@
 \begin{center}
-%\includegraphics[width=8cm]{../figures/invquant_mean}
+\includegraphics[width=8cm]{../figures/invquant_yhat_mean}
 \caption{\label{fig:invquant_mean} invquant profile for mean prediction on X-axis}
@@
 \begin{center}
-%\includegraphics[width=8cm]{../figures/invquant_pg}
-\caption{\label{fig:invquant_pg} invquant profile for product groups on X-axis}
+\includegraphics[width=8cm]{../figures/invquant_dayofweek}
+\caption{\label{fig:invquant_dayofweek} invquant profile for day of week on X-axis}
 \end{center}
@@
 
-... quantitative CDF accuracy ...
+\textcolor{red}{... quantitative CDF accuracy ...}
 

### 2020-07-21 code/california_housing.py

@@ -1,375 +1,0 @@
-"""
-Cond-dist predictions for the California Housing dataset.
-
-Source and problem description:
-
-* http://www.dcc.fc.up.pt/~ltorgo/Regression/cal_housing.html
-* Hastie, The Elements of Statistical Learning, 10.14.1
-
-
-Why are feature bins filled with **unnecessarily heterogeneous data** harmful
-in prediction projects?
-
-* Many machine-learning algorithms such as Neurobayes without internal boost
-  can only consider axis-parallel cuts through the feature space.
-
-  Often target averages inside these subspaces are used for the prediction,
-  in particular averages over such heterogeneous feature bins.
-
-* Such averages may not be meaningful or even **misleading**. For example, if
-  you lumped together apples and peas, then the average size is neither a good
-  predictor for apples nor for peas.
-
-  Taking the median is not much better.
-
-* Heterogeneous feature bins often lead to the famous **XOR problem** in
-  machine learning which is notoriously hard to solve for many
-  machine learning algorithms.
-
-* In **power-law distributions**, taking averages is very misleading, e.g. if
-  you lump together Harry Potter books with specialist books about twelve tone
-  music, then taking the average sales leads to very bad predictions.
-"""
-from __future__ import print_function, division
-
-import numpy as np
-import pandas as pd
-import cPickle
-
-import matplotlib.pylab as plt
-import matplotlib.ticker as ticker
-
-import sklearn.datasets
-import sklearn.cross_validation
-import sklearn.neighbors
-import sklearn.ensemble
-
-from nbpy import cyclic_boosting, pipeline, x_transformer, binning
-from nbpy.cyclic_boosting.plots import plot_analysis
-from nbpy import flags, meta_estimator, on_cols
-from nbpy.smoothing.onedim import IsotonicRegressor
-
-from nbpy.ext import predict_as_transform, plots
-
-
-USE_PICKLE = False
-GRADBOOST = False
-
-
-def get_data(test_size):
-    """Get the data for training and test.
-
-    For the parameters, see :func:`sklearn.cross_validation.train_test_split`.
-
-    :return: ``X_train, X_test, y_train, y_test``, the X's being Pandas
[…]

### 2020-07-21 code/kaggle_demand.py

@@ -9,10 +9,14 @@ mpl.use('Agg')
 import matplotlib.pyplot as plt
 import matplotlib.ticker as ticker
 
+from scipy.stats import nbinom
+from scipy.stats import binned_statistic
+
 from nbpy import binning, pipeline, cyclic_boosting
 from nbpy import flags
 from nbpy.smoothing.onedim import SeasonalSmoother, IsotonicRegressor
 from nbpy.cyclic_boosting.plots import plot_analysis
+from nbpy.cyclic_boosting.nbinom import CBNBinomC
 
 from IPython import embed
 
@@ -26,7 +30,7 @@ def plot_CB(filename, plobs, binner):
         )
 
 
-def plot_factors(est, X):
+def plot_factors(est, X, samples):
     factors = {}
     for feature in est.ests[-1][1].features:
         ft = "" if feature.feature_type is None else "_{}".format(
@@ -37,7 +41,7 @@ def plot_factors(est, X):
             ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False)) - 1.
     df_factors = pd.DataFrame(factors)
 
-    ax = df_factors.loc[[0,10000,50000]].plot(kind='barh', title="factors", legend=True, fontsize=15)
+    ax = df_factors.loc[samples].plot(kind='barh', title="factors", legend=True, fontsize=15)
     fig = ax.get_figure()
     ax.axvline(0, color="gray")
     ax.set_yticks([])
@@ -49,11 +53,45 @@ def plot_factors(est, X):
     fig.savefig('plots/factors.pdf')
 
 
+def plot_cdf(n, p):
+    plt.figure()
+    x = np.linspace(0.0, 100.0, num=100)
+    plt.plot(x, nbinom.cdf(x, n, p))
+    plt.tight_layout()
+    plt.savefig('plots/cdf.pdf')
+
+
+def plot_cdf_truth(cdf_truth):
+    plt.figure()
+    plt.hist(cdf_truth, bins=100)
+    plt.tight_layout()
+    plt.savefig('plots/cdf_truth.pdf')
+
+
+def plot_invquants(X, variable):
+    means_result = binned_statistic(X[variable], [X['cdf_truth']<=0.1, X['cdf_truth']<=0.3, X['cdf_truth']<=0.5, X['cdf_truth']<=0.7, X['cdf_truth']<=0.9, X['cdf_truth']<=0.97], bins=100, statistic='mean')
[…]

### 2020-07-21 LaTex/paper.tex
minor rewording

@@
 While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying random distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
-We present an extension to our earlier work on a supervised machine learning method called Cyclic Boosting to predict full individual probability density functions in form of negative binomial distributions, that are fully explainable on the individual prediction level.
+Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates.
+Additionally, each single prediction is explainable which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
 Also, we describe new techniques for both qualitative and quantitative evaluation of predicted probability density functions by comparison with observed data.
@@
 
-\subsection{Cyclic Boosting Recap}
+\subsection{Cyclic Boosting Algorithm}
 \label{sec:CB}
 
-In our earlier work \cite{Wick2019}, we have presented a new machine learning algorithm, Cyclic Boosting, which can be categorized as a generalized additive model with a cyclic coordinate descent optimization featuring a boosting-like update of parameters.
+Cyclic Boosting \cite{Wick2019} is a novel machine learning algorithm that can be categorized as a generalized additive model with a cyclic coordinate descent optimization featuring a boosting-like update of parameters.
 The major benefits of Cyclic Boosting are that it is not only extremely performant in practical applications, like demand forecasting, at large scale, but also fully explainable.

### 2020-07-23 LaTex/paper.tex
new data set (including model code and plots) and more explanations on math

@@
 This means we use two subsequent, independent Cyclic Boosting models in order to estimate individual PDFs, the first to estimate the mean and the second to estimate the variance.
-The features may or may not differ between the mean and variance estimation models.
-But in any case, as there is a strong correlation between mean and variance, it is beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
+The features may or may not differ between the mean and variance estimation models, and one might even include the corresponding predicted mean from the mean model as feature in the variance model.
 The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.
@@
 
-Cyclic Boosting in a modified form of its multiplicative regression mode (briefly described in sec.
-\ref{sec:CB}), namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
+Cyclic Boosting in a modified form of the one briefly described in sec.
+\ref{sec:CB}, namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
 
-For regression, the parametrization of the negative binomial mass function can be specified as \cite{hilbe2011negative}:
+For regression of a target variable $y$ taking the values $y = 0, 1, 2, ...$, the parametrization of the negative binomial mass function by means of mean $\mu$ and dispersion parameter $r$ can be specified as \cite{hilbe2011negative}:
 
@@
 
-with mean $\mu$ and dispersion parameter $r$. The target variable $y$ takes the values $y = 0, 1, 2, ...$.
-
-The variance estimation is achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $r_i$ according to  \eqn \eqref{eqn:r}, over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
-
-\begin{equation} \label{eqn:loss_likelihood}
-L(r) = -\mathcal{L}(r) = -\ln \sum_i \mathrm{NB}(y_i; \hat{\mu_i}, r_i)
-\end{equation}
-
-As stated in \eqn \eqref{eqn:r}, the values $r_i$ are defined by the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$.
-For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls.
-The dispersion parameter $r$ is bound to the interval $[1, \infty]$.
-
-\begin{equation} \label{eqn:r}
-r_i = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
-\end{equation}
-
-As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence.
-Unlike in the original multiplicative regression mode of Cyclic Boosting, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search.
-All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
-
-Finally, the variance $\hat{\sigma}^2_i$ can be calculated from the dispersion parameter $\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}.
-And so, together with the estimated mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of the negative binomial distribution NB for the prediction of an individual observation $i$ are given.
+Bounding the inverse of the dispersion parameter $1/r$ to the interval $[0, 1]$ (corresponding to bounding $r$ to the interval $[1, \infty]$), the variance $\sigma^2$ can be calculated from $\mu$ and $r$ via:
 
@@
 \end{equation}
+
+An estimation of the dispersion parameter $\hat{r}$ can be achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to the Cyclic Boosting parameters $f^k_j$ constituting the model of $\hat{r_i}$ according to \eqn \eqref{eqn:r}, over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
+
+\begin{equation} \label{eqn:loss_likelihood}
+L(r) = -\mathcal{L}(r) = -\ln \sum_i \mathrm{NB}(y_i; \hat{\mu_i}, \hat{r_i})
+\end{equation}
+
+\begin{equation} \label{eqn:r}
+\hat{r_i} = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
+\end{equation}
+
+So, the values $\hat{r_i}$ are estimated by means of the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$.
+For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls.
+Just like in sec.
+\ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and neutral factor $1$.
+Analog to the classification mode of Cyclic Boosting described in \cite{Wick2019}, the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function.
+
+As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence.
+Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec.
+\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search.
+All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
+
+Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}.
+And so, together with the estimated mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of the negative binomial distribution NB for the prediction of an individual observation $i$ are given.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/cdf}
-\caption{\label{fig:cdf_example} CDF example}
+\includegraphics[width=4cm]{../figures/pdf}
+\includegraphics[width=4cm]{../figures/cdf}
+\caption{\label{fig:pdf_example} PDF/CDF example}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/invquant_yhat_mean}
-\caption{\label{fig:invquant_mean} invquant profile for mean prediction on X-axis}
+\caption{\label{fig:invquant_mean} invquant profile for mean predictions of all product-location-day combinations on X-axis}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/invquant_dayofweek}
-\caption{\label{fig:invquant_dayofweek} invquant profile for day of week on X-axis}
+\caption{\label{fig:invquant_dayofweek} invquant profile for day of week of one product over all locations and days on X-axis}
 \end{center}

### 2020-07-23 code/kaggle_demand.py

@@ -1,379 +1,0 @@
-import sys
-
-import pandas as pd
-import numpy as np
-import datetime
-
-import matplotlib as mpl
-mpl.use('Agg')
-import matplotlib.pyplot as plt
-import matplotlib.ticker as ticker
-
-from scipy.stats import nbinom
-from scipy.stats import binned_statistic
-
-from nbpy import binning, pipeline, cyclic_boosting
-from nbpy import flags
-from nbpy.smoothing.onedim import SeasonalSmoother, IsotonicRegressor
-from nbpy.cyclic_boosting.plots import plot_analysis
-from nbpy.cyclic_boosting.nbinom import CBNBinomC
-
-from IPython import embed
-
-
-def plot_CB(filename, plobs, binner):
-    for i, p in enumerate(plobs):
-        plot_analysis(
-            plot_observer=p,
-            file_obj="plots/" + filename + "_{}".format(i), use_tightlayout=False,
-            binners=[binner]
-        )
-
-
-def plot_factors(est, X, samples):
-    factors = {}
-    for feature in est.ests[-1][1].features:
-        ft = "" if feature.feature_type is None else "_{}".format(
-            feature.feature_type)
-
-        factors[
-            "_".join(feature.feature_group) + ft
-            ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False)) - 1.
-    df_factors = pd.DataFrame(factors)
-
-    ax = df_factors.loc[samples].plot(kind='barh', title="factors", legend=True, fontsize=15)
-    fig = ax.get_figure()
-    ax.axvline(0, color="gray")
-    ax.set_yticks([])
-
-    def setfactorformat(x, pos):
-        return '%1.2f' % (x + 1)
-    ax.xaxis.set_major_formatter(ticker.FuncFormatter(setfactorformat))
-
-    fig.savefig('plots/factors.pdf')
-
-
-def plot_cdf(n, p):
-    plt.figure()
-    x = np.linspace(0.0, 100.0, num=100)
-    plt.plot(x, nbinom.cdf(x, n, p))
-    plt.tight_layout()
-    plt.savefig('plots/cdf.pdf')
-
-
-def plot_cdf_truth(cdf_truth):
-    plt.figure()
-    plt.hist(cdf_truth, bins=100)
-    plt.tight_layout()
-    plt.savefig('plots/cdf_truth.pdf')
[…]

### 2020-07-23 code/walmart_demand.py

@@ -1,0 +1,545 @@
+import sys
+
+import pandas as pd
+import numpy as np
+import datetime
+from dateutil.easter import easter
+
+import matplotlib as mpl
+mpl.use('Agg')
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+
+from scipy.stats import nbinom, poisson
+from scipy.stats import binned_statistic
+
+from nbpy import binning, pipeline, cyclic_boosting
+from nbpy import flags
+from nbpy.smoothing.onedim import SeasonalSmoother, IsotonicRegressor
+from nbpy.cyclic_boosting.plots import plot_analysis
+from nbpy.cyclic_boosting.nbinom import CBNBinomC
+from nbpy.cyclic_boosting import CBFixedVarianceRegressor
+from nbpy.x_transformer import MultiLabelEncoder
+from nbpy.cyclic_boosting.price import CBExponential
+
+from IPython import embed
+
+
+def plot_CB(filename, plobs, binner):
+    for i, p in enumerate(plobs):
+        plot_analysis(
+            plot_observer=p,
+            file_obj="plots/" + filename + "_{}".format(i), use_tightlayout=False,
+            binners=[binner]
+        )
+
+
+def plot_factors(est, X, samples):
+    factors = {}
+    for feature in est.ests[-1][1].features:
+        ft = "" if feature.feature_type is None else "_{}".format(
+            feature.feature_type)
+
+        factors[
+            "_".join(feature.feature_group) + ft
+            ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False)) - 1.
+    df_factors = pd.DataFrame(factors)
+
+    ax = df_factors.loc[samples].plot(kind='barh', title="factors", legend=True, fontsize=15)
+    fig = ax.get_figure()
+    ax.axvline(0, color="gray")
+    ax.set_yticks([])
+
+    def setfactorformat(x, pos):
+        return '%1.2f' % (x + 1)
+    ax.xaxis.set_major_formatter(ticker.FuncFormatter(setfactorformat))
+
+    fig.savefig('plots/factors.pdf')
+
+
+def plot_pdf(n, p):
+    plt.figure()
+    x = range(30)
+    plt.plot(x, nbinom.pmf(x, n, p))
[…]

### 2020-08-05 LaTex/paper.tex
clearer evaluation section, first paragraphs of example section

@@
 \begin{abstract}
-Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes.
-While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying random distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
+Demand forecasting is a central component for many aspects of supply chain operations as it provides crucial input for subsequent decision making like ordering processes.
+While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
 Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates.
-Additionally, each single prediction is explainable which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
-Also, we describe new techniques for both qualitative and quantitative evaluation of predicted probability density functions by comparison with observed data.
+While metrics evaluating simple point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods.
+Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
 \end{abstract}
 
-{Keywords: \textbf{machine learning, demand forecasting}}
+{Keywords: \textbf{explainable machine learning, demand forecasting, probability distribution}}
 
-PDF
 \section{Introduction}
@@
 Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.
-The ordering decision process is complicated by a range of factors: Even in the case of accurate  demand forecasts, the decision maker has to balance conflicting metrics to reach an optimal decision: Ordering to few items may result in stockout situations leading to unrealized demand and unsatisfied customers.
+The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision: Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers.
 Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
 This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
-These items typically have a shelf-life of less than a business day to a few business days at most, with a continuous spectrum depending on the exact item.
+These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item.
 In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured.
@@
 It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.).
-While the probability distribution describing the demand can be attributed to a given family or parametrization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors.
+While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors.
 The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy.
@@
 in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
-It aims to avoid estimating the underlying probability distribution and use the available data to derive the operational decisions directly.
+It aims to avoid estimating the underlying probability distribution for demand and use the available data to derive the operational decisions (the order quantity) directly.
 An overview of a range approaches can also be found in \cite{huber2019data}.
@@
 Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks.
-First of all, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand.
+First, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand.
 This can be used to simulate the performance on a per-item level and e.g.
@@
 However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
-Using an operational quantity such as the order quantity will in most cases be an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
+Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
 From a practical perspective, separating the demand forecast from the operational decisions (i.e.
@@
 
-In the following, we will demonstrate how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer.
+This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer.
 Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
@@
 
-Additionally, we show how the cumulative distribution function (CDF) or inverse quantiles can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution.
-This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer.
+Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution.
+This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer or gain further insights into customer behavior using for example causal modelling.
 
@@
 
-To predict an individual PDF by means of a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted.
-In the following, we describe such a parametric approach with the model assumption of a negative binomial probability distribution.
-For a PDF prediction under a negative binomial model assumption, there is the need for the estimation of two parameters, for example its mean and its variance.
+To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted.
+As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand.
+Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance.
+At least in principle, any method can be used.
+However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth.
+It should be noted that these parameters are not independent between products.
+For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment.
+This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically.
+Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches.
+Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm.
+This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance.
+The features may or may not differ between the mean and variance estimation models.
+It can also be beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
+The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.
 
-This can be done using two independent models, one to estimate the mean and the other for the variance.
-At least in principle, any method can be used.
-However, as argued above, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as product ID, date, and store location) and may depend on  a multitude of describing variables (features).
-Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm.
+In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD models using Cyclic Boosting.
 
-This means we use two subsequent, independent Cyclic Boosting models in order to estimate individual PDFs, the first to estimate the mean and the second to estimate the variance.
-The features may or may not differ between the mean and variance estimation models, and one might even include the corresponding predicted mean from the mean model as feature in the variance model.
-The assigned mean and variance predictions can then be used to generate individual PDFs according to a functional negative binomial distribution assumption for each sample.
-While any machine learning algorithm outputting a conditional mean as prediction may be used to get an approximate estimate for the mean of an underlying negative binomial distribution, it is preferable to choose an algorithm using a negative binomial assumption for its inner regularization methods, like Cyclic Boosting in multiplicative regression mode.
-
-In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict the variance of a negative binomial PDF along with its mean individually for each sample by means of Cyclic Boosting.
-
-\subsection{Cyclic Boosting Algorithm}
+\subsection{Cyclic Boosting Algorithm: Mean estimation}
 \label{sec:CB}
 
-Cyclic Boosting \cite{Wick2019} is a novel machine learning algorithm that can be categorized as a generalized additive model with a cyclic coordinate descent optimization featuring a boosting-like update of parameters.
-The major benefits of Cyclic Boosting are that it is not only extremely performant in practical applications, like demand forecasting, at large scale, but also fully explainable.
-Traditional machine learning algorithms are typically "black box'' approaches, where the individual decision cannot be explained.
-Cyclic Boosting on the other hand allows to follow how each individual prediction was made.
-The main idea of this algorithm is the following: When used for regression, denoted as multiplicative regression mode, the predictions $\hat{y_i}$ of the  target variable $Y \in [0,\infty)$ can be calculated in the following way:
+Cyclic Boosting \cite{Wick2019} is a type of generalized additive models using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters.
+Major benefits of Cyclic Boosting are its accuracy, performance even at large scale and providing fully explainable predictions which are of vital importance in practical applications, in particular for multi-national retail vendors.
 
+The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
+The global mean $\mu$ is determined from all target values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
+Single data records, for example the sales of a specific SKU along with all relevant features, are indexed by $i$.
+The individual predictions $\hat{y_i}$  can then be calculated as:
 \begin{equation} \label{eqn:cb}
@@
 \end{equation}
+The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges.
+During training,  regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm.
+The deviation of each factor from $f^k_j=1$ can then be used to explain how a  specific feature contributes to each individual prediction.
 
-The values $f^k_j$ are the model parameters that are determined from features $j$.
-Each feature is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
-The global mean $\mu$ is determined from all target values $y$ observed in the data.
-The factors $f^k_j$ are determined iteratively until convergence is reached and regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm.
-The deviations of the different factors from $f^k_j=1$ are then directly interpretable in terms of how much a specific feature contributes to each individual prediction.
-
-The following meta-algorithm is used to determine the model parameters $f^k_j$ from the training data:
+In detail, the following meta-algorithm describes how the model parameters $f^k_j$ are obtained from the training data:
 \begin{enumerate}
-\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and
-features $j$.}
+\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
 \item{Initialize the factors $f^k_j \leftarrow 1$}
@@
 \begin{equation} \label{factors}
-g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}
-\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
+g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
 \end{equation}
-This means $g$ is the factor that is multiplied to $f_{t-1}$ in each iteration. Here, $\hat{y}_\tau$ is calculated according to eqn. \ref{eqn:cb} with the current values of the aggregated factors $f$:
+Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration.
+The current prediction, $\hat{y}_\tau$, is calculated according to eqn.
+\ref{eqn:cb} with the current values of the aggregated factors $f$:
 \begin{equation} \label{factors3}
@@
 \end{equation}
-To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ employs $f^k_{j,t-1}$ in the calculation of $\hat{y}$.
-For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).
-}
+To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$.
+For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
 \item{Quit when stopping criteria, e.g.
@@
 
-\subsection{Variance Estimation by Cyclic Boosting in Negative Binomial Width Mode}
+\subsection{Cyclic Boosting Algorithm: Width estimation}
+\label{sec:cb_width}
 
-Cyclic Boosting in a modified form of the one briefly described in sec.
-\ref{sec:CB}, namely the negative binomial width mode presented in the following, can be used for the second machine learning model to predict the negative binomial variance associated with the mean predicted by the first model.
+In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model.
+In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm like described in the following.
 
-For regression of a target variable $y$ taking the values $y = 0, 1, 2, ...$, the parametrization of the negative binomial mass function by means of mean $\mu$ and dispersion parameter $r$ can be specified as \cite{hilbe2011negative}:
-
+When looking at the demand of individual SKUs, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
 \begin{equation} \label{eqn:nbinom}
@@
 \end{equation}
+where $\mu$ is the mean of the distribution and $r$ a dispersion parameter.
 
-Bounding the inverse of the dispersion parameter $1/r$ to the interval $[0, 1]$ (corresponding to bounding $r$ to the interval $[1, \infty]$), the variance $\sigma^2$ can be calculated from $\mu$ and $r$ via:
-
+By bounding the inverse of the dispersion parameter $1/r$ to the interval $[0, 1]$ (corresponding to bounding $r$ to the interval $[1, \infty]$), the variance $\sigma^2$ can be calculated from $\mu$ and $r$ via:
 \begin{equation} \label{eqn:variance_r}
@@
 
-An estimation of the dispersion parameter $\hat{r}$ can be achieved by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, expressed as negative log-likelihood function of a negative binomial distribution, with respect to the Cyclic Boosting parameters $f^k_j$ constituting the model of $\hat{r_i}$ according to \eqn \eqref{eqn:r}, over all input samples $i$, where the mean values $\hat{\mu_i}$ are fixed to the corresponding predictions from the first machine learning model outputting the conditional mean.
+The estimate of the dispersion parameter $\hat{r}$ can then be calculated by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, which is expressed as negative log-likelihood function of a negative binomial distribution.
+Using Cyclic Boosting, the minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained in the previous step (described in sec.
+\ref{sec:CB}).
 
@@
 
-So, the values $\hat{r_i}$ are estimated by means of the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$.
-For any concrete observation $i$, the index $k$ of the bin is determined by the observation of $x_{j,i}$ and the subsequent look-up into which bin this observation falls.
-Just like in sec.
-\ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and neutral factor $1$.
-Analog to the classification mode of Cyclic Boosting described in \cite{Wick2019}, the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function.
+In other words, the values $\hat{r_i}$ are estimated via learning the Cyclic Boosting model parameters $f^k_j$ for each feature $j$ and bin $k$ from data.
+For any concrete observation $i$, the index $k$ of the bin is determined by the value of the feature $x_{j,i}$ and the subsequent look-up into which bin this observation falls.
+Like in sec.
+\ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and again values deviating from $f^k_j=1$ can be used to explain the relative importance of a specific feature contributing to individual predictions.
+Note that the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function in the same way as explained in  \cite{Wick2019} when Cyclic Boosting is used for classification tasks.
 
-As described in \cite{Wick2019}, the parameter estimation in Cyclic Boosting is an iterative method corresponding to a cyclic coordinate descent, processing one feature with all its bins at a time until convergence.
+The Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached.
 Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec.
-\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and requires a numerical method, for example a random search.
+\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example using a random search.
 All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
 
-Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ by means of \eqn \eqref{eqn:variance_r}.
-And so, together with the estimated mean parameter $\hat{\mu_i}$ from the previous mean estimation model, all parameters of the negative binomial distribution NB for the prediction of an individual observation $i$ are given.
+Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ using \eqn \eqref{eqn:variance_r}.
+Using  the individual predicted mean $\hat{\mu_i}$ from the first step, the model is fully specified for each individual prediction $i$.
 
@@
 
-In the following, we will show how histograms of the CDF and inverse quantiles can be used to determine the accuracy of the predicted PDFs.
+In the following, we will show how different visualizations of the observed cumulative distribution function (CDF) values can be used to evaluate the quality of the predicted PDFs.
 Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.
@@
 
-As a first step, we use the inverse quantile function or probability integral transform (PIT), see e.g.
-\cite{Angus1994,casella2002statistical}.
-This offers a first qualitative assessment by looking at the distribution of the quantiles of the predicted density via the CDF.
-In case of a correctly calibrated predicted PDF, this distribution is expected to be uniform and any deviation can be interpreted as a hint that the predicted probability function is not correct \cite{diebold1998vevaluating}.
+In the simplest case, we only have one model with one set of model parameters to cover all predictions.
+In this case, the evaluation of the full probability distribution is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
+Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
 
-\subsubsection{Histogram of the CDF}
+In practical applications however, we have a large number of effective prediction models, since although we always use the same model parameterization, such as the negative binomial distribution, its parameters have to be determined at the required level of granularity.
+For example, for daily orders, we need to predict the parameters of the negative binomial distribution for each location, sales day, and product.
+Unlike the simple case discussed above, where we had many observations to compare the prediction model to, we now have just a single observation per prediction, meaning that we cannot use statistical tests directly.
+
+\subsubsection{Histogram of CDF Observations}
 \label{sec:cdf_histo}
+
+For a first qualitative assessment, we make use of the probability integral transform, see e.g.
+\cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed.
+So, in case of a correctly calibrated PDF prediction, the distribution of the corresponding actually observed CDF values of the individual PDF predictions is expected to be uniform, {\em regardless} of the shape of the predicted distribution, and any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.
 
@@
 
-If the CDF is continuous and strictly increasing, then the inverse of the CDF $F^{-1}(y)$ exists and is a unique real-valued number $x$ for each $y \in [0,1]$ so that we can write $F(x) = y$.
-The latter defines the inverse quantile function because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
+If the CDF is continuous and strictly increasing, then the inverse of the CDF, $F^{-1}(y)$, exists and is a unique real-valued number $x$ for each $y \in [0,1]$, so that we can write $F(x) = y$.
+The latter defines the inverse quantile function, because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
 
@@
 \includegraphics[width=8cm]{../figures/PdfCdf}
-\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution funtion of a normal distrubution.}
+\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution function of a normal distribution.}
 \end{center}
@@
 Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF.
-This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s\in [0,1]$ in which $s$ is defined and $s$ can be interpreted as the cumulative distribution of its own probability distribution:
+This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s \in [0,1]$ in which $s$ is defined, and $s$ can be interpreted as the cumulative distribution of its own probability distribution:
 
@@
 
-In case of discrete probability functions such as the negative binomial function, the same argument still holds but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
+In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
 \cite[p.
@@
 
-A histogram of the quantiles (or PIT values) of each prediction is therefore expected to be uniformly distributed in $[0,1]$ if the predicted probability distribution $f(x)$ is correctly calibrated.
-
-This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of quantiles (or PIT values) for five different cases.
-If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution as shown in the last part of the figure.
-The first two plots of \fig \ref{fig:cdf_histos} indicate a bias in the variance estimation and the third and fourth plot indicate a bias in the mean
-estimation.
+A histogram of the observed quantiles (or CDF values) for each individual PDF prediction is therefore expected to be uniformly distributed in $[0,1]$, if the predicted probability distribution $f(x)$ is correctly calibrated.
+This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed quantiles for five different cases.
+If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution.
+If the mean or the variance are not estimated correctly, the resulting distribution will show a distinct deviation from this uniform behavior.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/cdf_histos}
-\caption{\label{fig:cdf_histos} ...}
+\includegraphics[scale=0.4]{../figures/CDFHistos_uniform} \includegraphics[scale=0.4]{../figures/CDFHistos_broad} \includegraphics[scale=0.4]{../figures/CDFHistos_narrow} \includegraphics[scale=0.4]{../figures/CDFHistos_over} \includegraphics[scale=0.4]{../figures/CDFHistos_under}
+\caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
+correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
 \end{center}
@@
 
-\subsubsection{Inverse Quantile Plot}
-Evaluating the full probability function in the simplest case, where we have only one model that can be used to predict all observations, is straight-forward: We could fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
-Then, we could directly compare the model curve with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
-In practical applications however, we have a large number of effective prediction models, since although we always use the same model parametrization, such as the negative binomial distribution, its parameters have to be determined at the required level of granularity.
-For example, for daily orders, we need to predict the parameters of the negative binomial distribution for each location, sales day, and product.
-Unlike the simple case discussed initially, where we had many observations to compare the prediction model to, we now have just one observation per prediction, meaning that we cannot use statistical tests directly.
-Instead we use the fact that a correctly calibrated, predicted PDF should return a uniform cumulative density distribution {\em regardless} of the shape of the predicted distribution.
+\subsubsection{Quantile Profile Plot}
 
-We therefore proceed as follows: First, each individual predicted probability density distribution is transformed to its cumulative density distribution.
-In the example of the granularity from above, we would then have a CDF for each predicted PDF for each product, sales day, and sales location.
-The CDF values, also known as quantiles, of the corresponding truth values are then compared (in the sense of higher or lower) to expected quantile values and accordingly filled to a collection of profile plots, each of which representing one expected quantile value.
+We now refine the method from sec.
+\ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e.
+the sales records, with specified quantile values.
+For example, if we wanted to check for the quantile value $0.3$, we would look at the observed quantiles for each prediction and calculate the fraction of these being lower than $0.3$, expecting a fraction of $0.3$ for accurate PDF predictions.
+In order to judge quality of the full PDF predictions, we calculate these fractions for a range of different specified quantile values.
+When choosing which quantile to compare it is important to cover the tails of the distribution to make sure that even relatively rare events are estimated correctly by the machine learning algorithm.
+We call this method quantile profile plot.
 
-\fig \ref{fig:invquant_profiles} illustrates five different collections of inverse quantile profile plots (each collection comparing to 7 expected quantile values), each with only one bin for the sake of visualization, for separate sets of exemplary probability density estimation and observed data combinations.
-Each shown line represents the percentage of probability density estimation and observed data combinations for which the observed data point should be above and below the quantile of the predicted PDF indicated by that line (for example, the 0.50 line, corresponding to the median, indicates that 50 percent of all probability density estimation and observed data combinations in a given set of data should fall above the line, and 50 percent should fall below the line), respectively.
-The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of probability density estimations.
-For a set of accurate PDF predictions, one expects a uniform shape, like illustrated by the final far-right column of shaded circles.
-Here, all shaded circles exactly fall on the line, indicating that both mean and variance estimations, as well as the choice of a negative binomial distribution as underlying PDF, were correct.
-Conversely, the leftmost two columns of shaded circles illustrate a case for which the variance estimation has a bias, and the center and center-right columns of shaded circles illustrate cases for which the mean estimation has a bias.
+\fig \ref{fig:quant_profiles} illustrates five different collections of quantile profile plots (each collection comparing to 7 specified quantile values), for separate sets of exemplary PDF estimations and observed data combinations.
+The dashed horizontal lines indicates the fraction we expect, i.e.
+the specified quantile value, if the predictions are correct.
+For example, for the median, the line at 0.5 indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line.
+The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations.
+In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.
 
@@
 \includegraphics[width=8cm]{../figures/invquant_profiles}
-\caption{\label{fig:invquant_profiles} ...}
+\caption{\label{fig:quant_profiles} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events.
+In the leftmost two columns of shaded circles illustrate the behavior if the estimate of the variance is biased, the center and center-right columns of shaded circles illustrate cases for which the mean estimation is biased and the far right column shows the expected behaviour if all predictions are correct.}
 \end{center}
@@
 
-The two advantages of this method, as compared to the simpler method presented in section \ref{sec:cdf_histo}, are that an inverse quantile plot supports the qualitative evaluation of the predicted individual PDFs not only globally but (1) for different specified quantiles (potentially hinting to deviations in e.g.
-the tails of the distributions) and (2) in dependence of arbitrary variables of the data set (potentially hinting to deviations in e.g.
-specific categories of a feature).
+In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec.
+\ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population.
+However, calculating the fractions of the quantile profile plots globally, i.e.
+over all samples in the data set, might not reveal certain deficiencies for a subset, e.g.
+a specific store or product group in the example of retail demand forecasts.
+Therefore, we combine the approach described above with the method of profile plots, where the quantity on the x-axis of the quantile profile plot can be any variable of the data set at hand.
+Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
+
+In summary, this approach has two major benefits compared to the method discussed in sec.
+\ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
+Second, by showing the dependency from the (arbitrary) variable on the x-axis, the quantile profile plot reveals deviations of the predicted PDF from the acuals for different parts, e.g.
+specific categories, of that variable.
 Two examples for this can be found in figures \ref{fig:invquant_mean} and \ref{fig:invquant_dayofweek} in the next section.
@@
 
-The methods described so far allow a detailed qualitative evaluation of PDF predictions.
-However, in order to also quantify the quality of the PDF predictions, a measure of the deviation of the PDF predictions from the optimal outcome given the observed target data is needed.
-To achieve this, we compare the CDF histogram of the predicted PDFs with the uniform distribution and define a prediction accuracy interval between 0 and 1.
-Several different methods can be used to compute the deviance between two probability distributions.
-Good choices are the \textbf{1st Wasserstein distance} \cite{olkin1982}, the \textbf{Kullback-Leibler divergence} \cite{kullback1951}, and the \textbf{Jensen-Shannon divergence}  \cite{dagan1997}.
+The methods discussed  so far allow a detailed qualitative evaluation of PDF predictions.
+However, in order to also quantify the quality of the PDF predictions, a metric is required that assesses the difference between the distributions.
+Therefore, we want to compare the CDF histogram of the predicted PDFs with the expected  uniform distribution and define a prediction accuracy interval between 0 and 1 where an accuracy of 1 indicates that both distributions agree perfectly.
+Several approaches that measure the difference between two probability distributions are suggested in the literature such as  the first Wasserstein distance \cite{olkin1982}, the Kullback-Leibler divergence \cite{kullback1951}, and the Jensen-Shannon divergence  \cite{dagan1997}.
 
-The 1st Wasserstein distance, also known as earth mover distance (EMD), represents a distance function defined between two probability distributions on a given metric space, and for our purposes here can defined by
-
+The first Wasserstein distance, also known as earth mover distance (EMD), represents a distance  defined between two probability distributions on a given metric space, and for our purposes here can defined by:
 \begin{equation}
@@
 \end{equation}
-
 where $F_P(X)$ and $F_Q(X)$ are the CDFs of the two PDFs $P(X)$ and $Q(X)$, respectively, and $x_k$ denotes the average value of $X$ in bin $k$, with $X$ being divided in $N$ bins.
+Compared to the other metrics given below, the first Wasserstein or earth mover distance has the additional benefit, that it takes a specific metric space into account, meaning that it depends on the distance of potential deviations.
 
 The Kullback-Leibler divergence is a measure of how one probability distribution diverges from a second expected probability distribution.
-For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-Leibler divergence from Q to P is defined as
-
+For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-Leibler divergence from Q to P is defined as:
 \begin{equation}
@@
 \end{equation}
-
-where $\log$ can be either base-2 or natural logarithm.
+where the logarithm can be either base-2 or the natural logarithm.
 
 The Jensen-Shannon divergence (JSD) can be seen as a symmetrized and smoothed version of the Kullback-Leibler divergence:
-
 \begin{equation}
@@
 \end{equation}
-
 where $M = 0.5  \cdot (P + Q)$.
-
-The EMD has the advantage over the mentioned divergences that it takes an underlying metric space into account, meaning that it depends on the distance of potential deviations.
 
@@
 
-\textcolor{red}{... predict an individual demand volume for different product-location-date combinations ...}
+We use data from a Kaggle online competition about estimating unit sales of Walmart retail goods \cite{kaggle_data} to demonstrate forecasting of demand for different product-location-date combinations in form of full individual PDFs by means of Cyclic Boosting and subsequent qualitative and quantitative evaluation of the PDF predictions.
+The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date, while the field sales represents the target to predict.
+In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}.
+Besides the key and target fields (\texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}), we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.
 
-\subsection{Mean and Variance Estimations}
+\subsection{Mean Estimation}
+\label{sec:example_mean}
 
-\textcolor{red}{...}
+For our Cyclic Boosting mean estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several variables describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), and a flag denoting a promotion.
+We also include various two-dimensional combinations of the mentioned features, where one of the two dimensions being \texttt{store\_id} or \texttt{item\_id} to learn characteristics of individual locations and products.
+Furthermore, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}.
+We plan to report on the inner workings of this method in a future work.
+
+\textcolor{red}{... training and prediction periods ... residual correction ...}
+
+\textcolor{red}{... results: mad, mean(y) ... example time series plot ...}
 
@@
 \includegraphics[width=8cm]{../figures/mean_prediction}
-\caption{\label{fig:mean_prediction} time series of mean prediction and actuals for one item in all stores}
+\caption{\label{fig:mean_prediction} \textcolor{red}{time series of mean prediction and actuals for one item in all stores}}
 \end{center}
 \end{figure}
+
+\subsection{Variance Estimation}
+
+For our Cyclic Boosting variance estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, a flag denoting a promotion, and the insample prediction of the mean model, and the individual correction factor of the residual correction step described in \ref{sec:example_mean}.
+
+\textcolor{red}{... same training and prediction periods ... no residual correction ...}
+
+\textcolor{red}{... pdf/cdf example for visualization...}
 
@@
 \includegraphics[width=4cm]{../figures/cdf}
-\caption{\label{fig:pdf_example} PDF/CDF example}
+\caption{\label{fig:pdf_example} \textcolor{red}{PDF/CDF example}}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/cdf_truth}
-\caption{\label{fig:cdf_demand} CDF histogram}
+\caption{\label{fig:cdf_demand} \textcolor{red}{CDF histogram}}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/invquant_yhat_mean}
-\caption{\label{fig:invquant_mean} invquant profile for mean predictions of all product-location-day combinations on X-axis}
+\caption{\label{fig:invquant_mean} \textcolor{red}{invquant profile for mean predictions of all product-location-day combinations on x-axis}}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/invquant_dayofweek}
-\caption{\label{fig:invquant_dayofweek} invquant profile for day of week of one product over all locations and days on X-axis}
+\caption{\label{fig:invquant_dayofweek} \textcolor{red}{invquant profile for day of week of one product over all locations and days on x-axis}}
 \end{center}
@@
 Furthermore, we have presented new qualitative and quantitative methods for evaluating predictions in form of full PDFs.
-For qualitative evaluation, it is important to check the full PDF, especially its tails that are often crucial for subsequent decision making, as well as investigate aggregations of individual PDFs (to reduce the uncertainty of the evaluation) dependent on specific variables, and we have proposed a novel profile approach called inverse quantile plot for this.
+For qualitative evaluation, it is important to check the full PDF, especially its tails that are often crucial for subsequent decision making, as well as investigate aggregations of individual PDFs (to reduce the uncertainty of the evaluation) dependent on specific variables, and we have proposed a novel profile approach called quantile profile plot for this.
 For quantitative evaluation, it is always desirable to have a single number to compare different models in terms of accuracy, and we have suggested to use the deviance between the CDF histogram of the predicted PDFs and the uniform distribution to compare PDF accuracies.
@@
 \section{Profile Histograms}
+\label{sec:profile}
 
@@
 Profile histograms are specifically designed to address this shortcoming.
-Intuitively, profile histograms are a one-dimensional representation of the two-dimensional scatter plot and are obtained  in the following way: The variable on the $x$ axis is discretizised into a suitable range of bins.
+Intuitively, profile histograms are a one-dimensional representation of the two-dimensional scatter plot and are obtained  in the following way: The variable on the $x$ axis is discretized into a suitable range of bins.
 The exact choice of binning depends on the problem at hand.

### 2020-08-06 LaTex/paper.tex
first draft for example section

@@
 The features may or may not differ between the mean and variance estimation models.
-It can also be beneficial to include the corresponding predicted mean from the mean model as feature in the variance model.
 The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.
@@
 \subsection{Quantitative Evaluation of PDF Predictions}
+\label{sec:cdf_acc}
 
@@
 The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date, while the field sales represents the target to predict.
-In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}.
+In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}, where we use all data before 2016 for the training and all data from 2016 for the testing of our model.
 Besides the key and target fields (\texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}), we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.
@@
 We also include various two-dimensional combinations of the mentioned features, where one of the two dimensions being \texttt{store\_id} or \texttt{item\_id} to learn characteristics of individual locations and products.
-Furthermore, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}.
+Furthermore, to capture the price sensitivity as feature, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}.
 We plan to report on the inner workings of this method in a future work.
 
-\textcolor{red}{... training and prediction periods ... residual correction ...}
+Unlike most state-of-the-art time series forecasting methods, we do not include lagged target information, for example via stacking of exponentially weighted moving average features, in our model.
+The reason for this is that including such information makes the learning of exogenous effects, e.g.
+promotions or events, much harder, as the model tends to rely mainly on temporal confounding.
+So, to instead strive for direct learning of causal effects and in turn improve on explainability, mid- and long-term forecasting quality, and predictability of rare events, it is beneficial to not include lagged target information in the machine learning model.
+In order to capture recent trends that are not reflected in the exogenous features of the model, we apply an individual residual correction on each of the predictions of the machine learning model, which accounts for deviations between the exponentially weighted moving average of the predictions and targets of each product-location combination over the corresponding past.
+Again, we plan to give more insights on the reasoning for such a residual correction in a future work.
 
-\textcolor{red}{... results: mad, mean(y) ... example time series plot ...}
+Just to give a rough verification for the accuracy of the mean predictions, we report the MAD and MSE averaged over the full used data set in the test period to be $1.68$ and $10.20$, respectively, while the average of the target, i.e.
+the sales, is $3.28$.
+\fig \ref{fig:mean_prediction} shows an example for the time series of both predictions and sales for one specific item, namely \texttt{FOODS\_3\_586} (which we will use throughout the rest of this example), summed over all 10 stores during the test period.
 
@@
 \includegraphics[width=8cm]{../figures/mean_prediction}
-\caption{\label{fig:mean_prediction} \textcolor{red}{time series of mean prediction and actuals for one item in all stores}}
+\caption{\label{fig:mean_prediction} Time series of mean prediction and sales in test period for one specific item summed over all stores.}
 \end{center}
@@
 
-For our Cyclic Boosting variance estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, a flag denoting a promotion, and the insample prediction of the mean model, and the individual correction factor of the residual correction step described in \ref{sec:example_mean}.
-
-\textcolor{red}{... same training and prediction periods ... no residual correction ...}
-
-\textcolor{red}{... pdf/cdf example for visualization...}
+For our Cyclic Boosting variance estimation model, the training and test periods are the same as for the mean estimation model, and the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}, meaning that the mean predictions are created in an insample way for the training period.
+We use a different set of features, namely the following: categorical variables for \texttt{store\_id} and \texttt{item\_id} and a flag denoting a promotion.
+An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_586} in store \texttt{WI\_1} on 2016-05-22 is shown in \fig \ref{fig:pdf_example}.
 
@@
 \includegraphics[width=4cm]{../figures/cdf}
-\caption{\label{fig:pdf_example} \textcolor{red}{PDF/CDF example}}
+\caption{\label{fig:pdf_example} Predicted PDF (left) and CDF (right) for a specific product-location-date combination.}
 \end{center}
@@
 
-\textcolor{red}{...}
+\fig \ref{fig:cdf_demand} shows the histogram of CDF observations according to the method described in sec.
+\ref{sec:cdf_histo} for item \texttt{FOODS\_3\_586} in the test period, where a significant deviation from the uniform distribution, expected for optimal PDF predictions, can be seen.
+The shape of the deviation points towards a slight underprediction of the mean estimation (see \fig \ref{fig:cdf_histos}), which is also visible in the time series plot of \fig \ref{fig:mean_prediction} (e.g.
+the high peak during two days end of April).
 
@@
 \includegraphics[width=8cm]{../figures/cdf_truth}
-\caption{\label{fig:cdf_demand} \textcolor{red}{CDF histogram}}
+\caption{\label{fig:cdf_demand} Histogram of all observed CDF values of the individual PDF predictions for a specific product in the test period, to be compared to a uniform distribution.}
 \end{center}
 \end{figure}
+
+\fig \ref{fig:invquant_mean} and \fig \ref{fig:invquant_dayofweek} show quantile profile plots for different variables on the x-axis, namely mean predictions and day of week, respectively.
+In the former case all product-location-day combinations in the test period are used to generate the statistics in the plot, while in the latter case all locations and days in the test period for item \texttt{FOODS\_3\_586} are aggregated.
+It can be seen that the different weekdays show different patterns and that there are slightly larger deviations for higher mean predictions (with fewer samples though).
 
@@
 \includegraphics[width=8cm]{../figures/invquant_yhat_mean}
-\caption{\label{fig:invquant_mean} \textcolor{red}{invquant profile for mean predictions of all product-location-day combinations on x-axis}}
+\caption{\label{fig:invquant_mean} Quantile profile plot for mean predictions on the x-axis aggregated over all product-location-day combinations in the test period.}
 \end{center}
@@
 \includegraphics[width=8cm]{../figures/invquant_dayofweek}
-\caption{\label{fig:invquant_dayofweek} \textcolor{red}{invquant profile for day of week of one product over all locations and days on x-axis}}
+\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the x-axis aggregated over all locations and days in the test period for a specific product.}
 \end{center}
@@
 
-\textcolor{red}{... quantitative CDF accuracy ...}
+The quantitative results for the CDF accuracies of our PDF predictions for the different metrics described in sec.
+\ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf:acc}.
+As benchmark, we compare these to the results of a Poisson model assumption needing only the mean as single parameter.
+Our negative binomial PDF predictions show a significant improvement over the simpler Poisson model, which uses the same mean predictions, for each of the metrics.
+
+\begin{table}[h!]
+\begin{center}
+\caption{CDF accuracies for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
+\label{tab:cdf:acc}
+\begin{tabular}{c|c|c}
+ & \textbf{NBD} & \textbf{Poisson} \\
+\hline
+\textbf{EMD} & 0.65 & 0.71 \\
+\textbf{KL\_2} & 0.62 & 0.79 \\
+\textbf{KL\_e} & 0.74 & 0.86 \\
+\textbf{JSD\_2} & 0.92 & 0.94 \\
+\textbf{JSD\_e} & 0.95 & 0.96
+\end{tabular}
+\end{center}
+\end{table}
 

### 2020-08-06 code/walmart_demand.py

@@ -433,8 +433,8 @@ def cb_width_model():
     fp['correction_factor'] = flags.IS_CONTINUOUS | flags.HAS_MISSING | flags.MISSING_NOT_LEARNED
 
     features = [
-        'yhat_mean_feature',
-        'correction_factor',
+#        'yhat_mean_feature',
+#        'correction_factor',
 #        'dayofweek',
         'store_id',
         'item_id',
@@ -515,9 +515,17 @@ def main(args):
     plot_cdf(X['n'][359323], X['p'][359323])
 
     X['cdf_truth'] = nbinom.cdf(X['y'], X['n'], X['p'])
-    cdf_accuracy(X['cdf_truth'][mask], 'wasserstein_2')
     X['cdf_truth_poisson'] = poisson.cdf(X['y'], X['yhat_mean'])
+    cdf_accuracy(X['cdf_truth'][mask], 'wasserstein_2')
     cdf_accuracy(X['cdf_truth_poisson'][mask], 'wasserstein_2')
+    cdf_accuracy(X['cdf_truth'][mask], 'kullback_2')
+    cdf_accuracy(X['cdf_truth_poisson'][mask], 'kullback_2')
+    cdf_accuracy(X['cdf_truth'][mask], 'kullback_e')
+    cdf_accuracy(X['cdf_truth_poisson'][mask], 'kullback_e')
+    cdf_accuracy(X['cdf_truth'][mask], 'jensen_2')
+    cdf_accuracy(X['cdf_truth_poisson'][mask], 'jensen_2')
+    cdf_accuracy(X['cdf_truth'][mask], 'jensen_e')
+    cdf_accuracy(X['cdf_truth_poisson'][mask], 'jensen_e')
 
 #    plot_cdf_truth(X['cdf_truth'][mask])
 #    plot_cdf_truth(X['cdf_truth_poisson'][mask])

### 2020-08-08 LaTex/paper.tex
improved model, adapted description

--- 2020-08-05 LaTex/paper.tex
@@
 the sales records, with specified quantile values.
-For example, if we wanted to check for the quantile value $0.3$, we would look at the observed quantiles for each prediction and calculate the fraction of these being lower than $0.3$, expecting a fraction of $0.3$ for accurate PDF predictions.
-In order to judge quality of the full PDF predictions, we calculate these fractions for a range of different specified quantile values.
-When choosing which quantile to compare it is important to cover the tails of the distribution to make sure that even relatively rare events are estimated correctly by the machine learning algorithm.
-We call this method quantile profile plot.
+In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example for the predicted demand for a single product sold on a single day in a given sales location.
+Each of these predicted negative binomial PDFs is then transformed to its CDF.
+Note that for simplicity, we always refer to the negative binomial model in this description, however, the general approach is valid for any PDF.
+
+Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample.
+For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values of observed sales records being lower/higher than $0.5$.
+In other words, in case of the median, 50\% of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50\% above.
+
+In order to judge whether the overall shape of the predicted distributions is predicted correctly, we repeat this procedure for a range of quantiles, for example $q = 0.1, 0.2, \ldots, 0.9$.
+However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say $q = 0.95$ and $q = 0.99$.
+In the following, we call this method {\em quantile profile plot}.
 
@@
 \ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
-Second, by showing the dependency from the (arbitrary) variable on the x-axis, the quantile profile plot reveals deviations of the predicted PDF from the acuals for different parts, e.g.
+Second, by showing the dependency from the (arbitrary) variable on the x-axis, the quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
 specific categories, of that variable.
@@
 However, in order to also quantify the quality of the PDF predictions, a metric is required that assesses the difference between the distributions.
-Therefore, we want to compare the CDF histogram of the predicted PDFs with the expected  uniform distribution and define a prediction accuracy interval between 0 and 1 where an accuracy of 1 indicates that both distributions agree perfectly.
+Therefore, we want to compare the CDF histogram of the predicted PDFs with the expected  uniform distribution and define a metric in the range between 0 and 1, such that the metric takes the value of 1 when both distributions agree perfectly.
 Several approaches that measure the difference between two probability distributions are suggested in the literature such as  the first Wasserstein distance \cite{olkin1982}, the Kullback-Leibler divergence \cite{kullback1951}, and the Jensen-Shannon divergence  \cite{dagan1997}.
@@
 We use data from a Kaggle online competition about estimating unit sales of Walmart retail goods \cite{kaggle_data} to demonstrate forecasting of demand for different product-location-date combinations in form of full individual PDFs by means of Cyclic Boosting and subsequent qualitative and quantitative evaluation of the PDF predictions.
-The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date, while the field sales represents the target to predict.
-In the following, we use data from 2013-01-01 to 2016-05-22 for all 10 available stores and 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3}.
-Besides the key and target fields (\texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}), we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.
+The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date.
+The target $y$ which we need to predict is the number of sales of a given product in a given store on a specific day, denoted by \texttt{sales}.
+In the following, we use data from 2013-01-01 to 2016-05-22, that describe the sales of 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3} in 10 stores.
+All data before 2016 are used as the training data and the data from 2016 are used as an independent test or validation set.
+Besides the fields used to identify an individual sales record and the corresponding observed sales value, namely \texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}, we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.
 
@@
 
-For our Cyclic Boosting mean estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several variables describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), and a flag denoting a promotion.
-We also include various two-dimensional combinations of the mentioned features, where one of the two dimensions being \texttt{store\_id} or \texttt{item\_id} to learn characteristics of individual locations and products.
-Furthermore, we include an exponential price-demand elasticity sub-estimator based on the ratio of reduced and normal price and individualized by \texttt{store\_id}.
-We plan to report on the inner workings of this method in a future work.
+The first model, based on the Cyclic Boosting approach described above, is used to predict the mean of an assumed negative binomial distribution, and we use the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time-series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
+We also include various two-dimensional combinations of these features.
+In these cases, one of the two dimensions is either \texttt{store\_id} or \texttt{item\_id}, allowing the machine learning model to learn characteristics of individual locations and products.
 
-\textcolor{red}{... training and prediction periods ... residual correction ...}
+Unlike most state-of-the-art time series forecasting methods, we do not include lagged target information, for example via stacking of exponentially weighted moving average features, in our model.
+The reason for this is that including such information makes the learning of exogenous effects, e.g.
+promotions or events, much harder, as the model tends to rely mainly on temporal confounding.
+Omitting these kind of variables improves the capability of the machine learning model to learn causal dependencies, which in turn improves the explainability of the model as well as the quality of the forecasts for mid- to long-term predictions and rare events.
+In order to capture recent trends that are not reflected in the exogenous features of the model, we apply an individual residual correction on each of the predictions of the machine learning model, which accounts for deviations between the exponentially weighted moving average (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past.
+To reflect a realistic replenishment scenario, we use a prediction horizon of two days for our example here, i.e.
+a target lag of two days for the training.
 
-\textcolor{red}{... results: mad, mean(y) ... example time series plot ...}
+\noindent
+Just to give a rough verification for the accuracy of the mean predictions, we report the MAD and MSE averaged over the full used data set in the test period to be $1.65$ and $10.15$, respectively, while the average of the target, i.e.
+the sales, is $3.28$.
+\fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.
 
@@
 \includegraphics[width=8cm]{../figures/mean_prediction}
-\caption{\label{fig:mean_prediction} \textcolor{red}{time series of mean prediction and actuals for one item in all stores}}
+\caption{\label{fig:mean_prediction} Time series of mean prediction and sales in test period summed over all products and stores.}
 \end{center}
@@
 
-For our Cyclic Boosting variance estimation model, we include the following features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, a flag denoting a promotion, and the insample prediction of the mean model, and the individual correction factor of the residual correction step described in \ref{sec:example_mean}.
+The second model, also based on Cyclic Boosting, is used to estimate the variance of the negative binomial distribution.
+The data are split into training and test set as above and in addition the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}.
+This effectively means that the mean predictions are created in-sample for the training period using the fully trained and validated model for the mean discussed above.
 
-\textcolor{red}{... same training and prediction periods ... no residual correction ...}
-
-\textcolor{red}{... pdf/cdf example for visualization...}
+In this model focusing on the variance, we use a different set of features: categorical variables for \texttt{store\_id} and \texttt{item\_id} and a flag denoting a promotion.
+An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_586} in store \texttt{WI\_1} on 2016-05-22 is shown in \fig \ref{fig:pdf_example}.
 
@@
 
-\textcolor{red}{...}
+\fig \ref{fig:cdf_demand} shows the histogram of CDF observations according to the method described in sec.
+\ref{sec:cdf_histo} for all product-location-day combinations in the test period.
+As benchmark, we compare the outcome of our negative binomial model to a simpler Poisson assumption, which has only a single model parameter, the mean.
+Using the same mean predictions for both negative binomial and Poisson model, the negative binomial PDF predictions are much closer to the uniform distribution, which we expect for optimal PDF predictions, than the Poisson PDF predictions, showing the effectiveness of our variance estimation.
+The only significant deviation of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$.
+As can be seen when excluding samples with mean predictions lower than $1.0$, this deviation stems from slow-sellers, which are prone to overdispersion.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/cdf_truth}
-\caption{\label{fig:cdf_demand} \textcolor{red}{CDF histogram}}
+\includegraphics[width=4cm]{../figures/cdf_truth_nbinom}
+\includegraphics[width=4cm]{../figures/cdf_truth_poisson}
+\includegraphics[width=4cm]{../figures/cdf_truth_nbinom_larger1}
+\includegraphics[width=4cm]{../figures/cdf_truth_poisson_larger1}
+\caption{\label{fig:cdf_demand} Histograms of ex post observed target CDF values of the corresponding individual PDF predictions (to be compared to a uniform distribution) for our negative binomial model (left) and a simpler Poisson model for comparison (right), using all product-location-day combinations in the test period (upper two plots).
+In order to show the effect of slow-sellers, the lower two plots exclude all samples with mean predictions lower than $1.0$.}
 \end{center}
 \end{figure}
+
+\fig \ref{fig:invquant_mean} and \fig \ref{fig:invquant_dayofweek} show quantile profile plots for different variables on the x-axis, namely mean predictions and day of week, respectively.
+All product-location-day combinations in the test period are used to generate the statistics in the plots.
+It can be seen that the different weekdays (from Monday as 0 to Sunday as 6) show slightly different patterns and that there are larger deviations for higher mean predictions (with fewer samples though).
 
@@
 \includegraphics[width=8cm]{../figures/invquant_dayofweek}
-\caption{\label{fig:invquant_dayofweek} \textcolor{red}{invquant profile for day of week of one product over all locations and days on x-axis}}
+\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the x-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period.}
 \end{center}
@@
 
-\textcolor{red}{... quantitative CDF accuracy ...}
+The quantitative results for the CDF accuracies of our PDF predictions for the different metrics described in sec.
+\ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf:acc}.
+As benchmark, we compare these to the results of a Poisson model assumption needing only the mean as single parameter.
+Our negative binomial PDF predictions show a significant improvement over the simpler Poisson model, which uses the same mean predictions, for each of the metrics.
+
+\begin{table}[h!]
+\begin{center}
+\caption{CDF accuracies for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
+\label{tab:cdf:acc}
+\begin{tabular}{c|c|c}
+ & \textbf{NBD} & \textbf{Poisson} \\
+\hline
+\textbf{EMD} & 0.965 & 0.852 \\
+\textbf{KL\_2} & 0.995 & 0.933 \\
+\textbf{KL\_e} & 0.997 & 0.954 \\
+\textbf{JSD\_2} & 0.999 & 0.982 \\
+\textbf{JSD\_e} & 0.999 & 0.987
+\end{tabular}
+\end{center}
+\end{table}
 

### 2020-08-08 code/walmart_demand.py

@@ -73,14 +73,14 @@ def plot_cdf(n, p):
     plt.savefig('plots/cdf.pdf')
 
 
-def plot_cdf_truth(cdf_truth):
+def plot_cdf_truth(cdf_truth, suffix):
     plt.figure()
-    plt.hist(cdf_truth, bins=20)
+    plt.hist(cdf_truth, bins=30)
     plt.tight_layout()
-    plt.savefig('plots/cdf_truth.pdf')
+    plt.savefig('plots/cdf_truth_' + suffix + '.pdf')
 
 
-def plot_invquants(X, variable):
+def plot_invquants(X, variable, suffix):
     means_result = binned_statistic(X[variable], [X['cdf_truth']<=0.1, X['cdf_truth']<=0.3, X['cdf_truth']<=0.5, X['cdf_truth']<=0.7, X['cdf_truth']<=0.9, X['cdf_truth']<=0.97], bins=20, statistic='mean')
     means10, means30, means50, means70, means90, means97 = means_result.statistic
 
@@ -96,7 +96,7 @@ def plot_invquants(X, variable):
     plt.errorbar(x=bin_centers, y=means97, linestyle='none', marker='s')
     plt.hlines([0.1, 0.3, 0.5, 0.7, 0.9, 0.97], bin_edges[0], bin_edges[-1])
     plt.tight_layout()
-    plt.savefig('plots/invquant_' + variable + '.pdf')
+    plt.savefig('plots/invquant_' + variable + '_' + suffix + '.pdf')
 
 
 def plot_timeseries(df, suffix, title=''):
@@ -125,12 +125,6 @@ def plotting(df, suffix=''):
     for name, group in predictions_grouped:
         ts_data = group.groupby(['date'])['y', 'yhat_mean'].sum().reset_index()
         plot_timeseries(ts_data, 'item_' + str(name) + suffix)
-
-
-def transform_nbinom(mean, var):
-    p = np.minimum(np.where(var > 0, mean / var, 1 - 1e-8), 1 - 1e-8)
-    n = np.where(var > 0, mean * p / (1 - p), 1)
-    return n, p
 
 
 def eval_results(yhat_mean, y):
@@ -309,7 +303,7 @@ def feature_properties():
     fp['dayofyear'] = flags.IS_CONTINUOUS | flags.IS_LINEAR
     fp['weekofmonth'] = flags.IS_ORDERED
[…]

### 2020-08-08 LaTex/paper.tex
small feature set fix

@@
 
-In this model focusing on the variance, we use a different set of features: categorical variables for \texttt{store\_id} and \texttt{item\_id} and a flag denoting a promotion.
+In this model focusing on the variance, we use the same set of features as for the mean model described above in sec.
+\ref{sec:example_mean}, except for dropping the two-dimensional combinations including individual event features.
 An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_586} in store \texttt{WI\_1} on 2016-05-22 is shown in \fig \ref{fig:pdf_example}.

### 2020-08-14 LaTex/paper.tex
some more explanations and cosmetics

@@
 
-\newcommand{\fig}{Fig. }
-\newcommand{\eqn}{Eqn. }
-\newcommand{\tab}{Tab. }
-\newcommand{\wrt}{w.r.t. }
-\newcommand{\etal}{ {\em et al. }}
+\newcommand{\fig}{Fig.}
+\newcommand{\eqn}{Eqn.}
+\newcommand{\tab}{Tab.}
+\newcommand{\wrt}{w.r.t.}
+\newcommand{\etal}{ {\em et al.}}
 
@@
 Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration.
-The current prediction, $\hat{y}_\tau$, is calculated according to eqn.
-\ref{eqn:cb} with the current values of the aggregated factors $f$:
+The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
 \begin{equation} \label{factors3}
@@
 54]{casella2002statistical}.
-In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the inverse quantiles or values of the CDF is filled using random numbers according to the intervals of the CDF.
-For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram is randomly chosen in the interval $[0,0.75]$.
-Proceeding similarly for all other observed values, the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.
+In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF.
+For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
+Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g.
+for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.
 
@@
 \begin{center}
-\includegraphics[scale=0.4]{../figures/CDFHistos_uniform} \includegraphics[scale=0.4]{../figures/CDFHistos_broad} \includegraphics[scale=0.4]{../figures/CDFHistos_narrow} \includegraphics[scale=0.4]{../figures/CDFHistos_over} \includegraphics[scale=0.4]{../figures/CDFHistos_under}
+\includegraphics[width=2.5cm]{../figures/CDFHistos_uniform}
+\includegraphics[width=2.5cm]{../figures/CDFHistos_broad}
+\includegraphics[width=2.5cm]{../figures/CDFHistos_narrow}
+\includegraphics[width=2.5cm]{../figures/CDFHistos_over}
+\includegraphics[width=2.5cm]{../figures/CDFHistos_under}
 \caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
@@
 Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample.
-For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values of observed sales records being lower/higher than $0.5$.
+For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec.
+\ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$.
 In other words, in case of the median, 50\% of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50\% above.
@@
 specific categories, of that variable.
-Two examples for this can be found in figures \ref{fig:invquant_mean} and \ref{fig:invquant_dayofweek} in the next section.
+Two examples for this can be found in Figs.
+\ref{fig:invquant_mean} and \ref{fig:invquant_dayofweek} in the next section.
 
@@
 
-We use data from a Kaggle online competition about estimating unit sales of Walmart retail goods \cite{kaggle_data} to demonstrate forecasting of demand for different product-location-date combinations in form of full individual PDFs by means of Cyclic Boosting and subsequent qualitative and quantitative evaluation of the PDF predictions.
-The fields identifying a unique sample of the data set are store (\texttt{store\_id}) and item identifiers (\texttt{item\_id}) as well as date.
-The target $y$ which we need to predict is the number of sales of a given product in a given store on a specific day, denoted by \texttt{sales}.
-In the following, we use data from 2013-01-01 to 2016-05-22, that describe the sales of 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3} in 10 stores.
+In the following, we describe how to use the approach outlined above in a practical setting.
+We use a public dataset obtained from a Kaggle online competition focusing in estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual SKUs for specific stores on specific days.
+For each demand forecast, Cyclic Boosting is used to predict the full probability density distribution of the expected demand at a granularity of (SKU, store, day) and use the methods described in Sec.
+\ref{sec:pdfEvaluation} to evaluate the quality of the individual forecasts.
+Each data record corresponding to an observed sales record is described by the following fields: the identifier of an individual store store (\texttt{store\_id}), the product identifier (\texttt{item\_id}) as well as the date.
+The target $y$, that we need to predict, is the number of sales of a given product in a given store on a specific day, denoted by \texttt{sales}.
+
+For our experiments, we use data from 2013-01-01 to 2016-05-22, that describe the sales of 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3} in 10 stores.
 All data before 2016 are used as the training data and the data from 2016 are used as an independent test or validation set.
-Besides the fields used to identify an individual sales record and the corresponding observed sales value, namely \texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}, we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} to build features for our machine learning models.
+Besides the fields used to identify an individual sales record and the corresponding observed sales value, namely \texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}, we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} in the forecasts and multiple features are built from these variables that are then used in the machine learning models.
 
@@
 
-The first model, based on the Cyclic Boosting approach described above, is used to predict the mean of an assumed negative binomial distribution, and we use the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time-series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
+As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event.
+As a first step, we use Cyclic Boosting to predict the mean of the distribution.
+
+\noindent
+This model uses the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time-series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
 We also include various two-dimensional combinations of these features.
@@
 Unlike most state-of-the-art time series forecasting methods, we do not include lagged target information, for example via stacking of exponentially weighted moving average features, in our model.
-The reason for this is that including such information makes the learning of exogenous effects, e.g.
-promotions or events, much harder, as the model tends to rely mainly on temporal confounding.
-Omitting these kind of variables improves the capability of the machine learning model to learn causal dependencies, which in turn improves the explainability of the model as well as the quality of the forecasts for mid- to long-term predictions and rare events.
+Although common in practice, including such variables makes the learning of exogenous effects, such as product promotions or the occurrence of special events, much harder.
+This is because when those variables are included, machine learning models tend to rely mainly on temporal confounding.
+Therefore, omitting these kind of variables improves the capability of the machine learning model to learn causal dependencies, which in turn improves the explainability of the model as well as the quality of the forecasts for mid- to long-term predictions and rare events.
 In order to capture recent trends that are not reflected in the exogenous features of the model, we apply an individual residual correction on each of the predictions of the machine learning model, which accounts for deviations between the exponentially weighted moving average (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past.
-To reflect a realistic replenishment scenario, we use a prediction horizon of two days for our example here, i.e.
-a target lag of two days for the training.
+We use the model to predict the expected demand two days into the future to reflect a realistic replenishment scenario, meaning that we use a target lag of two days for the training.
 
 \noindent
-Just to give a rough verification for the accuracy of the mean predictions, we report the MAD and MSE averaged over the full used data set in the test period to be $1.65$ and $10.15$, respectively, while the average of the target, i.e.
-the sales, is $3.28$.
+We use the mean absolute deviation (MAD) as well as the mean squared error (MSE) as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean.
+These metrics do not take the shape of the underlying probability distribution into account, but only compare the predicted mean to the observed number of sales.
+Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.15$.
+The mean of the target, i.e.
+the observed sales, is $3.28$ for this period.
 \fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.
@@
 
+\noindent
 In this model focusing on the variance, we use the same set of features as for the mean model described above in sec.
@@
 The only significant deviation of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$.
-As can be seen when excluding samples with mean predictions lower than $1.0$, this deviation stems from slow-sellers, which are prone to overdispersion.
+These cases are due to slow moving articles with a predicted mean lower than $1.0$, i.e.
+less than one item sold per store per day on average.
+These items have been removed in the lower part of  \fig \ref{fig:cdf_demand} and the resulting cumulative distribution is flat as expected.
+In practice, predicting the demand for these kind of items is very challenging as they are prone to over-dispersion and large fluctuations due to the discrete nature of the sales event.
 
@@
 
-\fig \ref{fig:invquant_mean} and \fig \ref{fig:invquant_dayofweek} show quantile profile plots for different variables on the x-axis, namely mean predictions and day of week, respectively.
-All product-location-day combinations in the test period are used to generate the statistics in the plots.
-It can be seen that the different weekdays (from Monday as 0 to Sunday as 6) show slightly different patterns and that there are larger deviations for higher mean predictions (with fewer samples though).
+Using the profile plots introduced earlier, the prediction quality of the full predicted probability distribution can be assessed in more detail.
+\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 10 columns correspond to mean predictions in the intervals $[0.0, 5.0]$, $(5.0, 10.0]$, $(10.0, 15.0]$, $(15.0, 20.0]$, $(20.0, 30.0]$, $(30.0, 40.0]$, $(40.0, 50.0]$, $(50.0, 60.0]$, $(60.0, 80.0]$, $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
+The considered quantile values are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
+We can see that for mean predictions higher than $80$, our negative binomial model deviates significantly from the expected behaviour, with the shape of the deviations pointing towards a general underprediction (see  \fig \ref{fig:quant_profiles}).
+In a real-live situation with a live supply chain project for a customer, a more detailed investigation of the root-causes would then start.
+However, it should be noted that due to this segregation, the statistics in each part of the relevant test sample becomes a limiting factor as well.
+This plot also illustrates the benefits of using the quantile profile plot, since we have seen that using more conventional approaches, such as \fig \ref{fig:mean_prediction} or even the lower part of \fig \ref{fig:cdf_demand}, the predictions appear reasonable, even when not relying on simple point metrics such as MAD or MSE.
+It is therefore paramount both to predict the full probability distribution instead of just a point estimate, such as the mean, as well as verifying a number of quantiles for each prediction to assess the quality of the prediction thoroughly.
+In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the used quantile values are the same as in \fig \ref{fig:invquant_mean}.
+Aggregating over all stores, items, and sales dates (independently for each day of the week), we can see that each quantile is predicted relatively well.
 
@@
 
-The quantitative results for the CDF accuracies of our PDF predictions for the different metrics described in sec.
-\ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf:acc}.
-As benchmark, we compare these to the results of a Poisson model assumption needing only the mean as single parameter.
-Our negative binomial PDF predictions show a significant improvement over the simpler Poisson model, which uses the same mean predictions, for each of the metrics.
+The quantitative results for the CDF accuracy of our PDF predictions for the different metrics described in sec.
+\ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf_acc}.
+As benchmark, we again compare against a Poisson model assumption using the same mean predictions.
+As expected from the qualitative findings in \fig \ref{fig:cdf_demand}, the negative binomial PDF predictions show a significant improvement over the simpler Poisson model.
 
@@
 \begin{center}
-\caption{CDF accuracies for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
-\label{tab:cdf:acc}
+\caption{CDF accuracy for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
+\label{tab:cdf_acc}
 \begin{tabular}{c|c|c}
@@
 \hline
-\textbf{EMD} & 0.965 & 0.852 \\
+\textbf{EMD} & 0.965 & 0.851 \\
 \textbf{KL\_2} & 0.995 & 0.933 \\
@@
 \section{Conclusion}
+Demand forecasting remains a crucial step in operational planning for retailers.
+Both from practical and theoretical perspectives, disentangling the forecast of demand from the operational decision making regarding order quantities has potentially significant advantages over an integrated approach using data to predict the resulting order quantities directly.
+When predicting future demand, both model-free and a model-based approaches can be used, where the model-based approach is generally more robust, as it is rooted in a theoretical understanding of the sales process.
+Compared to a model-free approach like quantile regression, the distributional assumption can drastically reduce the uncertainty of the resulting predictions.
+Using the Cyclic Boosting machine learning approach, the full probability density distribution can be predicted in a fully explainable way on the individual level.
+This allows to extract not only simple point estimates, such as the expected mean demand, but all relevant quantiles of the distribution.
 
-We have shown how to use two subsequent machine learning models for mean and variance estimation together with a negative binomial model assumption to come up with individual PDF predictions.
-This setup is especially useful for retail demand forecasting, where the true demand follows a negative binomial distribution quite closely.
-Compared to a model-free approach like quantile regression, the distributional assumption therefore drastically reduces the uncertainty of the resulting predictions, which are, by using Cyclic Boosting as underlying machine learning algorithm, fully explainable on the individual level.
-
-Furthermore, we have presented new qualitative and quantitative methods for evaluating predictions in form of full PDFs.
-For qualitative evaluation, it is important to check the full PDF, especially its tails that are often crucial for subsequent decision making, as well as investigate aggregations of individual PDFs (to reduce the uncertainty of the evaluation) dependent on specific variables, and we have proposed a novel profile approach called quantile profile plot for this.
-For quantitative evaluation, it is always desirable to have a single number to compare different models in terms of accuracy, and we have suggested to use the deviance between the CDF histogram of the predicted PDFs and the uniform distribution to compare PDF accuracies.
+Evaluating these PDF predictions becomes the critical question.
+Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions.
+Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general.
+Instead, both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions.
+Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of the probability integral transformation and the expected uniform distribution.
 

### 2020-08-14 code/walmart_demand.py

@@ -61,6 +61,7 @@ def plot_pdf(n, p):
     plt.figure()
     x = range(30)
     plt.plot(x, nbinom.pmf(x, n, p))
+    plt.xlabel("sales")
     plt.tight_layout()
     plt.savefig('plots/pdf.pdf')
 
@@ -69,6 +70,7 @@ def plot_cdf(n, p):
     plt.figure()
     x = np.linspace(0.0, 30.0, num=100)
     plt.plot(x, nbinom.cdf(x, n, p))
+    plt.xlabel("sales")
     plt.tight_layout()
     plt.savefig('plots/cdf.pdf')
 
@@ -76,25 +78,48 @@ def plot_cdf(n, p):
 def plot_cdf_truth(cdf_truth, suffix):
     plt.figure()
     plt.hist(cdf_truth, bins=30)
+    plt.xlabel("CDF values")
+    plt.ylabel("count")
     plt.tight_layout()
     plt.savefig('plots/cdf_truth_' + suffix + '.pdf')
 
 
-def plot_invquants(X, variable, suffix):
-    means_result = binned_statistic(X[variable], [X['cdf_truth']<=0.1, X['cdf_truth']<=0.3, X['cdf_truth']<=0.5, X['cdf_truth']<=0.7, X['cdf_truth']<=0.9, X['cdf_truth']<=0.97], bins=20, statistic='mean')
+def plot_invquants(X, variable, suffix, continuous=False):
+    cols = [X['cdf_truth']<=0.1, X['cdf_truth']<=0.3, X['cdf_truth']<=0.5, X['cdf_truth']<=0.7, X['cdf_truth']<=0.9, X['cdf_truth']<=0.97]
+    if continuous:
+        bins = [0., 5., 10., 15. ,20., 30., 40., 50., 60., 80., 100.]
+    else:
+        bins = 100
+    means_result = binned_statistic(X[variable], cols, bins=bins, statistic='mean')
+    std_result = binned_statistic(X[variable], cols, bins=bins, statistic='std')
     means10, means30, means50, means70, means90, means97 = means_result.statistic
+    std10, std30, std50, std70, std90, std97 = std_result.statistic
+    std10 = std10 / np.sqrt(len(X))
+    std30 = std30 / np.sqrt(len(X))
+    std50 = std50 / np.sqrt(len(X))
+    std70 = std70 / np.sqrt(len(X))
+    std90 = std90 / np.sqrt(len(X))
+    std97 = std97 / np.sqrt(len(X))
 
[…]

### 2020-08-15 code/walmart_demand.py
forgot code

@@ -92,25 +92,27 @@ def plot_invquants(X, variable, suffix, continuous=False):
         bins = 100
     means_result = binned_statistic(X[variable], cols, bins=bins, statistic='mean')
     std_result = binned_statistic(X[variable], cols, bins=bins, statistic='std')
+    count_result = binned_statistic(X[variable], cols, bins=bins, statistic='count')
     means10, means30, means50, means70, means90, means97 = means_result.statistic
     std10, std30, std50, std70, std90, std97 = std_result.statistic
-    std10 = std10 / np.sqrt(len(X))
-    std30 = std30 / np.sqrt(len(X))
-    std50 = std50 / np.sqrt(len(X))
-    std70 = std70 / np.sqrt(len(X))
-    std90 = std90 / np.sqrt(len(X))
-    std97 = std97 / np.sqrt(len(X))
+    count10, count30, count50, count70, count90, count97 = count_result.statistic
+    stdmean10 = std10 / np.sqrt(count10)
+    stdmean30 = std30 / np.sqrt(count30)
+    stdmean50 = std50 / np.sqrt(count50)
+    stdmean70 = std70 / np.sqrt(count70)
+    stdmean90 = std90 / np.sqrt(count90)
+    stdmean97 = std97 / np.sqrt(count97)
 
     bin_edges = means_result.bin_edges
     bin_centers = (bin_edges[:-1] + bin_edges[1:]) / 2.
 
     plt.figure()
-    plt.errorbar(x=bin_centers, y=means10, yerr=std10, linestyle='none', marker='v')
-    plt.errorbar(x=bin_centers, y=means30, yerr=std30, linestyle='none', marker='<')
-    plt.errorbar(x=bin_centers, y=means50, yerr=std50, linestyle='none', marker='.')
-    plt.errorbar(x=bin_centers, y=means70, yerr=std70, linestyle='none', marker='>')
-    plt.errorbar(x=bin_centers, y=means90, yerr=std90, linestyle='none', marker='^')
-    plt.errorbar(x=bin_centers, y=means97, yerr=std97, linestyle='none', marker='s')
+    plt.errorbar(x=bin_centers, y=means10, yerr=stdmean10, linestyle='none', marker='v')
+    plt.errorbar(x=bin_centers, y=means30, yerr=stdmean30, linestyle='none', marker='<')
+    plt.errorbar(x=bin_centers, y=means50, yerr=stdmean50, linestyle='none', marker='.')
[…]

### 2020-08-17 LaTex/paper.tex
reviewed abstract and introduction

@@
 \begin{abstract}
-Demand forecasting is a central component for many aspects of supply chain operations as it provides crucial input for subsequent decision making like ordering processes.
+Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes.
 While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
@@
 While metrics evaluating simple point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods.
-Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
+Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice, as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
 \end{abstract}
@@
 
-Demand forecasting is one of the main challenges for retailers and at the core of business operations.
-Due to its stochastic nature, demand is difficult to forecast as it depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
-In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions used in the replenishment process of the retailer.
+Demand forecasting is one of the main challenges for retailers and at the core of supply chain business operations.
+Due to its stochastic nature, demand is difficult to forecast.
+It depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
+In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer.
 Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.
-The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision: Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers.
-Ordering too many items results in excess inventory which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
+
+The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
+Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers.
+Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
 This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
@@
 
-The availability of Big Data allows capturing, storing and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information as well as a wide range of further variables.
-Modern machine learning algorithms can then be used to predict the per-item demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision.
-It is important to note that demand as a random variable is not identically and independently distributed (i.i.d.).
-While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity and other factors.
+The availability of Big Data allows capturing, storing, and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as a wide range of further variables.
+Modern machine learning algorithms can then be used to predict the per-item-location demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision.
+Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations.
+While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors.
 The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy.
 This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required.
-For example, the probability distribution governing the demand of a particular item is specific to the item, date and retail location for which the forecast is made and depends on a wide range of further influencing factors.
+For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.
 
@@
 \ref{sec:pdfEstimation}.
-After that we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec.
+After that, we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec.
 \ref{sec:pdfEvaluation}.

### 2020-08-18 LaTex/paper.tex
reviewed literature section

@@
 The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
-Ordering too few items may result in stock-out situations leading to unrealized demand and unsatisfied customers.
+Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
 Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
@@
 
-Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand  distribution or an integrated approach that seeks to map directly from the available data (historic sales records and further variables) to the ordering decision.
-This approach is often referred to as "data-driven newsvendor'' and discussed  e.g.
-in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
-It aims to avoid estimating the underlying probability distribution for demand and use the available data to derive the operational decisions (the order quantity) directly.
-An overview of a range approaches can also be found in \cite{huber2019data}.
-
-Although this approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision (the order quantity), it faces several drawbacks.
-First, the full probability distribution for the demand of a specific item at a given sales location and business day includes all available information including the uncertainty of the modelled demand.
-This can be used to simulate the performance on a per-item level and e.g.
-optimize the impact on business strategy decisions on conflicting metrics such as stock out- and waste-rate.
-By forecasting the full demand distribution as opposed to point estimators such as the expected demand or the median of the distribution, the forecast quality can be evaluated for quantiles of the distribution, including the often extensive tails of the distribution.
-Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e.
-the mechanism behind the stochastic behavior of the customer demand.
-However, modelling demand directly is vital if a causal analysis is planned at a later stage or independently, for example to study the effect promotions, advertisements,  price changes or other influencing factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
-Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy of the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
-From a practical perspective, separating the demand forecast from the operational decisions (i.e.
-calculating the  order quantities for the next delivery cycle) also allows to evaluate longer-term planning and reduces the complexity as it avoids coupling the complex delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
-This also allows to share long-term demand predictions with other business units or external vendors  and wholesalers to ease their planning for the production and supply-chain processes upstream of the retailer.
-From the perspective of industrial practice, modelling the demand separately from deriving the subsequent orders  has the additional benefit that multiple retail chains can benefit from any improvement in the model description even if the concrete retailers are unrelated to each other.
-For example, if a particular effect was identified at some retailer A and included in the machine learning model, the improvement can be rolled out immediately or on request to all other retailers using this system on a planetary scale without having to adapt the underlying machine learning model for each retailer individually.
-
-In contrast to the direct mapping the observed data to ordering decisions, more traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g.
+Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision.
+More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g.
 \cite{silver1998} for an overview.
@@
 Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}.
-Perishable goods are well described by the "newsvendor-problem'' \cite{Edgeworth}, where in the simplest case all stock perishes at the end of the selling period (e.g.
+Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g.
 a business day).
@@
 Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.
+The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g.
+in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
+It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e.
+the order quantity) directly.
+An overview of a range of different approaches can also be found in \cite{huber2019data}.
+Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages.
+First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g.
+optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e.
+lost sales) and waste rate.
+From a practical perspective, separating the demand forecast from the operational decisions (i.e.
+calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
+It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer.
+From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other.
+Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e.
+the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
+Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
 
-The main objective in any of these approaches is to determine the underlying demand distribution.
+So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution.
 The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g.
@@
 the calculation of the optimal order quantity.
+Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately.
+Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model.
+This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case.
+And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level.
+Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method.
+Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.
 
-However, in order to be able to fully optimize the operational decision, it is critical that one indeed reconstructs a full demand distribution.
-This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine-learning approaches, will not suffice.
-Additionally, we need to consider that demand is not i.i.d., but depends on external factors such as season, weather, product life-cycle, and is also actively changed by the retailer by setting a specific price, offering rebates or running advertisements.
-Additionally, the demand implicitly depends on the location of the retail outlet as well as the specifics of that location, such as product  assortment influencing the choice of possible replacement articles and many more.
-These complications are the main reason we cannot treat the replenishment process as $n$ independent newsvendor-type problems.
-
+In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution.
+This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice.
 Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc.
 into account.
-This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution from which an empirical probability distribution can be interpolated.
-Using a dedicated neural network \cite{Feindt2006190}, either the  full probability distribution or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
+This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
+Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
 Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}.
-This approach is computationally favourable, as fewer parameters need to be estimated compared to the case of the full distribution.
+This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated.
 Empirically, one can determine the best fitting distribution from data \cite{adan1995}.
@@
 The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}.
-This distribution arises if the Poisson parameter $\mu$ is a random variable itself that follows a Gamma distribution.
-The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution for which $\mu = \sigma^2$.
+This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
+The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$.
 Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
@@
 This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer.
-Typically this means that the full demand distribution has to be estimated per SKU for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
-In contrast to using a "black-box'' machine learning model, using Cyclic Boosting allows to interpret how each individual prediction was made and what the most important variables are per individual prediction.
+Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
+In contrast to using a "black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.
 
 Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution.
-This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as  order quantities as well as strategic business decisions by the retailer or gain further insights into customer behavior using for example causal modelling.
+This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as order quantities as well as strategic business decisions by the retailer, and to gain further insights into customer behavior using for example causal modeling.
 

### 2020-08-19 LaTex/paper.tex
reviewed Cyclic Boosting section

@@
 Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.
+However, it should be noted that demand is not independent between different samples for a given date.
+For example, a promotion applied to one product can affect the sales of related products within the assortment.
+This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.
+
 The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g.
@@
 \cite{shapiro2014} for an overview.
-However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price or advertisements influence, and therefore are highly correlated with, the demand.
+However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price, advertisements, or weather influence, and therefore are highly correlated with, the demand.
 Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e.
@@
 To predict an individual PDF using a parametric approach, one has to rely on a model assumption about the underlying distribution of the random variable to be predicted.
-As discussed earlier, the negative binomial probability distribution (NBD) is well routed in theoretical arguments to model customer demand.
-Its parameters can be modelled by two independent models, one to estimate the mean and the other for the variance.
+As discussed earlier, the NBD is well routed in theoretical arguments to model customer demand.
+Its parameters can be modeled by two independent models, one to estimate the mean and the other for the variance.
 At least in principle, any method can be used.
-However, in the case of demand forecasting, each prediction is highly specific to the circumstances in which it is used (such as SKU, opening day, and store location) and may depend on a multitude of describing variables or features such as sales location, weather, price, and so forth.
-It should be noted that these parameters are not independent between products.
-For example, a promotion applied to one product can, at least in principle, affect the sales of related products within the assortment.
-This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modelled holistically.
-Machine learning algorithms are ideally suited for this task and in the following we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches.
+However, as discussed in sec.
+\ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches.
 Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm.
-This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance.
-The features may or may not differ between the mean and variance estimation models.
+This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance, and the features may or may not differ between the mean and variance estimation models.
 The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.
 
-In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD models using Cyclic Boosting.
+In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD using two Cyclic Boosting models.
 
@@
 
-Cyclic Boosting \cite{Wick2019} is a type of generalized additive models using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters.
-Major benefits of Cyclic Boosting are its accuracy, performance even at large scale and providing fully explainable predictions which are of vital importance in practical applications, in particular for multi-national retail vendors.
+Cyclic Boosting is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters.
+Major benefits of Cyclic Boosting are its accuracy, performance, even at large scale, and providing fully explainable predictions, which are of vital importance in practical applications.
 
 The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
-The global mean $\mu$ is determined from all target values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
-Single data records, for example the sales of a specific SKU along with all relevant features, are indexed by $i$.
+The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
+Single data records, for example the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
 The individual predictions $\hat{y_i}$  can then be calculated as:
@@
 
-When looking at the demand of individual SKUs, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
+When looking at the demand of individual product-location-date combinations, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
 \begin{equation} \label{eqn:nbinom}

### 2020-08-27 LaTex/paper.tex
reworked model and lots of descriptions

@@
 Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes.
-While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often mere point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
-Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions in form of negative binomial distributions can be predicted instead of simple point estimates.
-While metrics evaluating simple point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods.
-Additionally, each single prediction obtained with this framework is explainable, which is a major benefit in practice, as individual forecasts can be understood by the practitioner and ``black-box'' models can be avoided.
+While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often just point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
+Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted instead of single numbers.
+While metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods.
+Additionally, each single prediction obtained with this framework is explainable.
+This is a major benefit in particular for practitioners, as this allows them to avoid ``black-box'' models and understand the contributing factors for each individual prediction.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
 
-Demand forecasting is one of the main challenges for retailers and at the core of supply chain business operations.
-Due to its stochastic nature, demand is difficult to forecast.
-It depends on many influencing factors and the realized demand can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
-In order to make operational decisions, an optimal point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer.
+Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations.
+Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
 Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.
-
-The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
-Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
-Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
-This situation is particularly noticeable in the so called "ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
-These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item.
-In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured.
-Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled.
-The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from "never-out-of-stock'' items to a service level exceeding e.g.
-90\%.
-
-The availability of Big Data allows capturing, storing, and processing a vast amount of data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as a wide range of further variables.
-Modern machine learning algorithms can then be used to predict the per-item-location demand distribution, corrected for censored data, from which an optimal point estimator can be derived to be used in the subsequent ordering decision.
 Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations.
@@
 
+Thanks to the availability of Big Data today, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed.
+Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data.
+Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately.
+Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them.
+This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series.
+Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.
+
+In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer.
+The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
+Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
+Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
+This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
+These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item.
+In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured.
+Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled.
+The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90\%.
+So, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.
+
 The remainder of the paper is organized as follows: We first review the relevant literature and existing work in sec.
@@
 \ref{sec:pdfEstimation}.
-After that, we describe methods for the qualitative and quantitative evaluation of PDF predictions in sec.
+After that, we describe novel techniques for the qualitative and quantitative evaluation of PDF predictions in sec.
 \ref{sec:pdfEvaluation}.
@@
 Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}.
-Perishable goods are well described by the "newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g.
+Perishable goods are well described by the ``newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g.
 a business day).
@@
 For example, a promotion applied to one product can affect the sales of related products within the assortment.
-This implies that the demand forecasts cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.
+This implies that the demand forecasts for multiple items or a given assortment range cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.
 
-The direct approach is often referred to as "data-driven newsvendor'' and discussed e.g.
+The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g.
 in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
-It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e.
-the order quantity) directly.
+It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly.
 An overview of a range of different approaches can also be found in \cite{huber2019data}.
 Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages.
-First, demand forecasts in form of PDFs can be used to simulate the performance on an individual level, and e.g.
-optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e.
-lost sales) and waste rate.
-From a practical perspective, separating the demand forecast from the operational decisions (i.e.
-calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
+First, demand forecasts in form of full PDFs can be used to simulate the performance on an individual level, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate.
+From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
 It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer.
-From the perspective of industrial practice, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the concrete retailers are unrelated to each other.
-Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e.
-the mechanism behind the stochastic behavior of the customer demand, which is vital if a causal analysis is planned, for example a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
-Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that may not be able to be fully controlled for.
+From the perspective of the industrial practice of a supplier of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other.
+Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand.
+This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
+Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.
 
-So, the main objective in the solution of the inventory control problem is to determine the underlying demand distribution.
-The simplest approach is to just use the observed sales events and forecast these as a time series (see e.g.
-\cite{alwan2016}) or via sample average approximation (SAA), see e.g.
-\cite{shapiro2014} for an overview.
-However, these approaches do not make use of any data apart from the sales record themselves, although we know that many variables such as price, advertisements, or weather influence, and therefore are highly correlated with, the demand.
-Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e.
-the calculation of the optimal order quantity.
-Often, there are many individual time series, for example millions of item-store combinations to be forecasted daily in retail demand forecasting, and traditional, univariate time series forecasting methods, like autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}, operate on each of these time series separately.
-Machine learning, on the contrary, optimizes all individual time series together, improving the forecast quality of the individual time series by exploiting commonalities between them and therefore reducing the variance of the model.
-This can be interpreted as kind of hierarchical learning, one example of a hierarchical level being the sum of product sales over all locations in the retail demand forecasting use case.
-And this reduction of variance does not come with an increase of bias, as usual with the bias-variance tradeoff, because we can still take into account the effects on the finest level.
-Besides the reduction in variance, this modeling on various hierarchical levels can also be interpreted as kind of transfer learning, improving the generalizability of the method.
-Furthermore, machine learning enables a natural way to include many exogenous variables, simply as features, what reduces also the bias of the model.
+Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution.
+The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see e.g.
+\cite{shapiro2014} for an overview, or as a time series (see e.g.
+\cite{alwan2016}), for example by means of autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}.
+However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements, or weather influence, and others are highly correlated with the demand.
+Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity.
+As described in sec.
+\ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches and therefore are the method of choice in most use cases today (see e.g.
+\cite{tsforecasting,ZHANG199835,rnn,transformer}).
 
@@
 Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
-In contrast to using a "black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.
+In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.
 
@@
 \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches.
-Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ''off-the-shelf'' machine learning algorithm.
-This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance, and the features may or may not differ between the mean and variance estimation models.
+Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm.
+This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance.
+Hereby, the features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model.
 The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.
@@
 The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
-Single data records, for example the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
+Single data records, for example, the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
 The individual predictions $\hat{y_i}$  can then be calculated as:
@@
 \end{equation}
-The factors $f^k_j$ are the model parameters that are determined iteratively from the features, and are determined iteratively until the algorithm converges.
+The factors $f^k_j$ are the model parameters that are determined iteratively from the features until the algorithm converges.
 During training, regularization techniques are applied to avoid overfitting and improve the generalization ability of the algorithm.
@@
 For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
-\item{Quit when stopping criteria, e.g.
-the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
+\item{Quit when stopping criteria, e.g., the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
 \end{enumerate}
@@
 In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model.
-In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm like described in the following.
-
-When looking at the demand of individual product-location-date combinations, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
+In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm as follows: When looking at the demand of individual product-location-date combinations (meaning the sales record of a specific item sold on a specific day at a specific sales location), the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
 \begin{equation} \label{eqn:nbinom}
@@
 \begin{equation} \label{eqn:r}
-\hat{r_i} = 1 + \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
+\hat{r_i} = 1 +  \frac{1}{\prod \limits_{j=1}^p f^k_j} \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
 \end{equation}
@@
 Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec.
-\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example using a random search.
+\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example, using a random search.
 All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
@@
 \begin{center}
-\includegraphics[width=2.5cm]{../figures/CDFHistos_uniform}
-\includegraphics[width=2.5cm]{../figures/CDFHistos_broad}
-\includegraphics[width=2.5cm]{../figures/CDFHistos_narrow}
-\includegraphics[width=2.5cm]{../figures/CDFHistos_over}
-\includegraphics[width=2.5cm]{../figures/CDFHistos_under}
+\includegraphics[width=2.5cm]{../figures/cdf_truth_uniform}
+\includegraphics[width=2.5cm]{../figures/cdf_truth_broad}
+\includegraphics[width=2.5cm]{../figures/cdf_truth_narrow}
+\includegraphics[width=2.5cm]{../figures/cdf_truth_over}
+\includegraphics[width=2.5cm]{../figures/cdf_truth_under}
 \caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
@@
 We now refine the method from sec.
-\ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e.
-the sales records, with specified quantile values.
-In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example for the predicted demand for a single product sold on a single day in a given sales location.
+\ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e., the sales records, with specified quantile values.
+In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example, for the predicted demand for a single product sold on a single day in a given sales location.
 Each of these predicted negative binomial PDFs is then transformed to its CDF.
@@
 
-\fig \ref{fig:quant_profiles} illustrates five different collections of quantile profile plots (each collection comparing to 7 specified quantile values), for separate sets of exemplary PDF estimations and observed data combinations.
-The dashed horizontal lines indicates the fraction we expect, i.e.
-the specified quantile value, if the predictions are correct.
-For example, for the median, the line at 0.5 indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line.
-The observation of the number of samples, indicated with shaded circles, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations.
+\fig \ref{fig:invquant_example} illustrates five different collections of quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations.
+The dashed horizontal lines indicates the fraction we expect, i.e., the specified quantile value, if the predictions are correct.
+For example, for the median, the line at $q = 0.5$ indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line.
+The observation of the number of samples, indicated with different markers, that do in fact fall above and below a particular line, then allows the evaluation of the accuracy of PDF estimations.
 In case that the PDFs are not estimated correctly, the fractions will deviate from their expected values and the corresponding profile plot allows to judge whether for example the tails of the predicted distribution describing rare events are particularly problematic.
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/invquant_profiles}
-\caption{\label{fig:quant_profiles} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events.
-In the leftmost two columns of shaded circles illustrate the behavior if the estimate of the variance is biased, the center and center-right columns of shaded circles illustrate cases for which the mean estimation is biased and the far right column shows the expected behaviour if all predictions are correct.}
+\includegraphics[width=8cm]{../figures/invquant_example}
+\caption{\label{fig:invquant_example} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events.
+In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
 \end{center}
@@
 \ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population.
-However, calculating the fractions of the quantile profile plots globally, i.e.
-over all samples in the data set, might not reveal certain deficiencies for a subset, e.g.
-a specific store or product group in the example of retail demand forecasts.
-Therefore, we combine the approach described above with the method of profile plots, where the quantity on the x-axis of the quantile profile plot can be any variable of the data set at hand.
+However, calculating the fractions of the quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts.
+Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand.
 Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
@@
 \ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
-Second, by showing the dependency from the (arbitrary) variable on the x-axis, the quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
+Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
 specific categories, of that variable.
@@
 In the following, we describe how to use the approach outlined above in a practical setting.
-We use a public dataset obtained from a Kaggle online competition focusing in estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual SKUs for specific stores on specific days.
-For each demand forecast, Cyclic Boosting is used to predict the full probability density distribution of the expected demand at a granularity of (SKU, store, day) and use the methods described in Sec.
+We use a public dataset obtained from a Kaggle online competition focusing in estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual items for specific stores on specific days.
+For each demand forecast, Cyclic Boosting is used to predict the full probability density distribution of the expected demand at a granularity of (item, store, day) and use the methods described in sec.
 \ref{sec:pdfEvaluation} to evaluate the quality of the individual forecasts.
@@
 These metrics do not take the shape of the underlying probability distribution into account, but only compare the predicted mean to the observed number of sales.
-Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.15$.
+Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.09$.
 The mean of the target, i.e.
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/mean_prediction}
-\caption{\label{fig:mean_prediction} Time series of mean prediction and sales in test period summed over all products and stores.}
+\includegraphics[width=8cm]{../figures/ts_full}
+\caption{\label{fig:mean_prediction} Time series of mean predictions and sales in test period summed over all products and stores.}
 \end{center}
@@
 In this model focusing on the variance, we use the same set of features as for the mean model described above in sec.
-\ref{sec:example_mean}, except for dropping the two-dimensional combinations including individual event features.
-An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_586} in store \texttt{WI\_1} on 2016-05-22 is shown in \fig \ref{fig:pdf_example}.
+\ref{sec:example_mean}, except for dropping most of the two-dimensional combinations (only keeping \texttt{item\_id} - \texttt{store\_id}, \texttt{store\_id} - day of week, and \texttt{item\_id} - event type features) and adding the in-sample mean prediction as feature.
+An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_516} in store \texttt{TX\_3} on 2016-05-06 is shown in \fig \ref{fig:pdf_example}.
 
@@
 
+\fig \ref{fig:factor_plots} visualizes the fully explainable nature of the individual Cyclic Boosting predictions, both for the mean and the variance model, by showing the time series plots of the factor values constituting the corresponding predictions for item \texttt{FOODS\_3\_516} in store \texttt{TX\_3} for all days in the test period from beginning of February to end of April 2016.
+Hereby, factors of different individual features, as described above, are multiplied, according to \eqn \eqref{eqn:cb} for the mean predictions and \eqn \eqref{eqn:r} for the variance predictions, in order to represent the behavior of a feature group.
+For example, the \texttt{events} feature group corresponds to the combination of all the factors of the different event features and the \texttt{dayofweek} feature group includes the one-dimensional feature \texttt{dayofweek} and its two-dimensional combinations with \texttt{store\_id} and \texttt{item\_id}.
+
+\noindent
+The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values.
+The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl.
+The variance prediction for this day is mainly made up from two competing factors, the Super Bowl event driving higher dispersion parameter and the high mean prediction driving lower dispersion parameter.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4cm]{../figures/ts_item_16_store_6}
+\includegraphics[width=4cm]{../figures/factors_ts}
+\includegraphics[width=4cm]{../figures/factors_ts_width}
+\caption{\label{fig:factor_plots} Time series of mean predictions and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016.
+For both mean and variance predictions, the most granular constituting factors are aggregated, i.e.
+multiplied, for the sake of visualization.}
+\end{center}
+\end{figure}
+
 \subsection{Evaluation of PDF Predictions}
@@
 Using the same mean predictions for both negative binomial and Poisson model, the negative binomial PDF predictions are much closer to the uniform distribution, which we expect for optimal PDF predictions, than the Poisson PDF predictions, showing the effectiveness of our variance estimation.
-The only significant deviation of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$.
-These cases are due to slow moving articles with a predicted mean lower than $1.0$, i.e.
-less than one item sold per store per day on average.
-These items have been removed in the lower part of  \fig \ref{fig:cdf_demand} and the resulting cumulative distribution is flat as expected.
+
+\noindent
+While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$ as well as in the first bins of CDF values close to $0$.
+These cases are mainly due to slow moving articles with less than one item sold per store per day on average.
 In practice, predicting the demand for these kind of items is very challenging as they are prone to over-dispersion and large fluctuations due to the discrete nature of the sales event.
+In the lower part of \fig \ref{fig:cdf_demand}, the effects of these slow-sellers are reduced by excluding all samples with mean predictions lower than $1.0$.
+The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.
 
@@
 
-Using the profile plots introduced earlier, the prediction quality of the full predicted probability distribution can be assessed in more detail.
-\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 10 columns correspond to mean predictions in the intervals $[0.0, 5.0]$, $(5.0, 10.0]$, $(10.0, 15.0]$, $(15.0, 20.0]$, $(20.0, 30.0]$, $(30.0, 40.0]$, $(40.0, 50.0]$, $(50.0, 60.0]$, $(60.0, 80.0]$, $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
-The considered quantile values are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
-We can see that for mean predictions higher than $80$, our negative binomial model deviates significantly from the expected behaviour, with the shape of the deviations pointing towards a general underprediction (see  \fig \ref{fig:quant_profiles}).
+Using the profile plots introduced earlier, the prediction quality of the full predicted PDF can be assessed in more detail.
+In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the considered quantile values are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
+Again, we compare the outcome of our negative binomial model to a simpler Poisson assumption.
+Aggregating over all stores, items, and sales dates (independently for each day of the week), we can see that for the negative binomial model each quantile is predicted relatively well, except for Sundays showing a tendency for overprediction of the mean (see \fig \ref{fig:invquant_example}).
+The Poisson model, on the other hand, shows a significant pattern of too narrow PDF predictions across all days of the week.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4cm]{../figures/invquant_dayofweek_nbinom}
+\includegraphics[width=4cm]{../figures/invquant_dayofweek_poisson}
+\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
+\end{center}
+\end{figure}
+
+\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
+The used quantile values are the same as in \fig \ref{fig:invquant_dayofweek}.
+While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model.
+For mean predictions around $30$, the negative binomial model deviates significantly from the expected behaviour, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above).
+And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter.
+The deviations at very low predicted mean values reflect the complications of slow-sellers described in the context of CDF histograms above.
+
+\noindent
 In a real-live situation with a live supply chain project for a customer, a more detailed investigation of the root-causes would then start.
@@
 It is therefore paramount both to predict the full probability distribution instead of just a point estimate, such as the mean, as well as verifying a number of quantiles for each prediction to assess the quality of the prediction thoroughly.
-In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the used quantile values are the same as in \fig \ref{fig:invquant_mean}.
-Aggregating over all stores, items, and sales dates (independently for each day of the week), we can see that each quantile is predicted relatively well.
 
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/invquant_yhat_mean}
-\caption{\label{fig:invquant_mean} Quantile profile plot for mean predictions on the x-axis aggregated over all product-location-day combinations in the test period.}
+\includegraphics[width=4cm]{../figures/invquant_yhat_mean_nbinom}
+\includegraphics[width=4cm]{../figures/invquant_yhat_mean_poisson}
+\caption{\label{fig:invquant_mean} Quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
 \end{center}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=8cm]{../figures/invquant_dayofweek}
-\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the x-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period.}
-\end{center}
-\end{figure}
-
-The quantitative results for the CDF accuracy of our PDF predictions for the different metrics described in sec.
+The quantitative results for the CDF accuracy of our PDF predictions for different metrics (Wasserstein distance as well as Kullback-Leibler and Jensen-Shannon divergence with base-2 logarithm) described in sec.
 \ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf_acc}.
@@
 \hline
-\textbf{EMD} & 0.965 & 0.851 \\
-\textbf{KL\_2} & 0.995 & 0.933 \\
-\textbf{KL\_e} & 0.997 & 0.954 \\
-\textbf{JSD\_2} & 0.999 & 0.982 \\
-\textbf{JSD\_e} & 0.999 & 0.987
+\textbf{EMD} & 0.967 & 0.850 \\
+\textbf{KL\_2} & 0.994 & 0.933 \\
+\textbf{JSD\_2} & 0.998 & 0.981
 \end{tabular}
@@
 Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general.
-Instead, both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions.
+Instead, novel techniques using both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions.
 Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of the probability integral transformation and the expected uniform distribution.

### 2020-08-27 code/walmart_demand.py

@@ -34,7 +34,7 @@ def plot_CB(filename, plobs, binner):
         )
 
 
-def plot_factors(est, X, samples):
+def calculate_factors(est, X):
     factors = {}
     for feature in est.ests[-1][1].features:
         ft = "" if feature.feature_type is None else "_{}".format(
@@ -42,19 +42,95 @@ def plot_factors(est, X, samples):
 
         factors[
             "_".join(feature.feature_group) + ft
-            ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False)) - 1.
+            ] = np.exp(est.ests[-1][1]._pred_feature(X, feature, is_fit=False))
+    factors['date'] = X['date']
+    factors['yhat_mean'] = X['yhat_mean']
+    factors['y'] = X['y']
     df_factors = pd.DataFrame(factors)
 
-    ax = df_factors.loc[samples].plot(kind='barh', title="factors", legend=True, fontsize=15)
-    fig = ax.get_figure()
-    ax.axvline(0, color="gray")
-    ax.set_yticks([])
+    df_factors['item_store'] = df_factors['item_id'] * df_factors['store_id'] * df_factors['item_id_store_id']
+    df_factors['price'] = df_factors['promo'] * df_factors['price_ratio'] * df_factors['item_id_promo'] * df_factors['item_id_price_ratio']
+    df_factors['dayofweek'] = df_factors['dayofweek'] * df_factors['store_id_dayofweek'] * df_factors['item_id_dayofweek']
+    df_factors['weekofmonth'] = df_factors['weekofmonth'] * df_factors['store_id_weekofmonth']
+    df_factors['snap'] = df_factors['snap'] * df_factors['snap_dayofweek'] * df_factors['snap_item_id']
+    df_factors['seasonality'] = df_factors['td'] * df_factors['dayofyear'] * df_factors['month']
[…]

### 2020-08-30 LaTex/paper.tex
cosmetics, larger axis labels and legends, added authors

@@
 \author[1]{Felix Wick \thanks{felix.wick@blueyonder.com}}
-\author[2]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
-\author[1]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
+\author[3]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
 \author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
+\author[1]{Moritz Wolf \thanks{martin.hahn@blueyonder.com}}
+\author[2]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
+\author[1]{Michael Feindt \thanks{michael.feindt@blueyonder.com}}
 \affil[1]{\small Blue Yonder GmbH (Karlsruhe, Germany)}
-\affil[2]{\small IUBH Internationale Hochschule (Erfurt, Germany)}
+\affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
+\affil[3]{\small IUBH Internationale Hochschule (Erfurt, Germany)}
 
@@
 Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
-Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and in case the demand exceeds the current stock level the data become censored.
-Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and locations.
+Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and, in case the demand exceeds the current stock level, the data become censored.
+Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and sales locations.
 While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors.
-The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products as well as pursuing an active pricing strategy.
+The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products, as well as pursuing an active pricing strategy.
 This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required.
@@
 
-Thanks to the availability of Big Data today, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed.
+Because of the access to  Big Data, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed.
 Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data.
@@
 Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
-Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoilt items need to be disposed of at additional cost and potentially even environmental impact.
+Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact.
 This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
@@
 Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled.
-The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90\%.
-So, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.
+The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent.
+Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.
 
@@
 Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages.
-First, demand forecasts in form of full PDFs can be used to simulate the performance on an individual level, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate.
+First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate.
 From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
 It also allows to share long-term demand predictions with other business units or external vendors and wholesalers to ease their planning for the production and supply chain processes upstream of the retailer.
-From the perspective of the industrial practice of a supplier of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other.
+From the perspective of the industrial practice of a vendor of supply chain methods and tools, modeling the demand separately from deriving the subsequent orders has the additional benefit that multiple retail chains can benefit from any improvement in the model description, even if the specific retailers are unrelated to each other.
 Additionally, a purely data-driven approach going from the observed data directly to the operational decision (such as the order quantity) does not allow to analyze the data-generating process, i.e., the mechanism behind the stochastic behavior of the customer demand.
@@
 Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution.
-The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see e.g.
-\cite{shapiro2014} for an overview, or as a time series (see e.g.
-\cite{alwan2016}), for example by means of autoregressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}.
-However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements, or weather influence, and others are highly correlated with the demand.
+The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see, e.g., \cite{shapiro2014} for an overview, or as a time series (see, e.g., \cite{alwan2016}), for example by means of auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}.
+However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements,  weather influence, and others are highly correlated with the demand.
 Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity.
 As described in sec.
-\ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches and therefore are the method of choice in most use cases today (see e.g.
-\cite{tsforecasting,ZHANG199835,rnn,transformer}).
+\ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches, and therefore are the method of choice in most use cases today (see, e.g., \cite{tsforecasting,ZHANG199835,rnn,transformer}).
 
@@
 Instead, the choice of the demand distribution should be motivated by theoretic considerations.
-The discrete demand is typically modelled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}.
+The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}.
 This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
@@
 
-This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity needed by the retailer.
+This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity required by the retailer.
 Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
@@
 \ref{sec:LitRev}, machine learning algorithms are ideally suited for the task of demand forecasting and in the following, we will use the Cyclic Boosting algorithm to benefit in particular from explainable decisions rather than black-box approaches.
-Furthermore, the regularization approach used during training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm.
-This means we use two subsequent Cyclic Boosting models, in order to estimate the parameters of each individual PDF that we need to forecast, where the first model is used to estimate the mean and the second to estimate the variance.
-Hereby, the features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model.
+Furthermore, the regularization approach used during the training of the Cyclic Boosting algorithm allows a dedicated treatment of the underlying NBD model, which is another major benefit compared to a standard ``off-the-shelf'' machine learning algorithm.
+This means we use two subsequent Cyclic Boosting models in order to estimate the parameters of each individual PDF that we need to forecast.
+The first model is used to estimate the mean and the second to estimate the variance.
+The features may or may not differ between the mean and variance estimation models, and it can be beneficial to include the corresponding mean predictions as feature in the variance model.
 The assigned mean and variance predictions can then be used to generate individual PDFs using the parameterization of the NBD for each sample.
@@
 
-Cyclic Boosting is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters.
+Cyclic Boosting \cite{Wick2019} is a type of generalized additive model using a cyclic coordinate descent optimization and featuring a boosting-like update of parameters.
 Major benefits of Cyclic Boosting are its accuracy, performance, even at large scale, and providing fully explainable predictions, which are of vital importance in practical applications.
 
-The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behaviour of the feature.
+The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behavior of the feature.
 The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
-Single data records, for example, the sales of a specific product-location-date combination along with all relevant features, are indexed by $i$.
+Single data records, for example, the sales corresponding to a specific product-location-date combination along with all relevant features, are indexed by $i$.
 The individual predictions $\hat{y_i}$  can then be calculated as:
@@
 \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed.
-So, in case of a correctly calibrated PDF prediction, the distribution of the corresponding actually observed CDF values of the individual PDF predictions is expected to be uniform, {\em regardless} of the shape of the predicted distribution, and any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.
+We therefore expect that distribution of the corresponding actually observed CDF values of the individual PDF predictions is uniform if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
+Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.
 
+\noindent
 The CDF of a PDF $f(x)$ is defined as:
@@
 In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF.
-For example, if the sales of zero items accounts for 75\% of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
+For example, if the sales of zero items accounts for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
 Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g.
-for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values should again be uniform as in the case of a continuous PDF.
+for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform as in the case of a continuous PDF.
 
@@
 \ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$.
-In other words, in case of the median, 50\% of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50\% above.
+In other words, in case of the median, 50 percent of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50 percent above.
 
@@
 In the following, we call this method {\em quantile profile plot}.
+Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
 
@@
 Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand.
-Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
 
@@
 The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl.
-The variance prediction for this day is mainly made up from two competing factors, the Super Bowl event driving higher dispersion parameter and the high mean prediction driving lower dispersion parameter.
+The variance prediction for this day is driven by two competing factors:  the Super Bowl event itself drives  the  higher dispersion parameter higher, and the high mean prediction drives the lower dispersion parameter lower.
 
@@
 \fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
-The used quantile values are the same as in \fig \ref{fig:invquant_dayofweek}.
+The  quantile values are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}.
 While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model.
-For mean predictions around $30$, the negative binomial model deviates significantly from the expected behaviour, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above).
+For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above).
 And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter.
@@
 
-Evaluating these PDF predictions becomes the critical question.
+Once the full probability density functions are forecast, their evaluation poses significant challenges:
 Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions.
@@
 
-In many cases, scatter plots are used to study the behaviour of two distributions or sets of data points visually.
+In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually.
 However, even for moderate amount of data, this approach becomes quickly difficult.
@@
 This means that the bin-borders on the $X$ axis are used as constraints on the variable $Y$ and with these conditions applied, for example the sample mean of the selected $y$ values as well as the standard deviation are calculated.
-These location and dispersion metrics in each bin of $X$ are used to illustrate the behaviour of the variable $Y$ as the values of the variable $X$ change from bin to bin.
+These location and dispersion metrics in each bin of $X$ are used to illustrate the behavior of the variable $Y$ as the values of the variable $X$ change from bin to bin.
 The resulting profile histogram is shown in \fig \ref{fig:profile}.

### 2020-08-30 code/walmart_demand.py

@@ -105,8 +105,8 @@ def plot_factors_ts(df, filename):
 #    df['y'].plot(label="y")
 #    df['yhat_mean'].plot(label="yhat_mean")
 
-    plt.legend(loc=1)
-    plt.ylabel("factors")
+    plt.legend(loc=1, fontsize=13)
+    plt.ylabel("factors", fontsize=13)
     plt.tight_layout()
     plt.savefig(filename)
 
@@ -127,8 +127,8 @@ def plot_factors_ts_width(df, filename):
 #    df['check'].plot(label="check")
 #    df['c'].plot(label="c")
 
-    plt.legend(loc=1)
-    plt.ylabel("factors")
+    plt.legend(loc=1, fontsize=13)
+    plt.ylabel("factors", fontsize=13)
     plt.tight_layout()
     plt.savefig(filename)
 
@@ -137,7 +137,7 @@ def plot_pdf(n, p):
     plt.figure()
     x = range(30)
     plt.plot(x, nbinom.pmf(x, n, p))
-    plt.xlabel("sales")
+    plt.xlabel("sales", fontsize=15)
     plt.tight_layout()
     plt.savefig('plots/pdf.pdf')
 
@@ -146,7 +146,7 @@ def plot_cdf(n, p):
     plt.figure()
     x = range(30)
     plt.plot(x, nbinom.cdf(x, n, p))
-    plt.xlabel("sales")
+    plt.xlabel("sales", fontsize=15)
     plt.tight_layout()
     plt.savefig('plots/cdf.pdf')
 
@@ -155,9 +155,9 @@ def plot_cdf_truth(cdf_truth, suffix):
     plt.figure()
     plt.hist(cdf_truth, bins=30)
     if suffix in ['uniform', 'over', 'under', 'broad', 'narrow']:
-        plt.title(suffix)
-    plt.xlabel("CDF values")
-    plt.ylabel("count")
+        plt.title(suffix, fontsize=20)
+    plt.xlabel("CDF values", fontsize=15)
+    plt.ylabel("count", fontsize=15)
     plt.tight_layout()
     plt.savefig('plots/cdf_truth_' + suffix + '.pdf')
 
@@ -195,8 +195,8 @@ def plot_invquants(X, variable, suffix, continuous=False):
     plt.errorbar(x=bin_centers, y=means90, yerr=stdmean90, linestyle='none', marker='^', c='y')
[…]

### 2020-09-06 LaTex/paper.tex
forgot accuracy formula, some renaming

@@
 Many statistical and most machine learning methods do not provide a full probability density distribution as result.
-Instead, these methods typically predict a single numerical quantity that is then compared to the observed concrete realization of the random variable using metrics such as the mean squared error (MSE), mean absolute deviation (MAD) or others.
-In the setting of a retailer, the observed quantity is the sales of individual products (denoted by $y$) and most machine learning approaches would then predict a single number $\hat{y}$ as a direct estimate of the sales.
+Instead, these methods typically predict a single numerical quantity (denoted by $\hat{y}$) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the mean squared error (MSE), mean absolute deviation (MAD) or others.
+In the setting of a retailer, the observed quantity is the sales of individual products and most machine learning approaches would then predict a single number as a direct estimate of the sales.
 However, reducing the prediction to a single number does not allow to account for the uncertainty of the prediction or the dynamics of the system.
@@
 
-For an estimation of the determining parameters of an assumed functional form for the PDF, assessing the correctness of the PDF model output refers to the evaluation of the accuracy of the prediction of the different determining parameters.
+For an estimation of the determining parameters of an assumed functional form for the PDF, assessing the correctness of the PDF model output refers to the evaluation of the accuracy of the prediction of the different parameters.
 In the case of the negative binomial distribution used in this work, we have to verify that mean and variance are determined accurately, as well as checking that the choice of the underlying model can describe the observed data.
@@
 In the following, we will show how different visualizations of the observed cumulative distribution function (CDF) values can be used to evaluate the quality of the predicted PDFs.
-Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if the PDF is obtained empirically.
+Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if it is obtained empirically.
 
@@
 \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed.
-We therefore expect that distribution of the corresponding actually observed CDF values of the individual PDF predictions is uniform if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
+We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
 Any deviation can be interpreted as a hint that the predicted PDF is not fully correct \cite{diebold1998vevaluating}.
@@
 If the CDF is continuous and strictly increasing, then the inverse of the CDF, $F^{-1}(y)$, exists and is a unique real-valued number $x$ for each $y \in [0,1]$, so that we can write $F(x) = y$.
-The latter defines the inverse quantile function, because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
+The inverse of the CDF is also called the quantile function, because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
 
@@
 
-In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the inverse quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
+In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
 \cite[p.
@@
 Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g.
-for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform as in the case of a continuous PDF.
+for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.
 
-A histogram of the observed quantiles (or CDF values) for each individual PDF prediction is therefore expected to be uniformly distributed in $[0,1]$, if the predicted probability distribution $f(x)$ is correctly calibrated.
-This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed quantiles for five different cases.
+A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted probability distribution $f(x)$ is correctly calibrated.
+This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for five different cases.
 If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution.
@@
 \includegraphics[width=2.5cm]{../figures/cdf_truth_under}
-\caption{\label{fig:cdf_histos} Quantile distribution for different cases of estimating the full probability distribution compared to the expected distribution:
+\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
 correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
@@
 
-\subsubsection{Quantile Profile Plot}
+\noindent
+It should be noted that the method requires a sufficient sample size, as a too low number of observations of course leads to a discretization bias.
+
+\subsubsection{Inverse Quantile Profile Plot}
 
 We now refine the method from sec.
-\ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the quantiles of the observed events, i.e., the sales records, with specified quantile values.
+\ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the CDF values of the observed events, i.e., the sales records, with specified quantiles.
 In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example, for the predicted demand for a single product sold on a single day in a given sales location.
@@
 Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample.
-For example, if we wanted to check that the median of the distribution, corresponding to the $0.5$ quantile, is predicted correctly by the machine learning model, we would compare the quantile value $0.5$ to the ratio of quantile values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec.
+For example, if we wanted to check that the median of the distribution, corresponding to the quantile $0.5$, is predicted correctly by the machine learning model, we would compare the value $0.5$ to the ratio of CDF values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec.
 \ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$.
@@
 However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say $q = 0.95$ and $q = 0.99$.
-In the following, we call this method {\em quantile profile plot}.
+In the following, we call this method {\em inverse quantile profile plot}.
 Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
 
-\fig \ref{fig:invquant_example} illustrates five different collections of quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations.
+\fig \ref{fig:invquant_example} illustrates five different collections of inverse quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations.
 The dashed horizontal lines indicates the fraction we expect, i.e., the specified quantile value, if the predictions are correct.
@@
 \includegraphics[width=8cm]{../figures/invquant_example}
-\caption{\label{fig:invquant_example} Profile plots of quantiles comparing all predicted PDFs to the corresponding observed events.
+\caption{\label{fig:invquant_example} Inverse quantile profile plots comparing all predicted PDFs to the corresponding observed events.
 In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
@@
 In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec.
-\ref{sec:cdf_histo}), quantile profile plots do not work on individual PDF predictions but require a statistical population.
-However, calculating the fractions of the quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts.
-Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the quantile profile plot can be any variable of the data set at hand.
+\ref{sec:cdf_histo}), inverse quantile profile plots do not work on individual PDF predictions but require a statistical population.
+However, calculating the fractions of the inverse quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts.
+Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand.
 
 In summary, this approach has two major benefits compared to the method discussed in sec.
-\ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
-Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
+\ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the inverse quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
+Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the inverse quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
 specific categories, of that variable.
@@
 However, in order to also quantify the quality of the PDF predictions, a metric is required that assesses the difference between the distributions.
-Therefore, we want to compare the CDF histogram of the predicted PDFs with the expected  uniform distribution and define a metric in the range between 0 and 1, such that the metric takes the value of 1 when both distributions agree perfectly.
-Several approaches that measure the difference between two probability distributions are suggested in the literature such as  the first Wasserstein distance \cite{olkin1982}, the Kullback-Leibler divergence \cite{kullback1951}, and the Jensen-Shannon divergence  \cite{dagan1997}.
+Therefore, we want to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
+Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, or the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius.
 
-The first Wasserstein distance, also known as earth mover distance (EMD), represents a distance  defined between two probability distributions on a given metric space, and for our purposes here can defined by:
+Compared to the other mentioned methods, the first Wasserstein distance, representing the symmetric distance between two probability distributions on a given metric space, is more sensitive to smaller deviations, because it exhibits a linear behavior around zero (reflecting perfect agreement).
+Therefore, we focus on the first Wasserstein distance as the measure of difference in the following.
+For our purposes here, it can be defined by:
 \begin{equation}
-\text{EMD}(P, Q) = 2 \cdot \frac{\sum_{k=1}^N |F_P(x_k) - F_Q(x_k)|}{N},
+\text{EMD}(P, Q) = \frac{\sum_{k=1}^N |F_P(x_k) - F_Q(x_k)|}{N},
 \end{equation}
 where $F_P(X)$ and $F_Q(X)$ are the CDFs of the two PDFs $P(X)$ and $Q(X)$, respectively, and $x_k$ denotes the average value of $X$ in bin $k$, with $X$ being divided in $N$ bins.
-Compared to the other metrics given below, the first Wasserstein or earth mover distance has the additional benefit, that it takes a specific metric space into account, meaning that it depends on the distance of potential deviations.
 
-The Kullback-Leibler divergence is a measure of how one probability distribution diverges from a second expected probability distribution.
-For PDFs $P(X)$ and $Q(X)$, again with $X$ being divided in $N$ bins $k$, the Kullback-Leibler divergence from Q to P is defined as:
+Since $0.5$ represents the maximum value of the first Wasserstein distance (its minimum being zero), we define an accuracy measure for our PDF predictions in the range $[0, 1]$ by:
 \begin{equation}
-D_{\text{KL}}(P \parallel Q) = \sum _{k=1}^N P(x_k) \log \left({\frac{P(x_k)}{Q(x_k)}}\right),
+\text{accuracy} = 1 - 2 \cdot \text{EMD}
 \end{equation}
-where the logarithm can be either base-2 or the natural logarithm.
-
-The Jensen-Shannon divergence (JSD) can be seen as a symmetrized and smoothed version of the Kullback-Leibler divergence:
-\begin{equation}
-D_{\text{JSD}}(P \parallel Q) = 0.5  \cdot (D_{\text{KL}}(P \parallel M) + D_{\text{KL}}(P \parallel M)),
-\end{equation}
-where $M = 0.5  \cdot (P + Q)$.
 
@@
 \ref{sec:example_mean}, except for dropping most of the two-dimensional combinations (only keeping \texttt{item\_id} - \texttt{store\_id}, \texttt{store\_id} - day of week, and \texttt{item\_id} - event type features) and adding the in-sample mean prediction as feature.
-An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_516} in store \texttt{TX\_3} on 2016-05-06 is shown in \fig \ref{fig:pdf_example}.
+An example for the resulting PDF and CDF predictions for item \texttt{FOODS\_3\_516} in store \texttt{TX\_3} on 2016-05-06, together with the corresponding observed sales value, is shown in \fig \ref{fig:pdf_example}.
 
@@
 \includegraphics[width=4cm]{../figures/cdf}
-\caption{\label{fig:pdf_example} Predicted PDF (left) and CDF (right) for a specific product-location-date combination.}
+\caption{\label{fig:pdf_example} Predicted PDF (left) and CDF (right) for a specific product-location-date combination.
+The dashed vertical line in the CDF plot represents the actually observed sales, corresponding to a CDF value just under $0.3$.}
 \end{center}
@@
 
-Using the profile plots introduced earlier, the prediction quality of the full predicted PDF can be assessed in more detail.
-In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the considered quantile values are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
+Using the inverse quantile profile plots introduced earlier, the prediction quality of the full predicted PDF can be assessed in more detail.
+In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the considered quantiles are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
 Again, we compare the outcome of our negative binomial model to a simpler Poisson assumption.
@@
 \includegraphics[width=4cm]{../figures/invquant_dayofweek_poisson}
-\caption{\label{fig:invquant_dayofweek} Quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
+\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
 \end{center}
@@
 
-\fig \ref{fig:invquant_mean} shows the quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
-The  quantile values are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}.
+\fig \ref{fig:invquant_mean} shows the inverse quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
+The considered quantiles are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}.
 While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model.
@@
 \includegraphics[width=4cm]{../figures/invquant_yhat_mean_poisson}
-\caption{\label{fig:invquant_mean} Quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
+\caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
 \end{center}
@@
 
-The quantitative results for the CDF accuracy of our PDF predictions for different metrics (Wasserstein distance as well as Kullback-Leibler and Jensen-Shannon divergence with base-2 logarithm) described in sec.
-\ref{sec:cdf_acc}, calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf_acc}.
+The quantitative results for the CDF accuracy of our PDF predictions using the first Wasserstein distance as metric (as described in sec.
+\ref{sec:cdf_acc}) and calculated over all product-location-day combinations in the test period, can be found in \tab \ref{tab:cdf_acc}.
 As benchmark, we again compare against a Poisson model assumption using the same mean predictions.
@@
 \begin{center}
-\caption{CDF accuracy for negative binomial and Poisson PDF predictions for different metrics calculated over all product-location-day combinations in the test period.}
+\caption{Accuracy for negative binomial and Poisson PDF predictions, using the first Wasserstein distance as metric, calculated over all product-location-day combinations in the test period.}
 \label{tab:cdf_acc}
@@
 \hline
-\textbf{EMD} & 0.967 & 0.850 \\
-\textbf{KL\_2} & 0.994 & 0.933 \\
-\textbf{JSD\_2} & 0.998 & 0.981
+\textbf{EMD accuracy} & 0.967 & 0.850
 \end{tabular}
@@
 
-Once the full probability density functions are forecast, their evaluation poses significant challenges:
-Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions.
+Once the full probability density functions are forecasted, their evaluation poses significant challenges: Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions.
 Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general.
-Instead, novel techniques using both the histogram of the probability integral transformation as well as quantile profile histograms allow a detailed investigation into the behavior of the predictions.
-Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of the probability integral transformation and the expected uniform distribution.
+Instead, novel techniques exploiting the probability integral transform, namely the histogram of observed CDF values of the predicted individual PDFs as well as inverse quantile profile plots, allow a detailed investigation into the behavior of the predictions.
+Finally, a quantitative assessment resulting in a single number can be obtained using metrics such as the earth mover distance in a comparison between the histogram of observed CDF values and the expected uniform distribution.
 

### 2020-09-06 code/walmart_demand.py

@@ -142,10 +142,11 @@ def plot_pdf(n, p):
     plt.savefig('plots/pdf.pdf')
 
 
-def plot_cdf(n, p):
+def plot_cdf(n, p, y):
     plt.figure()
     x = range(30)
     plt.plot(x, nbinom.cdf(x, n, p))
+    plt.vlines(y, ymin=0, ymax=1, linestyles ="dashed")
     plt.xlabel("sales", fontsize=15)
     plt.tight_layout()
     plt.savefig('plots/cdf.pdf')
@@ -275,48 +276,49 @@ def eval_results(yhat_mean, y):
     print('mean(y): {}'.format(mean_y))
 
 
-def wasserstein_2(p, q, n_bins):
+def wasserstein(p, q, n_bins):
     cdf_p = np.cumsum(p)
     cdf_q = np.cumsum(q)
+    # scale by 1/0.5 (0.5 is maximum)
     wasser_distance = 2.0 * np.sum(np.abs(cdf_p - cdf_q)) / len(cdf_p)
     return wasser_distance * n_bins / (n_bins - 1.0)
 
 
-def kullback_2(p, q):
-    mask = p > 0  # limit x -> 0 of x log(x) is zero
-    return np.float(np.sum(p[mask] * (np.log2(p[mask]) - np.log2(q[mask]))))
+# def kullback_2(p, q):
+#     mask = p > 0  # limit x -> 0 of x log(x) is zero
+#     return np.float(np.sum(p[mask] * (np.log2(p[mask]) - np.log2(q[mask]))))
+#
+#
+# def jensen_2(p, q):
+#     m = (p + q) / 2.0
+#     return (kullback_2(p, m) + kullback_2(q, m)) / 2.0
+#
+#
+# def kullback_e(p, q):
+#     mask = p > 0  # limit x -> 0 of x log(x) is zero
+#     return np.float(np.sum(p[mask] * (np.log(p[mask]) - np.log(q[mask]))))
+#
+#
+# def jensen_e(p, q):
+#     m = (p + q) / 2.0
+#     return (kullback_e(p, m) + kullback_e(q, m)) / 2.0
 
 
-def jensen_2(p, q):
-    m = (p + q) / 2.0
-    return (kullback_2(p, m) + kullback_2(q, m)) / 2.0
-
-
-def kullback_e(p, q):
-    mask = p > 0  # limit x -> 0 of x log(x) is zero
-    return np.float(np.sum(p[mask] * (np.log(p[mask]) - np.log(q[mask]))))
-
-
[…]

### 2020-09-08 LaTex/paper.tex
added error band in plot, some minor review changes

@@
 In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer.
-The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot-sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
+The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
 Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
 Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact.
-This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh diary products, or certain meat products such as ground meat.
+This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh dairy products, or certain meat products such as ground meat.
 These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item.
@@
 \includegraphics[width=8cm]{../figures/PdfCdf}
-\caption{\label{fig:PdfCdf} Probability distribution function and cumulative distribution function of a normal distribution.}
+\caption{\label{fig:PdfCdf} Probability distribution function $f(x)$ and cumulative distribution function of a normal distribution $F(x)$.}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=2.5cm]{../figures/cdf_truth_uniform}
-\includegraphics[width=2.5cm]{../figures/cdf_truth_broad}
-\includegraphics[width=2.5cm]{../figures/cdf_truth_narrow}
-\includegraphics[width=2.5cm]{../figures/cdf_truth_over}
-\includegraphics[width=2.5cm]{../figures/cdf_truth_under}
+\includegraphics[scale=0.25]{../figures/cdf_truth_uniform}
+\includegraphics[scale=0.25]{../figures/cdf_truth_broad}
+\includegraphics[scale=0.25]{../figures/cdf_truth_narrow}
+\includegraphics[scale=0.25]{../figures/cdf_truth_over}
+\includegraphics[scale=0.25]{../figures/cdf_truth_under}
 \caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
@@
 \includegraphics[width=8cm]{../figures/invquant_example}
-\caption{\label{fig:invquant_example} Inverse quantile profile plots comparing all predicted PDFs to the corresponding observed events.
+\caption{\label{fig:invquant_example} Five different inverse quantile profile plots (separated into columns by dotted vertical lines), each comparing all its predicted PDFs to the corresponding observed events.
 In the leftmost two columns, the outcome is illustrated if the estimate of the variance is biased, the center and center-right columns illustrate cases for which the mean estimation is biased, and the far right column shows the expected behavior if all predictions are correct.}
@@
 As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event.
-As a first step, we use Cyclic Boosting to predict the mean of the distribution.
+As a first step, we use Cyclic Boosting (as described in sec.
+\ref{sec:CB}) to predict the mean of the distribution.
 
@@
 
-The second model, also based on Cyclic Boosting, is used to estimate the variance of the negative binomial distribution.
+The second model, based on Cyclic Boosting in its negative binomial width mode described in sec.
+\ref{sec:cb_width}, is used to estimate the variance of the negative binomial distribution.
 The data are split into training and test set as above and in addition the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}.
@@
 \includegraphics[width=4cm]{../figures/factors_ts_width}
-\caption{\label{fig:factor_plots} Time series of mean predictions and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016.
+\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016.
 For both mean and variance predictions, the most granular constituting factors are aggregated, i.e.
@@
 In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually.
-However, even for moderate amount of data, this approach becomes quickly difficult.
+However, even for moderate amount of data, this approach quickly becomes difficult.
 To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$.

### 2020-09-08 code/walmart_demand.py

@@ -220,6 +220,7 @@ def plot_invquants_examples():
     plt.errorbar(x=bins, y=means90, linestyle='none', marker='^', c='y')
     plt.errorbar(x=bins, y=means97, linestyle='none', marker='s', c='m')
     plt.hlines([0.1, 0.3, 0.5, 0.7, 0.9, 0.97], 'broad', 'uniform', color=['b', 'g', 'k', 'r', 'y', 'm'], linestyle='dashed')
+    plt.vlines([0.5, 1.5, 2.5, 3.5], 0, 1, linestyle='dotted')
     plt.ylabel("quantile", fontsize=15)
     plt.xticks(fontsize=15)
     plt.tight_layout()
@@ -231,6 +232,7 @@ def plot_timeseries(df, suffix, title=''):
     df.index = df['date']
     df['y'].plot(style='r', label="sales")
     df['yhat_mean'].plot(style='b', label="prediction")
+    plt.fill_between(df.index, df['yhat_mean'] - np.sqrt(df['yhat_var']), df['yhat_mean'] + np.sqrt(df['yhat_var']), alpha=0.2)
     plt.legend(fontsize=15)
 #    plt.title(title)
     plt.ylabel("sum", fontsize=15)
@@ -239,24 +241,24 @@ def plot_timeseries(df, suffix, title=''):
 
 
 def plotting(df, suffix=''):
-    df = df[['y', 'yhat_mean', 'item_id', 'store_id', 'date']]
+    df = df[['y', 'yhat_mean', 'yhat_var', 'item_id', 'store_id', 'date']]
 
-    ts_data = df.groupby(['date'])[['y', 'yhat_mean']].sum().reset_index()
+    ts_data = df.groupby(['date'])[['y', 'yhat_mean', 'yhat_var']].sum().reset_index()
     plot_timeseries(ts_data, 'full' + suffix, 'all')
 
     predictions_grouped = df.groupby(['store_id'])
     for name, group in predictions_grouped:
-        ts_data = group.groupby(['date'])['y', 'yhat_mean'].sum().reset_index()
+        ts_data = group.groupby(['date'])['y', 'yhat_mean', 'yhat_var'].sum().reset_index()
         plot_timeseries(ts_data, 'store_' + str(name) + suffix)
 
     predictions_grouped = df.groupby(['item_id'])
     for name, group in predictions_grouped:
-        ts_data = group.groupby(['date'])['y', 'yhat_mean'].sum().reset_index()
[…]

### 2020-09-10 LaTex/paper.tex
plot cosmetics

@@
 \author[1]{Michael Feindt \thanks{michael.feindt@blueyonder.com}}
-\affil[1]{\small Blue Yonder GmbH (Karlsruhe, Germany)}
+\affil[1]{\small Blue Yonder GmbH (Ohiostra{\ss}e 8, 76149 Karlsruhe, Germany)}
 \affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
@@
 \caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
-correct prediction (``uniform''), variance overestimated (``broad''), variance underestimated (``narrow''), mean overestimated (``over''), mean underestimated (``under'')}
+correct prediction (a: ``uniform''), variance overestimated (b: ``broad''), variance underestimated (c: ``narrow''), mean overestimated (d: ``over''), mean underestimated (e: ``under'')}
 \end{center}
@@
 \includegraphics[width=4cm]{../figures/cdf}
-\caption{\label{fig:pdf_example} Predicted PDF (left) and CDF (right) for a specific product-location-date combination.
+\caption{\label{fig:pdf_example} Predicted PDF (a) and CDF (b) for a specific product-location-date combination.
 The dashed vertical line in the CDF plot represents the actually observed sales, corresponding to a CDF value just under $0.3$.}
@@
 \includegraphics[width=4cm]{../figures/factors_ts_width}
-\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (upper left plot), constituting factors of the mean predictions (upper right plot), and constituting factors of the variance predictions (lower plot), for a specific product-location combination from beginning of February to end of April 2016.
+\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (a), constituting factors of the mean predictions (b), and constituting factors of the variance predictions (c), for a specific product-location combination from beginning of February to end of April 2016.
 For both mean and variance predictions, the most granular constituting factors are aggregated, i.e.
@@
 \includegraphics[width=4cm]{../figures/cdf_truth_poisson_larger1}
-\caption{\label{fig:cdf_demand} Histograms of ex post observed target CDF values of the corresponding individual PDF predictions (to be compared to a uniform distribution) for our negative binomial model (left) and a simpler Poisson model for comparison (right), using all product-location-day combinations in the test period (upper two plots).
-In order to show the effect of slow-sellers, the lower two plots exclude all samples with mean predictions lower than $1.0$.}
+\caption{\label{fig:cdf_demand} Histograms of ex post observed target CDF values of the corresponding individual PDF predictions (to be compared to a uniform distribution) for our negative binomial model (a) and a simpler Poisson model for comparison (b), using all product-location-day combinations in the test period.
+In order to show the effect of slow-sellers, both negative binomial (c) and Poisson model (d) histograms are also shown using samples with mean predictions higher than $1.0$ only.}
 \end{center}
@@
 \includegraphics[width=4cm]{../figures/invquant_dayofweek_poisson}
-\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
+\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
 \end{center}
@@
 \includegraphics[width=4cm]{../figures/invquant_yhat_mean_poisson}
-\caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (left) and a simpler Poisson model for comparison (right).}
+\caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
 \end{center}

### 2020-09-10 code/walmart_demand.py

@@ -91,6 +91,7 @@ def calculate_factors_width(est, X):
 
 def plot_factors_ts(df, filename):
     plt.figure()
+    ax = plt.axes()
     df.index = df['date']
 
     df['item_store'].plot(label="item_store")
@@ -106,13 +107,15 @@ def plot_factors_ts(df, filename):
 #    df['yhat_mean'].plot(label="yhat_mean")
 
     plt.legend(loc=1, fontsize=13)
-    plt.ylabel("factors", fontsize=13)
+    plt.ylabel("mean factors", fontsize=13)
+    plt.text(-0.1, -0.24, 'b)', fontsize=15, transform=ax.transAxes)
     plt.tight_layout()
     plt.savefig(filename)
 
 
 def plot_factors_ts_width(df, filename):
     plt.figure()
+    ax = plt.axes()
     df.index = df['date']
 
     df['item_store'].plot(label="item_store")
@@ -128,38 +131,55 @@ def plot_factors_ts_width(df, filename):
 #    df['c'].plot(label="c")
 
     plt.legend(loc=1, fontsize=13)
-    plt.ylabel("factors", fontsize=13)
+    plt.ylabel("variance factors", fontsize=13)
+    plt.text(-0.1, -0.24, 'c)', fontsize=15, transform=ax.transAxes)
     plt.tight_layout()
     plt.savefig(filename)
 
 
 def plot_pdf(n, p):
     plt.figure()
+    ax = plt.axes()
     x = range(30)
     plt.plot(x, nbinom.pmf(x, n, p))
     plt.xlabel("sales", fontsize=15)
-    plt.tight_layout()
+    plt.title('PDF', fontsize=15)
+    plt.text(-0.05, -0.15, 'a)', fontsize=15, transform=ax.transAxes)
+    plt.tight_layout(rect=(0,0,1,0.99))
     plt.savefig('plots/pdf.pdf')
 
 
 def plot_cdf(n, p, y):
     plt.figure()
+    ax = plt.axes()
     x = range(30)
     plt.plot(x, nbinom.cdf(x, n, p))
     plt.vlines(y, ymin=0, ymax=1, linestyles ="dashed")
     plt.xlabel("sales", fontsize=15)
-    plt.tight_layout()
+    plt.title('CDF', fontsize=15)
+    plt.text(-0.05, -0.15, 'b)', fontsize=15, transform=ax.transAxes)
+    plt.tight_layout(rect=(0,0,1,0.99))
[…]

### 2020-09-12 LaTex/paper.tex
review changes, bib cosmetics

@@
 The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g.
-in \cite{beutel2012safety,ban2019big,bertsimas2020predictive, oroojlooyjadid2020applying}.
+in \cite{beutel2012safety,ban2019big,bertsimas2020predictive,oroojlooyjadid2020applying}.
 It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly.
@@
 This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
-Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to  fully controlled for.
+Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to fully control for.
 
@@
 into account.
-This can be done in several ways: Quantile regression \cite{koenker2001, wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
+This can be done in several ways: Quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
 Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
@@
 This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
-The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\mu = \sigma^2$.
+The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$.
 Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
@@
 For a first qualitative assessment, we make use of the probability integral transform, see e.g.
-\cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed.
+\cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1.
 We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
@@
 However, calculating the fractions of the inverse quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts.
-Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand.
+Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand known at prediction time.
 
@@
 The methods discussed so far allow a detailed qualitative evaluation of PDF predictions.
-However, in order to also quantify the quality of the PDF predictions, a metric is required that assesses the difference between the distributions.
-Therefore, we want to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
+In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two probability distributions to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
 Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, or the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius.
@@
 
-Since $0.5$ represents the maximum value of the first Wasserstein distance (its minimum being zero), we define an accuracy measure for our PDF predictions in the range $[0, 1]$ by:
+Since $0.5$ represents the maximum value of the first Wasserstein distance when comparing any distribution in the support $[0, 1]$ to a flat distribution in the same interval (its minimum being zero), we define an accuracy measure for our PDF predictions in the range $[0, 1]$ by:
 \begin{equation}
@@
 \noindent
-While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the last bins of CDF values close to $1$ as well as in the first bins of CDF values close to $0$.
-These cases are mainly due to slow moving articles with less than one item sold per store per day on average.
-In practice, predicting the demand for these kind of items is very challenging as they are prone to over-dispersion and large fluctuations due to the discrete nature of the sales event.
-In the lower part of \fig \ref{fig:cdf_demand}, the effects of these slow-sellers are reduced by excluding all samples with mean predictions lower than $1.0$.
+While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction).
+The first case is mainly due to a slight zero inflation of actual sales.
+The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$.
 The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.

### 2020-09-12 LaTex/paper.tex
journal submission

@@
 \author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
-\author[1]{Moritz Wolf \thanks{martin.hahn@blueyonder.com}}
+\author[1]{Moritz Wolf \thanks{moritz.wolf@blueyonder.com}}
 \author[2]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
@@
 While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction).
-The first case is mainly due to a slight zero inflation of actual sales.
+The first case is mainly due to a slight zero-inflation of actual sales.
 The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$.
@@
 And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter.
-The deviations at very low predicted mean values reflect the complications of slow-sellers described in the context of CDF histograms above.
+The deviations at very low predicted mean values reflect the complications of zero-inflation and slow-sellers seen in the first and last bins of the CDF histograms above.
 

### 2020-09-21 LaTex/paper.tex
arxiv submit

@@
+% Springer
+%\documentclass[smallextended]{svjour3}
+
+%Standard Article
+%\documentclass[10pt,twoside=true, twocolumn]{article}
+
+% KOMA Article
 \documentclass[BCOR=1mm, DIV=calc,10pt,
@@
 
-\author[1]{Felix Wick \thanks{felix.wick@blueyonder.com}}
-\author[3]{Ulrich Kerzel \thanks{u.kerzel@iubh-fernstudium.de}}
-\author[1]{Martin Hahn \thanks{martin.hahn@blueyonder.com}}
-\author[1]{Moritz Wolf \thanks{moritz.wolf@blueyonder.com}}
-\author[2]{Trapti Singhal \thanks{trapti.singhal@blueyonder.com}}
-\author[1]{Michael Feindt \thanks{michael.feindt@blueyonder.com}}
+%Springer
+%\author{
+%\underline{Felix Wick} \and
+%Ulrich Kerzel \and
+%Martin Hahn \and
+%Moritz Wolf \and
+%Trapti Singhal \and
+%Michael Feindt}
+%
+%\institute{ 
+%F. Wick (corresponding author) \at 
+%Blue Yonder GmbH \\ \email{felix.wick@blueyonder.com}
+%\and
+%U. Kerzel \at 
+%IUBH Internationale Hochschule \\ \email{u.kerzel@iubh-fernstudium.de}
+%\and
+%M. Hahn \at
+%Blue Yonder GmbH \\ \email{martin.hahn@blueyonder.com}
+%\and
+%M. Wolf \at 
+%Blue Yonder GmbH \\ \email{moritz.wolf@blueyonder.com}
+%\and
+%T. Singhal \at
+%Blue Yonder India Private Limited \\ \email{trapti.singhal@blueyonder.com}
+%\and
+%M. Feindt \at
+%Blue Yonder GmbH \\ \email{michael.feindt@blueyonder.com}
+%}
+
+\author[1]{Felix Wick } %\thanks{felix.wick@blueyonder.com}}
+\author[3]{Ulrich Kerzel }%\thanks{u.kerzel@iubh-fernstudium.de}}
+\author[1]{Martin Hahn }%\thanks{martin.hahn@blueyonder.com}}
+\author[1]{Moritz Wolf }%\thanks{moritz.wolf@blueyonder.com}}
+\author[2]{Trapti Singhal }%\thanks{trapti.singhal@blueyonder.com}}
+\author[1]{Michael Feindt }%\thanks{michael.feindt@blueyonder.com}}
 \affil[1]{\small Blue Yonder GmbH (Ohiostra{\ss}e 8, 76149 Karlsruhe, Germany)}
@@
 \date{}
+% Springer
+%\titlerunning{Demand Forecasting as a Probability Distribution}
+%\authorrunning{F. Wick et al.}
 
@@
 {Keywords: \textbf{explainable machine learning, demand forecasting, probability distribution}}
+%Springer
+%\keywords{explainable machine learning \and demand forecasting \and probability distribution \and quantile plot}
 
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/PdfCdf}
+\includegraphics[width=8cm]{figs/PdfCdf}
 \caption{\label{fig:PdfCdf} Probability distribution function $f(x)$ and cumulative distribution function of a normal distribution $F(x)$.}
@@
 \begin{center}
-\includegraphics[scale=0.25]{../figures/cdf_truth_uniform}
-\includegraphics[scale=0.25]{../figures/cdf_truth_broad}
-\includegraphics[scale=0.25]{../figures/cdf_truth_narrow}
-\includegraphics[scale=0.25]{../figures/cdf_truth_over}
-\includegraphics[scale=0.25]{../figures/cdf_truth_under}
+\includegraphics[scale=0.25]{figs/cdf_truth_uniform}
+\includegraphics[scale=0.25]{figs/cdf_truth_broad}
+\includegraphics[scale=0.25]{figs/cdf_truth_narrow}
+\includegraphics[scale=0.25]{figs/cdf_truth_over}
+\includegraphics[scale=0.25]{figs/cdf_truth_under}
 \caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/invquant_example}
+\includegraphics[width=8cm]{figs/invquant_example}
 \caption{\label{fig:invquant_example} Five different inverse quantile profile plots (separated into columns by dotted vertical lines), each comparing all its predicted PDFs to the corresponding observed events.
@@
 \begin{center}
-\includegraphics[width=8cm]{../figures/ts_full}
+\includegraphics[width=8cm]{figs/ts_full}
 \caption{\label{fig:mean_prediction} Time series of mean predictions and sales in test period summed over all products and stores.}
@@
 \begin{center}
-\includegraphics[width=4cm]{../figures/pdf}
-\includegraphics[width=4cm]{../figures/cdf}
+\includegraphics[width=4cm]{figs/pdf}
+\includegraphics[width=4cm]{figs/cdf}
 \caption{\label{fig:pdf_example} Predicted PDF (a) and CDF (b) for a specific product-location-date combination.
@@
 \begin{center}
-\includegraphics[width=4cm]{../figures/ts_item_16_store_6}
-\includegraphics[width=4cm]{../figures/factors_ts}
-\includegraphics[width=4cm]{../figures/factors_ts_width}
+\includegraphics[width=4cm]{figs/ts_item_16_store_6}
+\includegraphics[width=4cm]{figs/factors_ts}
+\includegraphics[width=4cm]{figs/factors_ts_width}
 \caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (a), constituting factors of the mean predictions (b), and constituting factors of the variance predictions (c), for a specific product-location combination from beginning of February to end of April 2016.
@@
 \begin{center}
-\includegraphics[width=4cm]{../figures/cdf_truth_nbinom}
-\includegraphics[width=4cm]{../figures/cdf_truth_poisson}
-\includegraphics[width=4cm]{../figures/cdf_truth_nbinom_larger1}
-\includegraphics[width=4cm]{../figures/cdf_truth_poisson_larger1}
+\includegraphics[width=4cm]{figs/cdf_truth_nbinom}
+\includegraphics[width=4cm]{figs/cdf_truth_poisson}
+\includegraphics[width=4cm]{figs/cdf_truth_nbinom_larger1}
+\includegraphics[width=4cm]{figs/cdf_truth_poisson_larger1}
 \caption{\label{fig:cdf_demand} Histograms of ex post observed target CDF values of the corresponding individual PDF predictions (to be compared to a uniform distribution) for our negative binomial model (a) and a simpler Poisson model for comparison (b), using all product-location-day combinations in the test period.
@@
 \begin{center}
-\includegraphics[width=4cm]{../figures/invquant_dayofweek_nbinom}
-\includegraphics[width=4cm]{../figures/invquant_dayofweek_poisson}
+\includegraphics[width=4cm]{figs/invquant_dayofweek_nbinom}
+\includegraphics[width=4cm]{figs/invquant_dayofweek_poisson}
 \caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
@@
 \begin{center}
-\includegraphics[width=4cm]{../figures/invquant_yhat_mean_nbinom}
-\includegraphics[width=4cm]{../figures/invquant_yhat_mean_poisson}
+\includegraphics[width=4cm]{figs/invquant_yhat_mean_nbinom}
+\includegraphics[width=4cm]{figs/invquant_yhat_mean_poisson}
 \caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
@@
 
+% Authors must disclose all relationships or interests that 
+% could have direct or potential influence or impart bias on 
+% the work: 
+%
+%\section*{Declarations}
+%
+%\subsection*{Funding}
+%No external funding was obtained for this work.
+%
+%\subsection*{Conflict of interest}
+%
+%\noindent
+%Authors Felix Wick and Michael Feindt applied for a US patent ``A System and Method of Cyclic Boosting for Explainable Supervised Machine Learning''.
+%
+%\noindent
+%Authors Felix Wick, Martin Hahn, and Moritz Wolf applied for a US patent ``Causal Factor Machine Learning with Individual Negative Binomial Probability Density Function Estimation''.
+%
+%\noindent
+%Authors Felix Wick and Trapti Singhal applied for a US patent ``Evaluation of Predictions as Individual Probability Density Functions''.
+%
+%
+%\subsection*{Availability of data and material}
+%Data used in this work are public \cite{kaggle_data}
+%
+%\subsection*{Code availability}
+%The details of the code are available from \cite{Wick2019}
+%
+%\subsection*{Authors' contributions}
+%F. Wick: Code examples and manuscript text, U. Kerzel: manuscript text, M. Hahn, M. Wolf, T. Singhal, M. Feindt: revision
+
 \bibliography{paper}
+
+% Springer
+%\bibliographystyle{spmpsci}      % mathematics and physical sciences
 \bibliographystyle{ieeetr}
@@
 \begin{center}
-\includegraphics[scale=0.5]{../figures/marginal} \includegraphics[scale=0.5]{../figures/scatter}
+\includegraphics[scale=0.5]{figs/marginal} \includegraphics[scale=0.5]{figs/scatter}
 \caption{\label{fig:scatter} Marginal distribution and scatter plot of variables $X$ and $Y$.}
@@
 \begin{center}
-\includegraphics[scale=0.5]{../figures/profile}
+\includegraphics[scale=0.5]{figs/profile}
 \caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}

### 2020-12-20 LaTex/paper.tex
included temporal confounding, reviewed intro and lit rev

@@
-% Springer
-%\documentclass[smallextended]{svjour3}
-
 %Standard Article
@@
 
-\newcommand{\fig}{Fig.}
-\newcommand{\eqn}{Eqn.}
-\newcommand{\tab}{Tab.}
+\newcommand{\fig}{Fig.~}
+\newcommand{\eqn}{Eqn.~}
+\newcommand{\tab}{Tab.~}
 \newcommand{\wrt}{w.r.t.}
@@
 
-\title{Demand Forecasting of individual Probability Density Functions with Machine Learning}
+\title{Demand Forecasting of Individual Probability Density Functions with Machine Learning}
 
-%Springer
-%\author{
-%\underline{Felix Wick} \and
-%Ulrich Kerzel \and
-%Martin Hahn \and
-%Moritz Wolf \and
-%Trapti Singhal \and
-%Michael Feindt}
-%
-%\institute{ 
-%F. Wick (corresponding author) \at 
-%Blue Yonder GmbH \\ \email{felix.wick@blueyonder.com}
-%\and
-%U. Kerzel \at 
-%IUBH Internationale Hochschule \\ \email{u.kerzel@iubh-fernstudium.de}
-%\and
-%M. Hahn \at
-%Blue Yonder GmbH \\ \email{martin.hahn@blueyonder.com}
-%\and
-%M. Wolf \at 
-%Blue Yonder GmbH \\ \email{moritz.wolf@blueyonder.com}
-%\and
-%T. Singhal \at
-%Blue Yonder India Private Limited \\ \email{trapti.singhal@blueyonder.com}
-%\and
-%M. Feindt \at
-%Blue Yonder GmbH \\ \email{michael.feindt@blueyonder.com}
-%}
-
-\author[1]{Felix Wick } %\thanks{felix.wick@blueyonder.com}}
-\author[3]{Ulrich Kerzel }%\thanks{u.kerzel@iubh-fernstudium.de}}
-\author[1]{Martin Hahn }%\thanks{martin.hahn@blueyonder.com}}
-\author[1]{Moritz Wolf }%\thanks{moritz.wolf@blueyonder.com}}
-\author[2]{Trapti Singhal }%\thanks{trapti.singhal@blueyonder.com}}
-\author[1]{Michael Feindt }%\thanks{michael.feindt@blueyonder.com}}
+% Michael am Ende als "PI"
+\author[1]{Felix Wick}
+\author[3]{Ulrich Kerzel}
+\author[1]{Martin Hahn}
+\author[1]{Moritz Wolf}
+\author[2]{Trapti Singhal}
+\author[1]{Daniel Stemmer}
+\author[1]{Jakob Ernst}
+\author[1]{Michael Feindt}
 \affil[1]{\small Blue Yonder GmbH (Ohiostra{\ss}e 8, 76149 Karlsruhe, Germany)}
@@
 \date{}
-% Springer
-%\titlerunning{Demand Forecasting as a Probability Distribution}
-%\authorrunning{F. Wick et al.}
 
@@
 \begin{abstract}
-Demand forecasting is a central component for many aspects of supply chain operations, as it provides crucial input for subsequent decision making like ordering processes.
-While machine learning methods can significantly improve prediction accuracy over traditional time series forecasting, the calculated predictions are often just point estimations for the conditional mean of the underlying probability distribution, and the most powerful approaches, like deep learning, are usually opaque in terms of how its individual predictions can be interpreted.
-Using the novel supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted instead of single numbers.
-While metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare and this work proposes new techniques for both qualitative and quantitative evaluation methods.
-Additionally, each single prediction obtained with this framework is explainable.
-This is a major benefit in particular for practitioners, as this allows them to avoid ``black-box'' models and understand the contributing factors for each individual prediction.
+
+Demand forecasting is a central component of the replenishment process for retailers, as it provides crucial input for subsequent decision making like ordering processes.
+In contrast to point estimates, such as the conditional mean of the underlying probability distribution or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand.
+Whereas metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare, and this work proposes new techniques for both qualitative and quantitative evaluation methods.
+Using the supervised machine learning method ``Cyclic Boosting'', complete individual probability density functions can be predicted such that each prediction is fully explainable.
+This is of particular importance for practitioners, as it allows to avoid ``black-box'' models and understand the contributing factors for each individual prediction.
+Another crucial aspect in terms of both explainability and generalizability of demand forecasting methods is the limitation of the influence of temporal confounding, which is prevalent in most state of the art approaches.
+
 \end{abstract}
 
-{Keywords: \textbf{explainable machine learning, demand forecasting, probability distribution}}
-%Springer
-%\keywords{explainable machine learning \and demand forecasting \and probability distribution \and quantile plot}
+{Keywords: \textbf{explainable machine learning, retail demand forecasting, probability distribution, temporal confounding}}
 
 \section{Introduction}
-\label{sec:intro}
+\label{sec:Intro}
 
-Demand forecasting is one of the main challenges for retailers and is at the core of supply chain business operations.
-Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and can be interpreted as a random variable that is described by an appropriate probability density function (PDF).
-Demand estimation is further complicated by the fact that retailers typically only observe realized sales rather than the actual demand, and, in case the demand exceeds the current stock level, the data become censored.
-Demand forecasting is an application of time series forecasting and it is important to note that demand as a random variable is not identically and independently distributed (i.i.d.) between different items and sales locations.
-While the probability distribution describing the demand can be attributed to a given family or parameterization, the exact parameters vary: Seasonal effects, finite life cycles of products, and the introduction of new products influence the demand distribution, as well as the local weather at the retail location or the retail location itself in terms of size, assortment range, customer diversity, and other factors.
-The retailers themselves also actively influence demand by using advertisements to highlight products, offering rebates or discounts for specific products, as well as pursuing an active pricing strategy.
-This means that while we can generally assume that demand follows a specific type of probability distribution, its parameters are unique to the instance for which an estimate is required.
-For example, the probability distribution governing the demand of a particular item is specific to the item, date, and retail location for which the forecast is made, and depends on a wide range of further influencing factors.
+% specific quantile
+One of the main business operations of retailers is to ensure that sufficient goods are available in the store for customers to buy.
+This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler.
+In the second step, aimed at inventory control, we need to place the actual orders, which is further complicated by the fact that multiple operational constraints, such as lot-sizes or delivery schedules, need to be included in the final ordering decision.
+Unfortunately, from the perspective of the retailer, the two parts, estimating the future demand and optimization of the order quantity, do not factorize into two separate issues, but are inextricably linked.
+This is because the business strategy depends on the complete chain of demand estimation and subsequent replenishment process to control the inventory.
+We can understand this in the following way: In the simplest case, a retailer might want to optimize their profit, or equivalently the cost, by reducing the impact of perishable items that have not been sold at the end of the selling period, as well as the lost revenue by not being able to meet the customer demand within this period.
+This is known as the ``newsvendor problem'' \cite{Edgeworth}, see, e.g., \cite{Khouja1999537} for a detailed review.
+Assuming linear overage ($h$) and underage ($b$) cost, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known probability density function (PDF) describing demand as random variable can be calculated exactly.
+This quantile then corresponds to the point estimator corresponding to the numerical value of the demand prediction that can be used to determine the optimal order quantity.
 
-Because of the access to  Big Data, a vast amount of these data associated with demand, such as historic sales records, information about promotional events or advertisements, pricing information, local weather at retail locations, seasonal information, as well as many further variables, can be captured, stored, and processed.
-Different methods, such as traditional univariate time series forecasting or modern supervised machine learning algorithms, can then be used to predict the demand distribution for individual products sold at specific locations, corrected for censored data.
-Demand forecasting for retailers often requires to predict millions of item-store combinations daily, and traditional univariate time series forecasting methods operate on each of these many individual time series separately.
-Machine learning, on the contrary, optimizes all individual time series together, improving the generalizability of the method by exploiting common aspects between them.
-This leads to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series.
-Furthermore, machine learning enables a natural way to include the many exogenous variables influencing demand as features of the model, what reduces also the bias compared to traditional time series forecasting approaches which mainly rely on the historic sales.
+% simulations
+In reality, the situation is often more complex, as many factors may contribute to the cost function.
+Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time.
+However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF.
+Appendix \ref{sec:CostQuantile} demonstrates how the optimal point estimator can be calculated using the full PDF.
+For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, the calculations are straight-forward.
+If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches.
+Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different points, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change when choosing a different quantile.
+This can be done using, for example, a business oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business relevant KPIs for each setting, taking operational constraints into account.
+This can then be used, for example, to align the business strategy.
+Predicting the full demand PDF therefore allows the maximum flexibility for further optimizations, because we have direct access to all quantiles.
 
-In order to make operational decisions, an optimal demand point estimator has to be defined, that can be used to derive ordering decisions in the replenishment process of the retailer.
-The ordering decision process is complicated by a range of factors: Even in the case of perfect demand forecasts, the decision maker has to consider lot sizes defined by the wholesaler or manufacturer as well as to balance conflicting metrics to reach an optimal decision.
-Ordering too few items may result in out of stock situations leading to unrealized demand and unsatisfied customers.
-Ordering too many items results in excess inventory, which increases transport and storage costs and, in the case of perishable goods, excessive waste, as spoiled items need to be disposed of at additional cost and potentially even environmental impact.
-This situation is particularly noticeable in the so called ``ultra-fresh'' category, which includes items such as bakery products, ready-meals, fresh dairy products, or certain meat products such as ground meat.
-These items typically have a shelf-life ranging from less than a business day to a few business days at most, with a continuous spectrum in between, depending on the exact item.
-In many situations, additional constraints have to be considered to reach an optimal ordering decision: Delivery cycles of items may vary depending on the type of item and the wholesaler or manufacturer from which they are procured.
-Retailers also operate at a given service level to guarantee that a certain level of demand can be fulfilled.
-The exact service level typically depends on the overall business policy of the retailer and may also depend on individual products, ranging from ``never-out-of-stock'' items to a service level exceeding, e.g., 90 percent.
-Therefore, demand forecasts are ideally made in the form of full individual probability density distributions, from which optimal point estimators can be derived, which in turn serve as starting point to determine the subsequent ordering decisions.
+% PDF predictions
+Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated PDF.
+And as such, demand is neither identically and independently distributed (i.i.d.) between different products and sales locations, nor between different times.
+This means that, while we can generally assume that demand follows a specific type of PDF, its parameters are unique to the instance for which an estimate is required.
+For example, the PDF describing the demand of a particular product is specific to this product, date, and retail location for which the forecast is made.
+Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc.
+Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDF predictions rather than globally adjusted parameters for groups of SKUs or locations or sales days.
+
+% temporal confounding
+If we account for the cases where demand exceeds the current stock level, resulting in out-of-stock situations, the actual demand in the past can be approximated by realized sales.
+We can then understand the time-ordered observations of sales events as a time series of a random variable.
+It is therefore tempting to use established time series regression models to predict the future values of the expected demand.
+However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed variables $Y$ used in the time series, say $Y_{t}$ and $Y_{t-\tau}$ with lag $\tau$.
+In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion.
+This dependency creates a backdoor path between $Y_{t-\tau}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$.
+This is particularly important for the case of demand forecasting, since customer demand is not a natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc.
+If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelations, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable.
+Moreover, due to this disguising of actual causal dependencies, reliance on temporal confounding is also detrimental in terms of generalizability of the forecasting model.
+The issue of temporal confounding, together with a proposed solution, is discussed in more detail in sec.
+\ref{sec:ts_confounding}.
+
+% ML
+Demand forecasting for retailers often requires to predict millions of product-location combinations for each SKU each day.
+Univariate time series forecasting methods operate on each of these many individual time series separately.
+On the other hand, supervised machine learning (ML) algorithms optimize all individual time series together, which can enhance the generalizability of the method by exploiting common aspects between the various time series.
+This can lead to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series.
+Furthermore, machine learning provides a straight-forward way to include the many exogenous variables influencing demand as features (or covariates) of the model.
+This can also help to improve the forecast quality compared to methods mainly relying on autocorrelation of the historic sales.
+
+Hence, in order for a retailer to be able to use demand forecasts most effectively, we need to be able to:
+\begin{itemize}
+	\item {Predict the future demands on SKU level at required temporal granularity as complete individual PDFs.}
+	\item {Avoid temporal confounding in the model used for the prediction of the future demand.}
+	\item {Evaluate that the functional form of the PDF is correct.}
+	\item {Evaluate that the shape of the PDF is correct and that each part of the PDF models the observed data appropriately.}
+	\item {Ideally, each prediction is fully explainable, enabling operational experts to exactly understand what contributed to it.}
+\end{itemize}
 
@@
 \ref{sec:LitRev}.
-We then describe our method to predict individual negative binomial PDFs by means of a parametric approach including two distinct machine learning models for mean of variance in sec.
+We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean of variance in sec.
 \ref{sec:pdfEstimation}.
@@
 \ref{sec:pdfEvaluation}.
+This is followed by a detailed discussion on temporal confounding in sec.
+\ref{sec:ts_confounding}.
 And finally, we present a demand forecasting example to show an application of our methods in sec.
 \ref{sec:example}.
+
+\subsection*{Summary of Contributions}
+
+The contribution of this paper is two-fold:
+
+Using the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019}, we show how complete probability distributions of the future demand of individual products in retail stores can be predicted at the operationally required granularity.
+In particular, we show how using machine learning, the detrimental effects of temporal confounding, that are intrinsic to many methods used for time series forecasting, can be limited.
+The impact of temporal confounding in the context of time series forecasting has not been widely investigated so far in the scientific literature, and to the best of our knowledge this is the first discussion within the context of demand forecasting.
+
+Furthermore, we show how predictions of full individual probability density functions can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form as well as specific quantiles are predicted correctly.
+This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.
+
+The general methods developed in this paper can be applied to any setting that requires the prediction and evaluation of probability distributions from observational data, not just demand forecasting.
+However, to make the impact and benefits more tangible in a practical setting, we limit the discussion to demand forecasting in retail.
 
@@
 
-Inventory management offers a rich theory and the extensive body of research can be broadly grouped into two categories, where the inventory control problem is either based on some knowledge of the underlying demand distribution or an integrated approach that seeks to map directly from the observed data to the ordering decision.
-More traditional inventory control systems rely on the knowledge of the demand distribution in one form or another, see e.g.
-\cite{silver1998} for an overview.
-In $(s,S)$ type inventory control systems \cite{Scarf1958}, inventory levels are monitored at regular intervals and orders are dispatched once the inventory level reaches a minimal value $s$.
-In case of linear holding and shortage costs, such policies are optimal \cite{Scarf1959}, although perishable goods pose more challenges, see e.g.
-\cite{Nahmias1973,Nahmias1975,nahmias1978}.
-Additionally, service level constraints can be included in these kind of inventory control systems \cite{minner2010periodic}.
-Perishable goods are well described by the ``newsvendor problem'' \cite{Edgeworth}, where, in the simplest case, all stock perishes at the end of the selling period (e.g.
-a business day).
-For a detailed review of the newsvendor problem see e.g.
-\cite{Khouja1999537}.
-Assuming linear underage and overage costs $b,h >0$, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known demand distribution $f(D)$ can be calculated exactly.
-However, it should be noted that demand is not independent between different samples for a given date.
-For example, a promotion applied to one product can affect the sales of related products within the assortment.
-This implies that the demand forecasts for multiple items or a given assortment range cannot be treated as individual newsvendor-type predictions but need to be modeled holistically.
+Demand forecasting is one of the core operational activities of any customer-facing business and accurate demand predictions are vital to the optimization of the business operations.
+This is of particular concern for example for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}.
+It can be approached in a variety of ways, see, e.g., \cite{beheshti2015survey} for a survey of methods used in retail sales forecasting with a particular focus on fashion retail.
 
-The direct approach is often referred to as ``data-driven newsvendor'' and discussed e.g.
-in \cite{beutel2012safety,ban2019big,bertsimas2020predictive,oroojlooyjadid2020applying}.
-It aims to avoid estimating the underlying probability distribution for demand and use the available data (historic sales records and further variables) to derive the operational decisions (i.e., the order quantity) directly.
-An overview of a range of different approaches can also be found in \cite{huber2019data}.
+% univariate time series methods
+A popular approach is to treat the demand as a time series, i.e., as a time-ordered sequence of observed sales events.
+In this setting, the future demand can be predicted as a point estimate using auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins} or exponential smoothing \cite{croston1972forecasting, holt1957forecasting, Brown1963, Gardner1985} approaches.
+See also \cite{deGooijerHyndman2006} for a comprehensive overview over traditional time series forecasting methods.
+Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality or external factors in various ways, depending on the exact modeling approach taken.
+The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $y_t$ that is to be forecasted with some lag $\tau$ between subsequent events.
+Intuitively, we assume that  the sequence of events observed so far is a good predictor of future events.
+Examples for these methods in the context of demand forecasting can be found in, e.g., \cite{huber2017cluster, kalchschmidt2006forecasting, fattah2018forecasting, permatasari2018sales}.
+A different approach is to use a concept from engineering, the Kalman filter \cite{kalman1960new}, and apply it to statistical forecasting \cite{morrison1977kalman}.
+Examples for this method used to estimate demand are, e.g., \cite{mitropoulos1980using, tegene1991kalman, kandananond2014applying,jacobi2007water}.
+Starting from the concept of Kalman filters, state space models \cite{hyndman2008forecasting} relate the measured values over time to the evolution of (unknown) states of the considered system.
+A comparison of the performance of ARIMA and state space models in a retail setting can be found, e.g., in \cite{ramos2015performance}.
+Structural time series models \cite{StructuralTS} take a similar approach and decompose the observations into components for trend, seasonality, and others.
+A modern implementation is Facebook's ``Prophet''  \cite{taylor2018forecasting}.
+Several other approaches are known in the literature as well: K\"ok and Fisher \cite{kok2007demand} model the demand based on the average demand for a product, modified by a logistic regression for the number of customers in a store, purchase incidence, and product choice.
+Wang \etal \cite{wang2005demand} model the demand under the assumption of a Poisson process.
+
+% ML
+Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached using supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.
+For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting.
+Machine learning approaches can include multiple covariates.
+This allows to include not only past sales data, but also information about prices, promotions, or details about the article or store in the forecast.
+There are two main approaches for time series forecasting using  machine learning: We either assume that the variables are independent random variables (for both prediction target and potentially time-dependent features) for each time step of the different time series, or we consider the full sequences of time steps as a whole, resulting in a single sample (including prediction target and potentially time-dependent features) for each individual time series.
+For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model.
+For example, we could add exponentially smoothed moving averages with a range of dampening factors as features into the model.
+An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}.
+For the case of sequence samples, deep learning methods such recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used.
+These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events.
+Both concepts, LSTM cells and self-attention, aim at learning from the data the elements of the individual time series that are deemed important for the prediction of the future values by the algorithm, and implicitly rely on the autocorrelation between the observed events at different times as well.
+The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}.
+A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}.
+A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.
+
+% temporal confounding
+In many situations, relying on the exploitation of autocorrelation, either via traditional time series forecasting or machine learning approaches, is sufficient.
+However, this can lead to shortcomings in terms of explainability and generalizability due to temporal confounding.
+There are two main paths to avoid this: One way is to explicitly account for the confounders and extend the model to include them.
+The main challenge is then the discovery of temporal confounders and, for example, try to learn the causal structure from the observational data, see, e.g., \cite{malinsky2018causal,runge2018causal,Runge2019}.
+This topic is of particular importance in fields like medicine \cite{bica2020time} or climate research \cite{perrakis2014controlling}.
+Brodersen \etal \cite{Brodersen2015,GoogleCausalImpact} propose to use Bayesian structural time series models to estimate the causal effect of interventions on a time series by creating a counterfactual prediction that describes the behavior of the time series had the intervention not taken place.
+While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state at $Y_{t-1}$ to $Y_t$.
+Therefore, great care must be taken to avoid temporal confounding when setting up the states for structural time series models.
+Compared to fields like medicine, marketing, and many others that focus on the causal impact of few or just one intervention(s), retailers have to include many external effects that affect the predicted demand and are repeated frequently: Prices are changed dynamically, advertisements placed, new products are included in the assortment and others removed, etc.
+
+\noindent
+The alternative to the above approaches is to separate the autocorrelation from exogenous information via a chain of independent models (see sec. \ref{sec:ts_confounding}).
+
+% PDF predictions
+Common to all approaches discussed so far is that they predict a point estimate, essentially a single number that indicates the next value in the sequence, for example, the number of a specific product likely to be sold in a given store on a particular opening day.
+While operational decisions can be based on such a single number, this does not allow us to evaluate the uncertainty on such a prediction.
+Since the model parameters are not {\em a priori} known but estimated from data, the predicted values are generally associated with an uncertainty that needs to be quantified.
+Several methods have been developed, and Chatfield \cite{chatfield1993calculating} reviews the methods available at the time.
+In particular, bootstrapping methods for auto-regressive time series forecasting have been proposed, see, e.g.,\cite{masarotto1990bootstrap,mccullough1994bootstrapping,mccullough1996consistent,grigoletto1998bootstrap,thombs1990bootstrap,clements2001bootstrapping,Angus1994,pascual2004bootstrap,pascual2001effects,pascual2005bootstrap}.
+However, as already pointed out in \cite{chatfield1993calculating}, estimating the uncertainty of the point forecast is only one reason to go beyond a point estimate.
+For example, if we want to be able to evaluate strategies for a range of different outcomes, we need access to more information.
+In particular, if we do not wish to make the assumption that the point forecast is the mean of an underlying Gaussian distribution, that can be fully specified by the point forecast and an interval, we need to be able to compute any quantile of the distribution or estimate the full PDF directly.
+
+% quantile regression
+Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
+Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
+Focusing on time series, Hyndman \cite{hyndman1995highest} proposes to use either a simulation or bootstrapping framework to estimate the highest density forecast regions of non-normal time series.
+Tay \cite{tay2000density} summarizes the use of density forecasts in applications in finance and economics.
+Using RNNs, quantile regression for multiple forecasting horizons \cite{wen2017multi} has been used to predict the demand at the internet retailer Amazon.
+Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step.
+
+ % PDF assumption
+Alternatively, we can start from a model assumption for the demand distribution and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}.
+This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated.
+Empirically, we can determine the best fitting distribution from data \cite{adan1995}.
+However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes.
+Instead, the choice of the demand distribution should be motivated by theoretic considerations.
+The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}.
+This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
+The NBD has two parameters, the mean $\mu$ and the variance $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$.
+Hence, for each ordering decision, the model parameters $\mu$ and $\sigma^2$ need to be determined at the required granularity, typically for each item, sales location, and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
+
+% order optimization
+Finally, since from an operational perspective the retailer's focus is less on the demand prediction {\em per se} but on the decisions, in particular, the ordering decisions, that are derived from these, we need to consider the question if we should skip the separate step of estimating the demand and directly predict the optimal order quantity.
+This direct approach is often referred to as ``data-driven newsvendor'' in the recent literature, see, e.g., \cite{beutel2012safety, ban2019big, bertsimas2020predictive, oroojlooyjadid2020applying,huber2019data}.
+It aims to avoid estimating the underlying PDF for demand and use the available data (historic sales records and further variables)  to derive the operational decisions (i.e., the order quantity) directly.
 Although the integrated approach seems preferable at first glance, since it avoids determining the full demand distribution and results directly in the desired operational decision, the indirect approach via demand forecasts offers some substantial advantages.
-First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual item, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out of stock (i.e., lost sales) and waste rate.
+First, demand forecasts in form of full PDFs can be used to simulate the performance of the relevant metrics on the level of individual items, and, for example, optimize the impact on business strategy decisions on conflicting metrics such as out-of-stock (i.e., lost sales) and waste rate.
 From a practitioners perspective, separating the demand forecast from the operational decisions (i.e., calculating the order quantities for the next delivery cycle) enables longer-term planning and reduces the complexity, as it avoids coupling delivery schedules of multiple wholesalers and manufacturers with the forecast of customer demand.
@@
 This is crucial, for example, if a causal analysis is planned, such as a study of the effect of promotions, advertisements, price changes, or other demand shaping factors in either Pearl's do-calculus \cite{PearlCausality} or Rubin's potential outcomes framework \cite{rubin1974estimating}.
-Using an operational quantity such as the order quantity will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that one may not be able to fully control for.
-
-Therefore, the main objective in solving the inventory control problem is to determine the underlying demand distribution.
-The simplest approach is to use the observed sales records and forecast these via sample average approximation (SAA), see, e.g., \cite{shapiro2014} for an overview, or as a time series (see, e.g., \cite{alwan2016}), for example by means of auto-regressive integrated moving average (ARIMA) \cite{BoxJenkins}, exponential smoothing \cite{ExponentialSmoothing}, or structural time series models \cite{StructuralTS}.
-However, these approaches do not make use of any data apart from the sales records themselves, although we know that many variables such as price, advertisements,  weather influence, and others are highly correlated with the demand.
-Saghafian and Tomlin \cite{saghafian2016newsvendor} propose to include partial information about the distribution in the derivation of the operational decision, i.e., the calculation of the optimal order quantity.
-As described in sec.
-\ref{sec:intro}, modern supervised machine learning techniques can overcome many of the issues of the more traditional approaches, and therefore are the method of choice in most use cases today (see, e.g., \cite{tsforecasting,ZHANG199835,rnn,transformer}).
-
-In order to be able to fully optimize the operational decision, it is critical to reconstruct a full demand distribution.
-This also implies that a simple point-estimator, as provided by the most common statistical techniques and machine learning approaches, will not suffice.
-Instead, we need to determine the full demand distribution from data, conditional on the relevant variables such as date, location, and item, taking all auxiliary data such as article characteristics, pricing, advertisements, retail location details, etc.
-into account.
-This can be done in several ways: Quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
-Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
-Alternatively, one can assume a given demand model and fit the model parameters instead of reconstructing the complete distribution \cite{astonpr373, SALINAS20201181}.
-This approach is computationally favorable and usually more robust, as fewer parameters need to be estimated.
-Empirically, one can determine the best fitting distribution from data \cite{adan1995}.
-However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes.
-Instead, the choice of the demand distribution should be motivated by theoretic considerations.
-The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959,Ehrenberg1967,Ehrenberg1972,Chatfield1973,Schmittlein_1985}.
-This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
-The NBD has two parameters, $\mu$ and $ \sigma^2 > \mu$, and is over-dispersed compared to the Poisson distribution, for which $\sigma^2 = \mu$.
-Hence, for each ordering decision, the model parameters $\mu$ and $\sigma$ need to be determined for each item at the required granularity, typically for each sales location and ordering time, depending on all auxiliary data describing article details, retail location, and influencing factors such as pricing and advertisement information.
-
-\subsection*{Summary of Contributions}
-
-This work demonstrates how the explainable machine learning algorithm Cyclic Boosting \cite{Wick2019} can be used to model the demand distribution at the granularity required by the retailer.
-Typically, this means that the full demand PDF has to be estimated per product for each sales location and opening day, conditional on a wide range of variables such as weather, prices, promotions, etc.
-In contrast to using a ``black-box'' machine learning model, Cyclic Boosting allows to interpret each individual prediction in terms of influence of the different features.
-
-Additionally, we show how the cumulative distribution function (CDF) can be used to accurately asses the forecast quality of the full predicted demand distribution, including the tails of the distribution.
-This allows to verify that the predicted demand distribution accurately reflects the observed data and can hence be used both to derive operational decisions such as order quantities as well as strategic business decisions by the retailer, and to gain further insights into customer behavior using for example causal modeling.
+We should note that this is different compared to Granger causality \cite{Granger1969}, that seeks to determine the causal influence one time series has on another.
+In our case, price changes, promotions, and similar are interventions that we actively pursue in order to influence demand.
+In Pearl's do-calculus, we can express this, for example, as $P($demand $|do ($Price=2.99Euro$))$.
+While we can, of course, represent the prices as a chronologically ordered series of numbers, it is not really a time series but a sequence of interventions.
+Using an operational quantity, such as the order quantity, will in most cases act as an insufficient proxy for the quantity of interest (customer demand) and likely lead to unnecessary causal pathways that we may not be able to fully control for.
 
@@
 
-Many statistical and most machine learning methods do not provide a full probability density distribution as result.
+Many statistical and most machine learning methods do not provide a full PDF as result.
 Instead, these methods typically predict a single numerical quantity (denoted by $\hat{y}$) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the mean squared error (MSE), mean absolute deviation (MAD) or others.
@@
 In the following, we will show how different visualizations of the observed cumulative distribution function (CDF) values can be used to evaluate the quality of the predicted PDFs.
-Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a probability density distribution, even if it is obtained empirically.
+Although we limit the following discussion to the negative binomial model, the method can be applied generally to any representation of a PDF, even if it is obtained empirically.
 
@@
 In the simplest case, we only have one model with one set of model parameters to cover all predictions.
-In this case, the evaluation of the full probability distribution is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
+In this case, the evaluation of the full PDF is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
 Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
@@
 
-For a first qualitative assessment, we make use of the probability integral transform, see e.g.
-\cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1.
+For a first qualitative assessment, we make use of the probability integral transform, see e.g., \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1.
 We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
@@
 If the CDF is continuous and strictly increasing, then the inverse of the CDF, $F^{-1}(y)$, exists and is a unique real-valued number $x$ for each $y \in [0,1]$, so that we can write $F(x) = y$.
-The inverse of the CDF is also called the quantile function, because we can define the quantile $\tau$ of the probability distribution $f(x)$ as:
+The inverse of the CDF is also called the quantile function, because we can define the quantile $\tau$ of the PDF $f(x)$ as:
 
@@
 \includegraphics[width=8cm]{figs/PdfCdf}
-\caption{\label{fig:PdfCdf} Probability distribution function $f(x)$ and cumulative distribution function of a normal distribution $F(x)$.}
+\caption{\label{fig:PdfCdf} PDF $f(x)$ and CDF of a normal distribution $F(x)$.}
 \end{center}
@@
 
-We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e.
-$F:t \to s$.
+We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e., $F:t \to s$.
 Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF.
-This implies that the probability distribution of $s$, $g(s)$, is constant in the interval $s \in [0,1]$ in which $s$ is defined, and $s$ can be interpreted as the cumulative distribution of its own probability distribution:
+This implies that the PDF of $s$, $g(s)$, is constant in the interval $s \in [0,1]$ in which $s$ is defined, and $s$ can be interpreted as the CDF of its own PDF:
 
@@
 
-In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g.
-\cite[p.
+In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g., \cite[p.
 54]{casella2002statistical}.
@@
 For example, if the sales of zero items accounts for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
-Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g.
-for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.
+Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g., for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.
 
-A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted probability distribution $f(x)$ is correctly calibrated.
+A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted PDF $f(x)$ is correctly calibrated.
 This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for five different cases.
@@
 In summary, this approach has two major benefits compared to the method discussed in sec.
-\ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the inverse quantile profile plot reveals which part of the predicted PDF, such as the tails of the probability distribution, are particularly problematic.
-Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the inverse quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g.
-specific categories, of that variable.
+\ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the inverse quantile profile plot reveals which part of the predicted PDF, for example the tails, are particularly problematic.
+Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the inverse quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g., specific categories, of that variable.
 Two examples for this can be found in Figs.
@@
 The methods discussed so far allow a detailed qualitative evaluation of PDF predictions.
-In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two probability distributions to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
-Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, or the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius.
+In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two PDFs to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
+Several approaches that measure the difference between two PDFs are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius, or the energy distance \cite{SZEKELY20131249}.
 
-Compared to the other mentioned methods, the first Wasserstein distance, representing the symmetric distance between two probability distributions on a given metric space, is more sensitive to smaller deviations, because it exhibits a linear behavior around zero (reflecting perfect agreement).
+Compared to the Kullback-Leibler and Jensen-Shannon divergences, the first Wasserstein distance, representing the symmetric distance between two PDFs on a given metric space, is more sensitive to smaller deviations, because it exhibits a linear behavior around zero (reflecting perfect agreement).
 Therefore, we focus on the first Wasserstein distance as the measure of difference in the following.
@@
 
+\section{Temporal Confounding in Time Series Forecasting}
+\label{sec:ts_confounding}
+
+\subsection{Origin of Temporal Confounding}
+\label{sec:IntroTempConfound}
+
+The aim of time series forecasting in general is to predict the next values in a series of time-ordered events.
+The quantity we wish to predict is often called the ``prediction target'' or just ``target''.
+In the case of demand forecasting, the prediction target is the expected future demand, and we usually employ the time series of observed sales events in the past, possibly corrected for out-of-stock situations, to estimate the model parameters.
+Many time series forecasting methods exploit the autocorrelation between the values of the variable of interest at different times.
+Depending on the exact nature of the problem we consider in these approaches, different lags between the variable $Y$ describing the demand play an important role.
+For example, a weekly repeating pattern would lead to a strong autocorrelation for a lag of $\tau=7$ (days), meaning that the values of the variable $Y$ at $y_t$ and $y_{t-\tau=7}$ would be strongly autocorrelated.
+However, this autocorrelation between different times or lags of the target variable may not be a true causal dependency but a spurious correlation due to one or more temporal confounders.
+For example, if an exogenous variable has a direct causal impact at the variable at time, say,  $Y_{t-1}$, this external variable will often also have an impact at time $Y_t$.
+Intuitively, if the variable $Y$ was influenced yesterday by some external event, there is a high chance that we can observe this impact today as well.
+
+The effect of temporal confounding is illustrated in \fig \ref{fig:temporal_confounding}, where the variable $Y$ is observed at two different times $t-1$ and $t$ with values $y_{t-1}$ and $y_t$, and the variable $X$ is an external influencing factor that affects the variable $Y$ both at time $t-1$ and at time $t$.
+For example, we can imagine that $X$ represents the price of a product that is changed due to a special promotion at these two time steps.
+Looking at the causal structure of the directed acyclic graph (DAG) created by this dependency, we see that  $X$ takes the role of a (temporal) confounder that creates a backdoor path between $Y_{t-1}$ and $Y_t$.
+This in turn means that there is now a spurious (auto)correlation between $Y_{t-1}$ and $Y_t$, indicated by the dashed line in \fig \ref{fig:temporal_confounding}.
+Therefore, if we do not control for the temporal confounder $X$, but use the autocorrelation between $y_{t-1}$ and $y_t$ to forecast the next value $y_{t+1}$, we rely on a spurious correlation between $Y_{t-1}$ and $Y_t$ that is not ``real'' in the sense that it results from a causal influence of $Y_{t-1}$ on $Y_t$, but originates from the backdoor path via $X_{t, t-1}$.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=4cm]{figs/temporal_confounding}
+\caption{\label{fig:temporal_confounding} Temporal confounding by an exogenous variable $X_{t-1, t}$ stable over the time steps $t-1$ and $t$, creating a non-causal autocorrelation between $Y_{t-1}$ and $Y_t$.}
+\end{center}
+\end{figure}
+
+In the case of retail demand forecasting, we typically have many temporal confounders, as demand is influenced by a multitude of different causes such as prices, advertisements, weather, and many more.
+Some of these are due to interventions, whereas others are caused by external effects we cannot influence.
+For example, setting the price of a product or running an advertisement campaign is an intervention (in the sense of causal analysis) by the retailer, whereas the impact of the forecasted weather at the sales location and sales time on the expected demand is outside our control.
+And in addition to the spurious autocorrelations due to all these temporal confounders, there might also be a direct impact of $Y_{t-1}$ on $Y_t$, resulting in an autocorrelation originating from a ``real'' cause.
+If we can control for the confounder(s) $X$, we can remove, or at least limit, the spurious autocorrelation and in turn improve the forecasting method, especially in terms of its generalizability, by focusing on the description of the remaining ``real'' autocorrelation.
+
+\subsection{Separation of Causal Features from Autocorrelation in Machine Learning Models}
+
+% time series with ML and ewma features
+Using machine learning, we express the task of time series forecasting as a supervised learning approach, where the prediction target, i.e., the quantity we wish to forecast, is represented by the values of the time series of interest itself.
+In the case of demand forecasting, the future expected demand is the variable we wish to forecast, and we need to build feature variables, either by hand or using some automatic approach, to construct a supervised machine learning model.
+A common approach is to exploit the autocorrelation of the target time series variable to build dedicated feature time series from its summary statistics of the respective recent past.
+For example, we could use an exponentially weighted moving average (EWMA), calculated according to \eqn \eqref{eqn:ewma}, for each individual time series and choose several different values for the smoothing parameter $\alpha$, which is defined in the range $[0, 1]$.
+\begin{equation} \label{eqn:ewma}
+\begin{split}
+&\mathrm{EWMA}(x_{t+\mathrm{horizon}}) = \\ &\frac{x_t + (1 - \alpha) x_{t-1} + (1 - \alpha)^2 x_{t-2} + ... + (1 - \alpha)^t x_0}{1 + (1 - \alpha) + (1 - \alpha)^2 + ... + (1 - \alpha)^t}
+\end{split}
+\end{equation}
+We can then use these EWMA features in the supervised machine learning model that we use to predict the future value of the target variable, in our case, the future demand for a given product at a given location.
+When constructing these features, we need to keep in mind that we use a lag with respect to the target time series corresponding to the forecast horizon.
+For example, if we were to forecast three days ahead into the future, the lag of the feature time series with respect to the target time series has to be three as well.
+By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past, and use these as features for the machine learning model.
+In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the product we wish to forecast, or characteristics of the location at which the product is sold.
+
+% problem statement
+Although this approach is often taken in practice, it is prone to a detrimental effect of temporal confounding: Because we do not control for temporal confounders and exploit the autocorrelation of the target time series using its lagged information directly in form of, for example, an EWMA, we obscure the underlying causal dependencies and leave the temporal backdoor paths created by them open.
+This in turn makes learning the true causal dependencies between the exogenous variables and the quantity of interest much more difficult, because the machine learning algorithm can now learn the statistical dependencies of the spurious autocorrelation encoded in the time series features, rather than learn the more fundamental influences from variables such as price information, weather conditions, and similar.
+For example, if we wish that the model learns the dependency, say, the price of a product or a promotion has on the customer demand for the product, we provide this information twice: Once by explicitly adding the price or the promotion details as features to the machine learning model, and then again implicitly by using lagged target information, such as EWMA features of the observed sales in the past.
+Looking back at \fig \ref{fig:temporal_confounding}, the EWMA contains the spurious autocorrelation created by the backdoor path via the exogenous variable(s) (such as the price or promotion details).
+The machine learning algorithm is not aware of the causal structure and its influences on the time series, but only ``sees'' the correlation to the prediction variable.
+Therefore, the machine learning algorithm is prone to learn the spurious autocorrelation encoded in the EMWA features, as it cannot distinguish these from the true causal influences.
+
+As forecasting relies upon statistical dependencies, no matter if spurious or from true causal effects, this does not seem to be a problem for the predictive power at first glance.
+However, it can have severe impacts on the explainability and generalizability of the model, because the autocorrelation, being a composed effect of many causes, is less stable over time than the actual causal effects.
+A concrete weakness of the reliance on autocorrelation can be seen when looking at effects at isolated time steps.
+For example, if the target variable has a sharp peak that is induced by a sudden change of an exogenous variable with a causal influence on the target, the autocorrelation, lagged by its nature, would not help to predict the peak at the very day it occurs, but would tend to ``trick'' the model in predicting a delayed peak, which in turn reduces the accuracy of the prediction at other time steps after the actual peak as well.
+
+Note that the effect of temporal confounding does not only apply to this specific approach of building features such as the EMWA to be used in supervised machine learning, but to all time series forecasting methods that rely on the autocorrelation between events at various lags, including ARIMA, Holt-Winters, Kalman filters and the related variants, RNNs and its variants like LSTMs, and others summarized in sec.
+\ref{sec:LitRev}.
+
+% separate
+In order to avoid these issues, we suggest to separate the target autocorrelation from the exogenous dependencies.
+This can be done by avoiding any features in the machine learning model that include lagged target information.
+Using the example from above, this would mean that we include features such as the price of the product or promotion details, weather information, as well as other variables describing the characteristics of a given product at a specific sales location, but do not include lagged information from the sales time series, such as its EWMA or similar approaches.
+Therefore, the features do not include any variable that is directly constructed from the previously observed sales events.
+As a consequence, the machine learning model can learn from the true causal dependencies expressed in the feature variables, which in turn improves both the explainability of the predictions, as well as the generalizability of the prediction model, because the causal dependencies are not confused by open backdoor paths leading to spurious (auto)correlations between the features and the prediction target.
+
+However, we need to keep in mind that this approach of handling temporal confounders improves the prediction model used for demand forecasting, but is not yet sufficient for true causal inferences.
+For this, we would also have to control for the confounding effects between the different exogenous features of the machine learning model and its prediction target.
+
+\subsection{Autocorrelated Residual Correction}
+
+By including all available exogenous variables as features in the machine learning model but excluding any endogenous information from the target autocorrelation we can, at least partially, control for the temporal confounders.
+Any remaining autocorrelation between different lags of the time series of the residuals formed from the original time series and the time series of the forecasted value from the machine learning model discussed above can therefore originate from two reasons: Either there is a genuine autocorrelation between different lags $Y_{t-\tau}$ and $Y_t$ of the variable we wish to forecast, or we have not included all relevant causal variables in the machine learning model.
+To make use of this remaining autocorrelation, we suggest to apply a residual correction on each of the predictions of the machine learning model, based on the deviations in the recent past between the target and the predictions from the machine learning model.
+Such a residual correction heuristically captures the recent trends without modeling them explicitly, and can be seen as an empirical correction to improve the final forecast quality.
+
+% ewma
+One possibility for the implementation of the residual correction is to apply exponential smoothing.
+In the simplest case, this can be done as stated in eqn.
+\ref{eqn:emov_correction}, where the two EWMA terms are grouped by the individual time series and calculated according to eqn.
+\ref{eqn:ewma}.
+\begin{equation} \label{eqn:emov_correction}
+\hat{y} = \frac{\mathrm{EWMA}(y)}{\mathrm{EWMA}(\hat{y})} \cdot \hat{y}_{\mathrm{ML}}
+\end{equation}
+In most cases, simple exponential smoothing will be sufficient for the residual correction, since we have already incorporated all relevant external effects, including seasonalities like a dependency of the target value on the day of the week, in the machine learning model before, and only seek to capture the information that we have missed, because, for example, we do not have the relevant data to describe this as a further feature variable.
+However, we need to be aware that if we have missed a crucial temporal confounder when building the features for the machine learning model, applying this residual correction risks opening a backdoor path for this confounder, which in turn might result in slightly sub-optimal predictions, as discussed in sec.
+\ref{sec:IntroTempConfound}.
+We should therefore take care to use the relevant domain knowledge to capture the causal structure of the problem at hand as best as we can.
+On the other hand, the method proposed above does allow us to take temporal confounders into account at all, and we are only limited by the extend of our domain knowledge, or to capture the relevant data, to identify and include the temporal confounders as exogenous feature variables in the machine learning model.
+
 \section{Example: Demand Forecasting}
@@
 We use a public dataset obtained from a Kaggle online competition focusing in estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual items for specific stores on specific days.
-For each demand forecast, Cyclic Boosting is used to predict the full probability density distribution of the expected demand at a granularity of (item, store, day) and use the methods described in sec.
+For each demand forecast, Cyclic Boosting is used to predict the full probability distribution of the expected demand at a granularity of (item, store, day) and use the methods described in sec.
 \ref{sec:pdfEvaluation} to evaluate the quality of the individual forecasts.
@@
 As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event.
-As a first step, we use Cyclic Boosting (as described in sec.
+As a first step, we use Cyclic Boosting as the supervised machine learning method (as described in sec.
 \ref{sec:CB}) to predict the mean of the distribution.
@@
 \noindent
-This model uses the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time-series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
+This model uses the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
 We also include various two-dimensional combinations of these features.
@@
 
-Unlike most state-of-the-art time series forecasting methods, we do not include lagged target information, for example via stacking of exponentially weighted moving average features, in our model.
-Although common in practice, including such variables makes the learning of exogenous effects, such as product promotions or the occurrence of special events, much harder.
-This is because when those variables are included, machine learning models tend to rely mainly on temporal confounding.
-Therefore, omitting these kind of variables improves the capability of the machine learning model to learn causal dependencies, which in turn improves the explainability of the model as well as the quality of the forecasts for mid- to long-term predictions and rare events.
-In order to capture recent trends that are not reflected in the exogenous features of the model, we apply an individual residual correction on each of the predictions of the machine learning model, which accounts for deviations between the exponentially weighted moving average (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past.
-We use the model to predict the expected demand two days into the future to reflect a realistic replenishment scenario, meaning that we use a target lag of two days for the training.
+\noindent
+As discussed in sec.
+\ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past.
+To reflect a realistic replenishment scenario, we use the model to predict the expected demand two days into the future, meaning that we use a lag of two days for the calculation of  both the target EWMA and the prediction EWMA.
 
@@
 We use the mean absolute deviation (MAD) as well as the mean squared error (MSE) as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean.
-These metrics do not take the shape of the underlying probability distribution into account, but only compare the predicted mean to the observed number of sales.
+These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales.
 Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.09$.
-The mean of the target, i.e.
-the observed sales, is $3.28$ for this period.
+The mean of the target, i.e., the observed sales, is $3.28$ for this period.
 \fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.
@@
 \includegraphics[width=8cm]{figs/ts_full}
-\caption{\label{fig:mean_prediction} Time series of mean predictions and sales in test period summed over all products and stores.}
+\caption{\label{fig:mean_prediction} Time series of final mean predictions as well as sales in test period summed over all products and stores.}
 \end{center}
 \end{figure}
+
+\noindent
+To visualize the operation principle of the residual correction method, \fig \ref{fig:ts_res} shows the time series of sales and machine learning (before residual correction) as well as final predictions (after residual correction) for item \texttt{FOODS\_3\_516} in store \texttt{CA\_3} from beginning of November 2015 to end of April 2016, including all days in the test period in 2016 as well as the last two months of the training period before.
+Hereby, the machine learning predictions in the training period are in-sample predictions, which are just used to calculate the residual correction.
+A sudden change of the sales time series at the turn of the year is visible, which is seemingly not anticipated by the machine learning model, as its predictions are clearly off in the test period in 2016.
+However, the residual correction is able to account for the resulting deviations between sales and machine learning predictions, which can be seen by the much better agreement of the final predictions.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=8cm]{figs/ts_item_16_store_2_res}
+\caption{\label{fig:ts_res} Time series of mean predictions, before (machine learning prediction) and after (final prediction) residual correction, as well as sales for a specific product-location combination.
+In addition to the test period in 2016, the last two months of the training period are shown as well, in order to better demonstrate the workings of the residual correction.}
+\end{center}
+\end{figure}
+
+\noindent
+To give an indication of the effectiveness of the two-step approach, i.e., machine learning model without lagged target information and residual correction, we get the following metrics for a model without residual correction, but including the EWMA of the target (grouped by \texttt{store\_id} and \texttt{item\_id}, with a lag of two days, and a recursive smoothing factor of $0.25$) as a feature: MAD: $1.70$ and MSE: $11.04$.
+The value of the smoothing factor here differs from the one used in the residual correction, because this stricter damping improved the evaluation metrics.
+Including the day of the week as an additional group by dimension in the EWMA feature worsened the metrics even more.
 
@@
 For example, the \texttt{events} feature group corresponds to the combination of all the factors of the different event features and the \texttt{dayofweek} feature group includes the one-dimensional feature \texttt{dayofweek} and its two-dimensional combinations with \texttt{store\_id} and \texttt{item\_id}.
-
-\noindent
-The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values.
-The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl.
-The variance prediction for this day is driven by two competing factors:  the Super Bowl event itself drives  the  higher dispersion parameter higher, and the high mean prediction drives the lower dispersion parameter lower.
+Furthermore, the values of the residual correction on the mean predictions are included, which correspond to the differences between the shown machine learning and final predictions.
 
@@
 \includegraphics[width=4cm]{figs/factors_ts_width}
-\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions) and sales (a), constituting factors of the mean predictions (b), and constituting factors of the variance predictions (c), for a specific product-location combination from beginning of February to end of April 2016.
-For both mean and variance predictions, the most granular constituting factors are aggregated, i.e.
-multiplied, for the sake of visualization.}
+\caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions), before (machine learning prediction) and after residual correction (final prediction), as well as sales (a), constituting factors of the mean predictions (including the factor of the residual correction) (b), and constituting factors of the variance predictions (c), for a specific product-location combination from beginning of February to end of April 2016.
+For both mean and variance predictions, the most granular constituting factors are aggregated, i.e., multiplied, for the sake of visualization.}
 \end{center}
 \end{figure}
+
+\noindent
+The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the inverse dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values.
+The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl.
+The variance prediction for this day is driven by two competing factors: the Super Bowl event itself drives the variance up, while the high mean prediction (as feature) drives it down.
 
@@
 This plot also illustrates the benefits of using the quantile profile plot, since we have seen that using more conventional approaches, such as \fig \ref{fig:mean_prediction} or even the lower part of \fig \ref{fig:cdf_demand}, the predictions appear reasonable, even when not relying on simple point metrics such as MAD or MSE.
-It is therefore paramount both to predict the full probability distribution instead of just a point estimate, such as the mean, as well as verifying a number of quantiles for each prediction to assess the quality of the prediction thoroughly.
+It is therefore paramount both to predict the full PDF instead of just a point estimate, such as the mean, as well as verifying a number of quantiles for each prediction to assess the quality of the prediction thoroughly.
 
@@
 Both from practical and theoretical perspectives, disentangling the forecast of demand from the operational decision making regarding order quantities has potentially significant advantages over an integrated approach using data to predict the resulting order quantities directly.
-When predicting future demand, both model-free and a model-based approaches can be used, where the model-based approach is generally more robust, as it is rooted in a theoretical understanding of the sales process.
+When predicting future demand in the form of individual probability distributions, both model-free and a model-based approaches can be used, where the model-based approach is generally more robust, as it is rooted in a theoretical understanding of the sales process.
 Compared to a model-free approach like quantile regression, the distributional assumption can drastically reduce the uncertainty of the resulting predictions.
-Using the Cyclic Boosting machine learning approach, the full probability density distribution can be predicted in a fully explainable way on the individual level.
+Using the Cyclic Boosting supervised machine learning approach, the full probability density function can be predicted in a fully explainable way on the individual level.
 This allows to extract not only simple point estimates, such as the expected mean demand, but all relevant quantiles of the distribution.
 
-Once the full probability density functions are forecasted, their evaluation poses significant challenges: Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted probability distributions.
+Once the full individual probability distributions are forecasted, their evaluation poses significant challenges: Common metrics, such as the mean absolute deviation or mean squared error, only take point estimates into account and are generally not suitable for the evaluation of predicted PDFs.
 Furthermore, since the distributions are predicted for individual events, statistical techniques aiming at the comparison of distributions using many events cannot be used in general.
@@
 
-% Authors must disclose all relationships or interests that 
-% could have direct or potential influence or impart bias on 
-% the work: 
+An additional challenge for demand forecasting in specific and time series forecasting in general is the handling of the detrimental effects of temporal confounding, especially on explainability and generalizability of the employed models.
+Using a two-step model consisting of a machine learning model, including all available exogenous variables as features but excluding any lagged target information, to control for all known temporal confounders, and a subsequent empiric residual correction to capture any information on recent trends missed in the ML model, temporal confounding can be avoided, or at least mitigated.
+
+% Authors must disclose all relationships or interests that could have direct or potential influence or impart bias on the work: 
 %
@@
 %
+%\noindent
+%Authors Felix Wick and Daniel Stemmer applied for a US patent ``Causal Factor Machine Learning with Residual Time Series Correction''.
+%
 %
@@
 %\subsection*{Code availability}
-%The details of the code are available from \cite{Wick2019}
+%The details of the code are available from the authors
 %
 %\subsection*{Authors' contributions}
-%F. Wick: Code examples and manuscript text, U. Kerzel: manuscript text, M. Hahn, M. Wolf, T. Singhal, M. Feindt: revision
+%F. Wick: Code examples and manuscript text, U. Kerzel: manuscript text, M. Hahn, M. Wolf, T. Singhal, M. Feindt, D. Stemmer, J. Ernst: revision
 
@@
 
-% Springer
-%\bibliographystyle{spmpsci}      % mathematics and physical sciences
 \bibliographystyle{ieeetr}
 
-%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% Appendices
-%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \appendix
+
+\section{Optimal Point Estimator for Quadratic Cost Function}
+\label{sec:CostQuantile}
+
+Operational decisions ultimately need to be based on a single number, for example, if we want to order new goods, in the end we need a single number representing the expected demand, include operational constraints such as lot sizes and delivery schedules from the manufacturer or wholesaler to arrive at the optimal order quantity for the next delivery period.
+If we predict the future demand as a complete probability density distribution, we need to determine the optimal point estimator that ``reduces'' the PDF to the single best quantity.
+In an ideal case, the full cost function relating point estimator $p_i$ to observed (true) value $t_i$: $C(p_i,t_i)$.
+If the cost function is fully specified, we can obtain the optimal point estimator $p$ by minimizing the cost function with respect to the predicted probability distribution $f(t)$, i.e., we require that the first derivative of the cost function vanishes.
+\begin{equation}
+	\frac{\partial E[C(p,t)]}{\partial p} = 0
+\end{equation}
+Since the variables are random variables, we need to include the expectation value $E[\cdot]$.
+A common metric to evaluate point estimators is the mean squared error (MSE) which is a quadratic cost function given by $C(p,t) = (p-t)^2$.
+We can use this to derive the optimal point estimator associated with this particular cost function:
+\begin{align*}
+0 &= \frac{\partial E[C(p,t)]}{\partial p}   \\
+  &= \frac{\partial \int_{-\infty}^\infty  f(t) (p-t)^2 dt}{\partial p}  \\
+ &= 2 \int_{-\infty}^\infty  f(t) (p-t) dt
+\end{align*}
+Using the normalization of the probability distribution, i.e., $\int_{-\infty}^\infty  f(t) dt = 1$, the equation can be written as 
+\begin{equation}
+p = \int_{-\infty}^\infty  t f(t) dt
+\end{equation}
+Therefore, the optimal point estimator associated with a quadratic cost function is the mean of the predicted probability distribution for the demand.
+
+In the case of the original newsvendor problem, underage and overage cost are linear. In the special case that they are identical, the cost function is given by the MAD, i.e., $C(p,t) = |p-t|$.
+We can derive the optimal point estimator in the following way, using the shorthand notation $\partial_p X$ for $\partial/\partial p X$
+\begin{align*}
+0 &= \partial_p E[E(p,t)]  \\
+&= \partial_p \int_{-\infty}^\infty f(t) |p-t| dt \\
+&= \partial_p \left ( \int_{-\infty}^{x_p} f(t) |p-t| dt +  \int_{x_p}^\infty f(t) |p-t| dt \right ), \\
+\end{align*}
+where we have split the integration at some intermediate point $x_q$.
+This allows us to resolve the absolute value if we choose $x_q$ such that $t$ is always smaller than $p$ in the first integral and the other way around in the second.
+Hence,
+\begin{equation}
+0 = \partial_p \left ( \int_{-\infty}^\infty f(t) (p-t) dt -  \int_{-\infty}^\infty f(t) (p-t) dt \right ) \\
+\end{equation}
+We then execute the partial derivative with respect to $p$ and obtain:
+\begin{equation}
+0 = -  \int_{-\infty}^{x_p} f(t) dt + \int_{x_p}^\infty f(t)  dt
+\end{equation}
+which we can write as:
+\begin{equation}
+ \int_{-\infty}^{x_p} f(t) dt = \int_{x_p}^\infty f(t)  dt
+\end{equation}
+Therefore, the integral from $-\infty$ to the point $x_p$ is the same as from $x_p$ to $\infty$, which means that the point $x_p$ splits the integral in the interval $(-\infty, \infty)$ into two halves.
+This is the definition of the median and therefore a linear cost function with equal over- and underage costs are linked to the median of the predicted probability density function as optimal point estimator.
+If the underage and overage cost differ, effectively the slope of the two linear segments are no longer the same.
+This means that the resulting optimal point estimator is no longer the median but a different quantile.
 

### 2020-12-20 code/walmart_demand.py

@@ -101,6 +101,8 @@ def plot_factors_ts(df, filename):
     df['price'].plot(label="price")
     df['seasonality'].plot(label="seasonality")
     df['events'].plot(label="events")
+    df['correction_factor'].plot(style='b:', label="correction")
+#    df['sales_ewma'].plot(label="sales_ewma")
 
 #    df['check'].plot(label="check")
 #    df['y'].plot(label="y")
@@ -254,17 +256,21 @@ def plot_invquants_examples():
     plt.savefig('plots/invquant_example.pdf')
 
 
-def plot_timeseries(df, suffix, title='', textbox=None, errorband=False):
+def plot_timeseries(df, suffix, title='', textbox=None, errorband=False, split_line=False):
     plt.figure()
     ax = plt.axes()
     df.index = df['date']
     df['y'].plot(style='r', label="sales")
-    df['yhat_mean'].plot(style='b', label="prediction")
+    df['yhat_mean'].plot(style='b-.', label="final prediction")
+    if suffix != 'full':
+        df['yhat_mean_causal'].plot(style='b:', label="ML prediction")
     if errorband:
         plt.fill_between(df.index, df['yhat_mean'] - np.sqrt(df['yhat_var']), df['yhat_mean'] + np.sqrt(df['yhat_var']), alpha=0.2)
     plt.legend(fontsize=15)
 #    plt.title(title)
     plt.ylabel("sum", fontsize=15)
+    if split_line:
+        plt.vlines('2016-01-01', ymin=0, ymax=80, linestyles="dashed")
     if textbox is not None:
         plt.text(-0.1, -0.22, textbox, fontsize=15, transform=ax.transAxes)
     plt.tight_layout()
@@ -272,25 +278,37 @@ def plot_timeseries(df, suffix, title='', textbox=None, errorband=False):
 
 
 def plotting(df, suffix=''):
-    df = df[['y', 'yhat_mean', 'yhat_var', 'item_id', 'store_id', 'date']]
+    df = df[['y', 'yhat_mean', 'yhat_var', 'item_id', 'store_id', 'date', 'correction_factor']]
+    df.loc[df['date'] < '2016-01-01', 'correction_factor'] = 1
+    df['yhat_mean_causal'] = df['yhat_mean'] / df['correction_factor']
[…]

### 2020-12-21 LaTex/paper.tex
several changes after review

@@
 Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step.
+Another possibility to model the underlying data distribution without assumption of a fixed distribution class is via conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.
 
@@
 
-A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted PDF $f(x)$ is correctly calibrated.
-This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for five different cases.
-If both the choice of the model and the model parameters are estimated correctly, we would expect the uniform distribution.
+A histogram of the actually observed CDF values for each individual PDF prediction (see \fig \ref{fig:pdf_example} for an example) is therefore expected to be uniformly distributed in $[0,1]$, if the predicted PDF $f(x)$ is correctly calibrated, i.e., if both the choice of the model and the model parameters are estimated correctly.
 If the mean or the variance are not estimated correctly, the resulting distribution will show a distinct deviation from this uniform behavior.
+This is illustrated in \fig \ref{fig:cdf_histos}, which shows the distribution of observed CDF values for four different cases.
 
@@
 \begin{center}
-\includegraphics[scale=0.25]{figs/cdf_truth_uniform}
 \includegraphics[scale=0.25]{figs/cdf_truth_broad}
@@
 \includegraphics[scale=0.25]{figs/cdf_truth_under}
-\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution:
-correct prediction (a: ``uniform''), variance overestimated (b: ``broad''), variance underestimated (c: ``narrow''), mean overestimated (d: ``over''), mean underestimated (e: ``under'')}
+\caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution (dashed horizontal line in each of the plots): variance overestimated (a: ``broad''), variance underestimated (b: ``narrow''), mean overestimated (c: ``over''), mean underestimated (d: ``under'')}
 \end{center}
@@
 As discussed in sec.
-\ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with a recursive smoothing factor of $0.15$) of the predictions and targets of each product-location combination over the corresponding past.
+\ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with $\alpha=0.15$) of the predictions and targets of each product-location combination over the corresponding past.
+Empirically, a value of $\alpha=0.15$ gives good results here, though in a practical application, this parameter would have to be optimized using methods such as cross-validation according to a metric that reflects the objective of the business strategy.
 To reflect a realistic replenishment scenario, we use the model to predict the expected demand two days into the future, meaning that we use a lag of two days for the calculation of both the target EWMA and the prediction EWMA.
@@
 These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales.
-Using the independent test data, we obtain the following metrics: MAD: $1.65$ and MSE: $10.09$.
+Using the independent test data, we obtain the metrics as stated in the top line of \tab \ref{tab:mad_mse}.
 The mean of the target, i.e., the observed sales, is $3.28$ for this period.
 \fig \ref{fig:mean_prediction} shows the time series of both predictions and sales summed over all 100 products and 10 stores during the test period.
+
+\begin{table}[h!]
+\begin{center}
+\caption{Accuracy metrics for the predicted mean values for three different model setups: (a) ML model without lagged target information but subsequent residual correction, (b) ML model including lagged target information by means of two EMWA features without residual correction, and (c) ML model from (b) with residual correction.}
+\label{tab:mad_mse}
+\begin{tabular}{p{4cm}|c|c}
+ & \textbf{MAD} & \textbf{MSE} \\
+\hline
+\textbf{no EWMA features,\newline residual correction} & 1.65 & 10.09 \\
+\hline
+\textbf{EWMA features,\newline no residual correction} & 1.69 & 10.84 \\
+\hline
+\textbf{EWMA features,\newline residual correction} & 1.68 & 10.41
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 \caption{\label{fig:ts_res} Time series of mean predictions, before (machine learning prediction) and after (final prediction) residual correction, as well as sales for a specific product-location combination.
-In addition to the test period in 2016, the last two months of the training period are shown as well, in order to better demonstrate the workings of the residual correction.}
+In addition to the test period in 2016, the last two months of the training period are shown as well (the dashed vertical line marking the transition from training to test period), in order to better demonstrate the workings of the residual correction.}
 \end{center}
@@
 \noindent
-To give an indication of the effectiveness of the two-step approach, i.e., machine learning model without lagged target information and residual correction, we get the following metrics for a model without residual correction, but including the EWMA of the target (grouped by \texttt{store\_id} and \texttt{item\_id}, with a lag of two days, and a recursive smoothing factor of $0.25$) as a feature: MAD: $1.70$ and MSE: $11.04$.
-The value of the smoothing factor here differs from the one used in the residual correction, because this stricter damping improved the evaluation metrics.
-Including the day of the week as an additional group by dimension in the EWMA feature worsened the metrics even more.
+To give an indication of the effectiveness of the two-step approach, i.e., a machine learning model without lagged target information and residual correction as outlined in sec.
+\ref{sec:ts_confounding}, we compare its results (for the same training and test setup) with a model following the common practice of including the lagged target information as a variety of EMWA variables used as features in \tab \ref{tab:mad_mse}.
+For this, we include two EWMA features in the machine learning model, one grouped by \texttt{store\_id} and \texttt{item\_id} (with a lag of two days and $\alpha=0.25$), and one grouped by \texttt{store\_id}, \texttt{item\_id}, and the day of the week (with a lag of one week and $\alpha=0.05$), where the values for $\alpha$ are again found empirically.
+In \tab \ref{tab:mad_mse}, we also show the results after applying the same residual correction as before for the machine learning model without lagged target information on top of this model including the two EWMA features.
+The crucial point in this is that when applying the same residual correction method to both machine learning models with and without including lagged target information as EWMA features, the one without yields better results.
+Following our reasoning in sec.
+\ref{sec:ts_confounding}, this is due to detrimental temporal confounding effects disturbing the learning of the causal dependencies between the exogenous features and the target, which in turn leads to a worse generalizability of the machine learning model.
 

### 2020-12-21 code/walmart_demand.py

@@ -181,6 +181,7 @@ def plot_cdf_truth(cdf_truth, suffix, ordering=''):
     plt.xlabel("CDF values", fontsize=15)
     plt.ylabel("count", fontsize=15)
     plt.text(-0.15, -0.15, ordering, fontsize=15, transform=ax.transAxes)
+    plt.hlines(100000./30, xmin=0, xmax=1, linestyles ="dashed")
     plt.tight_layout(rect=(0,0,1,0.99))
     plt.savefig('plots/cdf_truth_' + suffix + '.pdf')
 
@@ -287,20 +288,20 @@ def plotting(df, suffix=''):
     ts_data = predictions.groupby(['date'])[['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal']].sum().reset_index()
     plot_timeseries(ts_data, 'full' + suffix, 'all')
 
-    predictions_grouped = df.groupby(['store_id'])
-    for name, group in predictions_grouped:
-        ts_data = group.groupby(['date'])['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal'].sum().reset_index()
-        plot_timeseries(ts_data, 'store_' + str(name) + suffix)
+    # predictions_grouped = df.groupby(['store_id'])
+    # for name, group in predictions_grouped:
+    #     ts_data = group.groupby(['date'])['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal'].sum().reset_index()
+    #     plot_timeseries(ts_data, 'store_' + str(name) + suffix)
 
     # predictions_grouped = df.groupby(['item_id'])
     # for name, group in predictions_grouped:
     #     ts_data = group.groupby(['date'])['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal'].sum().reset_index()
     #     plot_timeseries(ts_data, 'item_' + str(name) + suffix)
 
-    predictions_grouped = df.groupby(['item_id', 'store_id'])
-    for name, group in predictions_grouped:
-        ts_data = group.groupby(['date'])['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal'].sum().reset_index()
-        plot_timeseries(ts_data, 'item_' + str(name) + suffix)
+    # predictions_grouped = df.groupby(['item_id', 'store_id'])
[…]

### 2020-12-21 LaTex/paper.tex
nicer table

@@
 \begin{center}
-\caption{Accuracy metrics for the predicted mean values for three different model setups: (a) ML model without lagged target information but subsequent residual correction, (b) ML model including lagged target information by means of two EMWA features without residual correction, and (c) ML model from (b) with residual correction.}
+\caption{Accuracy metrics for the predicted mean values for three different model setups: a) ML model without lagged target information but subsequent residual correction, b) ML model including lagged target information by means of two EMWA features without residual correction, and c) ML model from b) with residual correction.}
 \label{tab:mad_mse}
-\begin{tabular}{p{4cm}|c|c}
- & \textbf{MAD} & \textbf{MSE} \\
+\begin{tabular}{c|p{4cm}|c|c}
+ & & \textbf{MAD} & \textbf{MSE} \\
 \hline
-\textbf{no EWMA features,\newline residual correction} & 1.65 & 10.09 \\
+a) & no EWMA features,\newline residual correction & 1.65 & 10.09 \\
 \hline
-\textbf{EWMA features,\newline no residual correction} & 1.69 & 10.84 \\
+b) & EWMA features,\newline no residual correction & 1.69 & 10.84 \\
 \hline
-\textbf{EWMA features,\newline residual correction} & 1.68 & 10.41
+c) & EWMA features,\newline residual correction & 1.68 & 10.41
 \end{tabular}

### 2020-12-23 LaTex/paper.tex
final review before resubmit

@@
 Demand forecasting is a central component of the replenishment process for retailers, as it provides crucial input for subsequent decision making like ordering processes.
-In contrast to point estimates, such as the conditional mean of the underlying probability distribution or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand.
+In contrast to point estimates, such as the conditional mean of the underlying probability distribution, or confidence intervals, forecasting complete probability density functions allows to investigate the impact on operational metrics, which are important to define the business strategy, over the full range of the expected demand.
 Whereas metrics evaluating point estimates are widely used, methods for assessing the accuracy of predicted distributions are rare, and this work proposes new techniques for both qualitative and quantitative evaluation methods.
@@
 One of the main business operations of retailers is to ensure that sufficient goods are available in the store for customers to buy.
-This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler.
+This entails two parts: In a first step, we need to estimate the future customer demand to be able to judge how many goods we will need to order from the supplier or wholesaler, given the current inventory level.
 In the second step, aimed at inventory control, we need to place the actual orders, which is further complicated by the fact that multiple operational constraints, such as lot-sizes or delivery schedules, need to be included in the final ordering decision.
@@
 Assuming linear overage ($h$) and underage ($b$) cost, the optimal quantile $q_{\mathrm{opt}} = {b}/{(b+h})$ of a known probability density function (PDF) describing demand as random variable can be calculated exactly.
-This quantile then corresponds to the point estimator corresponding to the numerical value of the demand prediction that can be used to determine the optimal order quantity.
+This quantile then corresponds to the numerical value of the demand PDF prediction that can be used to determine the optimal order quantity.
 
@@
 Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time.
-However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF.
-Appendix \ref{sec:CostQuantile} demonstrates how the optimal point estimator can be calculated using the full PDF.
-For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, the calculations are straight-forward.
+However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF, what is demonstrated in Appendix \ref{sec:CostQuantile}.
+For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, corresponding to the mean absolute deviation (MAD), the calculations are straight-forward.
 If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches.
@@
 This means that, while we can generally assume that demand follows a specific type of PDF, its parameters are unique to the instance for which an estimate is required.
-For example, the PDF describing the demand of a particular product is specific to this product, date, and retail location for which the forecast is made.
+For example, the PDF describing the demand of a particular product-location-date combination is specific to the product, date, and retail location for which the forecast is made.
 Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc.
-Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDF predictions rather than globally adjusted parameters for groups of SKUs or locations or sales days.
+Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDFs rather than globally adjusted parameters for groups of SKUs or locations or sales days.
 
@@
 It is therefore tempting to use established time series regression models to predict the future values of the expected demand.
-However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed variables $Y$ used in the time series, say $Y_{t}$ and $Y_{t-\tau}$ with lag $\tau$.
+However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed time series variable $Y$, say $Y_{t}$ and $Y_{t-\mathrm{lag}}$, where lag denotes a specific time difference between the two steps.
 In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion.
-This dependency creates a backdoor path between $Y_{t-\tau}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$.
-This is particularly important for the case of demand forecasting, since customer demand is not a natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc.
-If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelations, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable.
+This dependency creates a backdoor path between $Y_{t-\mathrm{lag}}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$.
+This is particularly important for the case of demand forecasting, since customer demand is not a mere natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc.
+If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelation, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable.
 Moreover, due to this disguising of actual causal dependencies, reliance on temporal confounding is also detrimental in terms of generalizability of the forecasting model.
@@
 % ML
-Demand forecasting for retailers often requires to predict millions of product-location combinations for each SKU each day.
+Demand forecasting for retailers often requires to predict millions of product-location combinations each day.
 Univariate time series forecasting methods operate on each of these many individual time series separately.
@@
 	\item {Avoid temporal confounding in the model used for the prediction of the future demand.}
-	\item {Evaluate that the functional form of the PDF is correct.}
-	\item {Evaluate that the shape of the PDF is correct and that each part of the PDF models the observed data appropriately.}
+	\item {Evaluate that each part of the PDF models the observed data appropriately.}
 	\item {Ideally, each prediction is fully explainable, enabling operational experts to exactly understand what contributed to it.}
@@
 \ref{sec:LitRev}.
-We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean of variance in sec.
+We then describe our method to predict individual negative binomial PDFs using  a parametric approach including two distinct machine learning models for mean and variance in sec.
 \ref{sec:pdfEstimation}.
@@
 See also \cite{deGooijerHyndman2006} for a comprehensive overview over traditional time series forecasting methods.
-Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality or external factors in various ways, depending on the exact modeling approach taken.
-The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $y_t$ that is to be forecasted with some lag $\tau$ between subsequent events.
+Essentially, these methods extrapolate the future values by building some form of regression model based on past observations, i.e., $y_{t+1} = f(y_t, y_{t-1}, \ldots)$, introducing dedicated functional components for trends, seasonality, or external factors in various ways, depending on the exact modeling approach taken.
+The main underlying concept behind these approaches is to exploit the autocorrelation between values of the variable $Y$ at different times $t$.
 Intuitively, we assume that  the sequence of events observed so far is a good predictor of future events.
@@
 An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}.
-For the case of sequence samples, deep learning methods such recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used.
+For the case of sequence samples, deep learning methods such as recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used.
 These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events.
@@
 Brodersen \etal \cite{Brodersen2015,GoogleCausalImpact} propose to use Bayesian structural time series models to estimate the causal effect of interventions on a time series by creating a counterfactual prediction that describes the behavior of the time series had the intervention not taken place.
-While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state at $Y_{t-1}$ to $Y_t$.
+While structural time series equations allow to model components such as trend, seasonality, and others explicitly, the fundamental form of the state equation is $Y_t = G_t Y_{t-1} + \epsilon_t$, where $\epsilon_t \sim \mathcal{N}(0,\sigma^2)$ is a white noise term and $G_t$ some matrix that describes the evolution from the state $Y_{t-1}$ to $Y_t$.
 Therefore, great care must be taken to avoid temporal confounding when setting up the states for structural time series models.
@@
 % quantile regression
-Generally, quantile regression \cite{koenker2001,wen2017} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
+Generally, quantile regression \cite{koenker2001} can be implemented in various frameworks and used to estimate a range of quantiles for each predicted distribution, from which an empirical probability distribution can be interpolated.
 Using a dedicated neural network \cite{Feindt2006190}, either the full PDF or a defined range of quantiles can be calculated directly from the data for each individual prediction without assuming an underlying model.
@@
 Using an attention-based architecture, the Temporal Fusion Transformer (TFT) \cite{lim2019temporal} can also predict a set of specified quantiles at each predicted time step.
-Another possibility to model the underlying data distribution without assumption of a fixed distribution class is via conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.
+Another possibility to model the underlying data distribution without assumption of a fixed distribution class is the usage of conditional normalizing flows \cite{rezende2016variational,timeseriesflow}, which aim to transform a simple initial density into a more complex one by applying a sequence of invertible transformations.
 
@@
 Empirically, we can determine the best fitting distribution from data \cite{adan1995}.
-However, given the stochastic nature of the demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes.
+However, given the stochastic nature of demand, such an empirically determined distribution is not expected to be stable and prone to sudden changes.
 Instead, the choice of the demand distribution should be motivated by theoretic considerations.
-The discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}.
+Discrete demand is typically modeled as a negative binomial distribution (NBD), also known as Gamma-Poisson distribution \cite{Ehrenberg1959, Ehrenberg1967, Ehrenberg1972, Chatfield1973, Schmittlein_1985}.
 This distribution arises if the Poisson parameter $\mu$ is a random variable itself, which follows a Gamma distribution.
@@
 
-In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBD using two Cyclic Boosting models.
+In the following, after a brief recap of the fundamental ideas of Cyclic Boosting, we describe a method to predict mean and variance for individual NBDs using two Cyclic Boosting models.
 
@@
 The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
-Single data records, for example, the sales corresponding to a specific product-location-date combination along with all relevant features, are indexed by $i$.
+Single data records, for example, the sales corresponding to a specific product-location-date combination (meaning the sales record of a specific item sold on a specific day at a specific sales location) along with all relevant features, are indexed by $i$.
 The individual predictions $\hat{y_i}$  can then be calculated as:
@@
 \end{equation}
-Here,  $g$ is a factor that is multiplied to $f_{t-1}$ in each iteration.
+Here,  $g$ is a factor that is multiplied to the corresponding $f_{t-1}$ in each iteration.
 The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
@@
 For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
-\item{Quit when stopping criteria, e.g., the maximum number of iterations or no further improvement of an error metric such as the mean absolute deviation (MAD) or mean squared error (MSE), are met at the end of a full feature cycle.}
+\item{Quit when stopping criteria, e.g., the maximum number of iterations or no further improvement of an error metric such as the MAD or MSE, are met at the end of a full feature cycle.}
 \end{enumerate}
@@
 In the previous section, the general Cyclic Boosting algorithm was used to estimate the mean of the NBD model.
-In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm as follows: When looking at the demand of individual product-location-date combinations (meaning the sales record of a specific item sold on a specific day at a specific sales location), the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
+In order to predict the variance of the NBD model (associated with the mean predicted before), we modify the algorithm as follows: When looking at the demand of individual product-location-date combinations, the target variable $y$ has the values $y = 0, 1, 2, ...$ and the NBD model can be parameterized as in \cite{hilbe2011negative}:
 \begin{equation} \label{eqn:nbinom}
@@
 
-The estimate of the dispersion parameter $\hat{r}$ can then be calculated by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, which is expressed as negative log-likelihood function of a negative binomial distribution.
-Using Cyclic Boosting, the minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained in the previous step (described in sec.
+The estimate of the dispersion parameter $\hat{r}$ can then be calculated by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, which is expressed as negative log-likelihood function of an NBD.
+The minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained in the previous step (described in sec.
 \ref{sec:CB}).
@@
 Like in sec.
-\ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$ and again values deviating from $f^k_j=1$ can be used to explain the relative importance of a specific feature contributing to individual predictions.
+\ref{sec:CB}, the model parameters $f^k_j$ correspond to factors with values in $[0, \infty]$, and again values deviating from $f^k_j=1$ can be used to explain the relative importance of a specific feature contributing to individual predictions.
 Note that the structure of \eqn \eqref{eqn:r} can be interpreted as inverse of a logit link function in the same way as explained in \cite{Wick2019} when Cyclic Boosting is used for classification tasks.
@@
 \ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example, using a random search.
-All other advantages of Cyclic Boosting, like for example individual explainability of predictions, remain valid for its negative binomial width mode.
+All other advantages of Cyclic Boosting, like individual explainability of predictions, remain valid for its negative binomial width mode.
 
 Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ using \eqn \eqref{eqn:variance_r}.
-Using  the individual predicted mean $\hat{\mu_i}$ from the first step, the model is fully specified for each individual prediction $i$.
+Using  the individual predicted mean $\hat{\mu_i}$ from the first step, the NBD model is fully specified for each individual prediction $i$.
 
@@
 Many statistical and most machine learning methods do not provide a full PDF as result.
-Instead, these methods typically predict a single numerical quantity (denoted by $\hat{y}$) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the mean squared error (MSE), mean absolute deviation (MAD) or others.
+Instead, these methods typically predict a single numerical quantity (denoted by $\hat{y}$) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the MSE, MAD, or others.
 In the setting of a retailer, the observed quantity is the sales of individual products and most machine learning approaches would then predict a single number as a direct estimate of the sales.
@@
 Instead, it is imperative to predict the full PDF for each prediction to be able to optimize the subsequent operational decision.
-Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or at least qualitative evaluation methods to assess whether the full distribution has been forecast correctly, in particular in the tails of the distribution.
+Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or even qualitative evaluation methods to assess whether the full distribution has been forecasted correctly, in particular in the tails of the distribution.
+
+In the simplest case, we only have one model with one set of model parameters to cover all predictions.
+In this case, the evaluation of the full PDF is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as an NBD with predicted parameters, that is used for all observations.
+Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
+
+In practical applications however, we have a large number of effective prediction models, since although we always use the same model parameterization, such as the NBD, its parameters have to be determined at the required level of granularity.
+For example, for daily orders, we need to predict the parameters of the NBD for each location, sales day, and product.
+Unlike the simple case discussed above, where we had many observations to compare the prediction model to, we now have just a single observation per prediction, meaning that we cannot use statistical tests directly.
 
 For an estimation of the determining parameters of an assumed functional form for the PDF, assessing the correctness of the PDF model output refers to the evaluation of the accuracy of the prediction of the different parameters.
-In the case of the negative binomial distribution used in this work, we have to verify that mean and variance are determined accurately, as well as checking that the choice of the underlying model can describe the observed data.
+In the case of the NBD used in this work, we have to verify that mean and variance are determined accurately, as well as checking that the choice of the underlying model can describe the observed data.
 
@@
 
-\subsection{Qualitative Evaluation of PDF Predictions}
-
-In the simplest case, we only have one model with one set of model parameters to cover all predictions.
-In this case, the evaluation of the full PDF is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as a negative binomial with predicted parameters, that is used for all observations.
-Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
-
-In practical applications however, we have a large number of effective prediction models, since although we always use the same model parameterization, such as the negative binomial distribution, its parameters have to be determined at the required level of granularity.
-For example, for daily orders, we need to predict the parameters of the negative binomial distribution for each location, sales day, and product.
-Unlike the simple case discussed above, where we had many observations to compare the prediction model to, we now have just a single observation per prediction, meaning that we cannot use statistical tests directly.
-
-\subsubsection{Histogram of CDF Observations}
+\subsection{Histogram of CDF Observations}
 \label{sec:cdf_histo}
@@
 Here, $F_X(x)$ is the CDF with $\lim_{x \to -\infty}F_X(x) = 0$ and $\lim_{x \to \infty}F_x(x) = 1$.
-The cumulative distribution describes the probability that a the variable has a value smaller than $x$ and intuitively represents the area under $f(x^\prime)$ up to a point $x$.
+The cumulative distribution describes the probability that the variable has a value smaller than $x$, and intuitively represents the area under $f(x^\prime)$ up to a point $x$.
 
 If the CDF is continuous and strictly increasing, then the inverse of the CDF, $F^{-1}(y)$, exists and is a unique real-valued number $x$ for each $y \in [0,1]$, so that we can write $F(x) = y$.
-The inverse of the CDF is also called the quantile function, because we can define the quantile $\tau$ of the PDF $f(x)$ as:
+The inverse of the CDF is also called the quantile function, because we can define the quantile $q$ of the PDF $f(x)$ as:
 
 \begin{equation}
-Q_\tau = F^{-1}(\tau)
+Q_q = F^{-1}(q)
 \end{equation}
 
-Using the example of the normal distribution with $\mathcal{N}(0,1)$ as shown in \fig \ref{fig:PdfCdf}, we can identify the median ($\tau = 0.5$) by first looking at the CDF in the lower part of the figure, look at $y=0.5$ on the $y$-axis and then identify the point on the $x$ axis for both the PDF $f(x)$ and the CDF $F(x)$ that correspond to the quantile $\tau$.
+Using the example of the normal distribution with $\mathcal{N}(0,1)$ as shown in \fig \ref{fig:PdfCdf}, we can identify the median ($q = 0.5$) by first looking at the CDF in the lower part of the figure, look at the value $0.5$ on the $y$ axis and then identify the point on the $x$ axis for both the PDF $f(x)$ and the CDF $F(x)$ that corresponds to the quantile $q$.
 In the case of the normal distribution, this is of course the central value at zero.
@@
 \includegraphics[width=8cm]{figs/PdfCdf}
-\caption{\label{fig:PdfCdf} PDF $f(x)$ and CDF of a normal distribution $F(x)$.}
+\caption{\label{fig:PdfCdf} PDF $f(x)$ and CDF $F(x)$ of a normal distribution.}
 \end{center}
@@
 We can then interpret the CDF as a new variable $s = F(t)$, meaning that $F$ becomes a transformation that maps $t$ to $s$, i.e., $F:t \to s$.
-Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$ and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF.
+Accordingly,  $\lim_{t \to -\infty}s(t) = 0$ and $\lim_{t \to \infty}s(t) = 1$, and $s$ can be intuitively interpreted as the fraction of the distribution of $t$ with values smaller than $t$ from the definition of the CDF.
 This implies that the PDF of $s$, $g(s)$, is constant in the interval $s \in [0,1]$ in which $s$ is defined, and $s$ can be interpreted as the CDF of its own PDF:
@@
 
-In case of discrete probability functions, such as the negative binomial function, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g., \cite[p.
+In case of discrete probability distributions, such as the NBD, the same argument still holds, but the definition of the quantile function is replaced by the generalized inverse: $F^{-1}(y) = \mathrm{\inf \left \{x : F(x)>y\right  \} }$ for $y \in [0,1]$, see e.g., \cite[p.
 54]{casella2002statistical}.
@@
 
-\noindent
 It should be noted that the method requires a sufficient sample size, as a too low number of observations of course leads to a discretization bias.
 
-\subsubsection{Inverse Quantile Profile Plot}
+\subsection{Inverse Quantile Profile Plot}
+\label{sec:invquant_plot}
 
@@
 \ref{sec:cdf_histo} by comparing (in the sense of higher or lower) the CDF values of the observed events, i.e., the sales records, with specified quantiles.
-In order to do so, we start again from the predicted values for the mean and variance from which a negative binomial distribution is constructed for each prediction, for example, for the predicted demand for a single product sold on a single day in a given sales location.
+In order to do so, we start again from the predicted values for the mean and variance from which an NBD is constructed for each prediction, for example, for the predicted demand for a specific product sold on a single day in a given sales location.
 Each of these predicted negative binomial PDFs is then transformed to its CDF.
@@
 
-Then we compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample.
+We then compare the actual observed sales value (corrected for censored data if necessary) to different quantiles of the corresponding predicted distribution for each data record and average over a larger data sample.
 For example, if we wanted to check that the median of the distribution, corresponding to the quantile $0.5$, is predicted correctly by the machine learning model, we would compare the value $0.5$ to the ratio of CDF values (again randomly chosen from the corresponding range of CDF values for discrete target values, as described in sec.
 \ref{sec:cdf_histo}) of observed sales records being lower/higher than $0.5$.
-In other words, in case of the median, 50 percent of the ex post observed target values should be observed below the median of the corresponding individual predicted PDF and 50 percent above.
+In other words, in case of the median, 50 percent of the ex post observed target values should be observed below the median of the corresponding individually predicted PDF and 50 percent above.
 
-In order to judge whether the overall shape of the predicted distributions is predicted correctly, we repeat this procedure for a range of quantiles, for example $q = 0.1, 0.2, \ldots, 0.9$.
-However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say $q = 0.95$ and $q = 0.99$.
+In order to judge whether the overall shape of the predicted distributions is estimated correctly, we repeat this procedure for a range of quantiles, for example, $q = 0.1, 0.2, \ldots, 0.9$.
+However, we are free to choose which quantiles to look at, and in specific situations it might be advisable to look at the tails of the distribution in more detail, to make sure that even relatively rare events are estimated correctly by the machine learning algorithm, and add more quantiles for comparison in the region between, say, $q = 0.95$ and $q = 0.99$.
 In the following, we call this method {\em inverse quantile profile plot}.
@@
 
-In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and compare to a uniform distribution (described in sec.
+In the same way as the method of filling all observed CDF values of the individual PDF predictions in a histogram and comparing to a uniform distribution (described in sec.
 \ref{sec:cdf_histo}), inverse quantile profile plots do not work on individual PDF predictions but require a statistical population.
 However, calculating the fractions of the inverse quantile profile plots globally, i.e., over all samples in the data set, might not reveal certain deficiencies for a subset, e.g., a specific store or product group in the example of retail demand forecasts.
-Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$-axis of the inverse quantile profile plot can be any variable of the data set at hand known at prediction time.
+Therefore, we combine the approach described above with the method of profile plots, where the quantity on the $x$ axis of the inverse quantile profile plot can be any variable of the data set at hand known at prediction time.
 
@@
 \ref{sec:cdf_histo}: First, by explicitly visualizing several different quantiles, the inverse quantile profile plot reveals which part of the predicted PDF, for example the tails, are particularly problematic.
-Second, by showing the dependency from the (arbitrary) variable on the $x$-axis, the inverse quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g., specific categories, of that variable.
+Second, by showing the dependency from the (arbitrary) variable on the $x$ axis, the inverse quantile profile plot reveals deviations of the predicted PDF from the actuals for different parts, e.g., specific categories, of that variable.
 Two examples for this can be found in Figs.
-\ref{fig:invquant_mean} and \ref{fig:invquant_dayofweek} in the next section.
+\ref{fig:invquant_dayofweek} and \ref{fig:invquant_mean} in the next section.
 
@@
 The methods discussed so far allow a detailed qualitative evaluation of PDF predictions.
-In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two PDFs to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
-Several approaches that measure the difference between two PDFs are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius, or the energy distance \cite{SZEKELY20131249}.
+In order to also quantify the quality of the PDF predictions, we use a metric assessing the difference between two probability distributions to compare the histogram of CDF observations of the predicted PDFs with the expected uniform distribution, and define an accuracy measure in the range between 0 and 1, which takes the value of 1 when both distributions agree perfectly.
+Several approaches that measure the difference between two probability distributions are suggested in the literature, such as the first Wasserstein distance \cite{olkin1982}, also known as earth mover distance (EMD), the Kullback-Leibler divergence \cite{kullback1951}, also known as relative entropy, the Jensen-Shannon divergence \cite{dagan1997}, also known as information radius, or the energy distance \cite{SZEKELY20131249}.
 
@@
 Many time series forecasting methods exploit the autocorrelation between the values of the variable of interest at different times.
-Depending on the exact nature of the problem we consider in these approaches, different lags between the variable $Y$ describing the demand play an important role.
-For example, a weekly repeating pattern would lead to a strong autocorrelation for a lag of $\tau=7$ (days), meaning that the values of the variable $Y$ at $y_t$ and $y_{t-\tau=7}$ would be strongly autocorrelated.
-However, this autocorrelation between different times or lags of the target variable may not be a true causal dependency but a spurious correlation due to one or more temporal confounders.
-For example, if an exogenous variable has a direct causal impact at the variable at time, say,  $Y_{t-1}$, this external variable will often also have an impact at time $Y_t$.
+Depending on the exact nature of the problem we consider in these approaches, different lags between the variable $Y$, describing the demand, play an important role.
+For example, a weekly repeating pattern would lead to a strong autocorrelation for a lag of $7$ (days), meaning that the values $y_t$ and $y_{t-7}$ of the variable $Y$ would be strongly autocorrelated.
+However, this autocorrelation between different times of the target variable may not be a true causal dependency, but a spurious correlation due to one or more temporal confounders.
+For example, if an exogenous variable has a direct causal impact on the variable $Y$ at time, say, $t-1$, this external variable will often also have an impact at time $t$.
 Intuitively, if the variable $Y$ was influenced yesterday by some external event, there is a high chance that we can observe this impact today as well.
@@
 This in turn means that there is now a spurious (auto)correlation between $Y_{t-1}$ and $Y_t$, indicated by the dashed line in \fig \ref{fig:temporal_confounding}.
-Therefore, if we do not control for the temporal confounder $X$, but use the autocorrelation between $y_{t-1}$ and $y_t$ to forecast the next value $y_{t+1}$, we rely on a spurious correlation between $Y_{t-1}$ and $Y_t$ that is not ``real'' in the sense that it results from a causal influence of $Y_{t-1}$ on $Y_t$, but originates from the backdoor path via $X_{t, t-1}$.
+Therefore, if we do not control for the temporal confounder $X$, but use the autocorrelation between $y_{t-1}$ and $y_t$ to forecast the next value $y_{t+1}$, we rely on a spurious correlation between $Y_{t-1}$ and $Y_t$ that is not ``real'' in the sense that it results from a causal influence of $Y_{t-1}$ on $Y_t$, but originates from the backdoor path via $X_{t-1, t}$.
 
@@
 \includegraphics[width=4cm]{figs/temporal_confounding}
-\caption{\label{fig:temporal_confounding} Temporal confounding by an exogenous variable $X_{t-1, t}$ stable over the time steps $t-1$ and $t$, creating a non-causal autocorrelation between $Y_{t-1}$ and $Y_t$.}
+\caption{\label{fig:temporal_confounding} Temporal confounding by an exogenous variable $X_{t-1, t}$ stable over the time steps $t-1$ and $t$, creating a spurious autocorrelation between $Y_{t-1}$ and $Y_t$.}
 \end{center}
@@
 By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past, and use these as features for the machine learning model.
-In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the product we wish to forecast, or characteristics of the location at which the product is sold.
+In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the products we wish to forecast, or characteristics of the location at which the product is sold.
 
@@
 Using the example from above, this would mean that we include features such as the price of the product or promotion details, weather information, as well as other variables describing the characteristics of a given product at a specific sales location, but do not include lagged information from the sales time series, such as its EWMA or similar approaches.
+
 Therefore, the features do not include any variable that is directly constructed from the previously observed sales events.
-As a consequence, the machine learning model can learn from the true causal dependencies expressed in the feature variables, which in turn improves both the explainability of the predictions, as well as the generalizability of the prediction model, because the causal dependencies are not confused by open backdoor paths leading to spurious (auto)correlations between the features and the prediction target.
+As a consequence, the machine learning model needs to learn from the true causal dependencies expressed in the exogenous features, which improves both the explainability of the predictions, as well as the generalizability of the prediction model, because the causal dependencies are not confused by additional spurious autocorrelations.
+It should be noted that variables describing certain seasonal effects, like the day of the week or the day of the year, can be included in the machine learning model without any negative effects.
+In fact, these represent seasonal causal dependencies of the target.
 
-However, we need to keep in mind that this approach of handling temporal confounders improves the prediction model used for demand forecasting, but is not yet sufficient for true causal inferences.
+We need to keep in mind that, although this approach of handling temporal confounders improves the prediction model used for demand forecasting, it is not yet sufficient for true causal inferences.
 For this, we would also have to control for the confounding effects between the different exogenous features of the machine learning model and its prediction target.
@@
 By including all available exogenous variables as features in the machine learning model but excluding any endogenous information from the target autocorrelation we can, at least partially, control for the temporal confounders.
-Any remaining autocorrelation between different lags of the time series of the residuals formed from the original time series and the time series of the forecasted value from the machine learning model discussed above can therefore originate from two reasons: Either there is a genuine autocorrelation between different lags $Y_{t-\tau}$ and $Y_t$ of the variable we wish to forecast, or we have not included all relevant causal variables in the machine learning model.
-To make use of this remaining autocorrelation, we suggest to apply a residual correction on each of the predictions of the machine learning model, based on the deviations in the recent past between the target and the predictions from the machine learning model.
+Any remaining autocorrelation between different lags of the time series of the residuals formed from the original time series and the time series of the forecasted values from the machine learning model discussed above can therefore originate from two reasons: Either there is a genuine autocorrelation between different lags $Y_{t-\mathrm{lag}}$ and $Y_t$ of the variable we wish to forecast, or we have not included all relevant causal variables in the machine learning model.
+To make use of this remaining autocorrelation, we suggest to apply a residual correction on each of the predictions of the machine learning model, based on the deviations in the recent past between the target and these predictions.
 Such a residual correction heuristically captures the recent trends without modeling them explicitly, and can be seen as an empirical correction to improve the final forecast quality.
@@
 \begin{equation} \label{eqn:emov_correction}
-\hat{y} = \frac{\mathrm{EWMA}(y)}{\mathrm{EWMA}(\hat{y})} \cdot \hat{y}_{\mathrm{ML}}
+\hat{y} = \frac{\mathrm{EWMA}(y)}{\mathrm{EWMA}(\hat{y}_{\mathrm{ML}})} \cdot \hat{y}_{\mathrm{ML}}
 \end{equation}
-In most cases, simple exponential smoothing will be sufficient for the residual correction, since we have already incorporated all relevant external effects, including seasonalities like a dependency of the target value on the day of the week, in the machine learning model before, and only seek to capture the information that we have missed, because, for example, we do not have the relevant data to describe this as a further feature variable.
+In most cases, simple exponential smoothing will be sufficient for the residual correction, since we have already incorporated all relevant external effects, including seasonalities like a dependency of the target on the day of the week, in the machine learning model before, and only seek to capture the information that we have missed, because, for example, we do not have the relevant data to describe this as a further feature variable.
 However, we need to be aware that if we have missed a crucial temporal confounder when building the features for the machine learning model, applying this residual correction risks opening a backdoor path for this confounder, which in turn might result in slightly sub-optimal predictions, as discussed in sec.
@@
 
-\section{Example: Demand Forecasting}
+\section{Example: Retail Demand Forecasting}
 \label{sec:example}
@@
 In the following, we describe how to use the approach outlined above in a practical setting.
-We use a public dataset obtained from a Kaggle online competition focusing in estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual items for specific stores on specific days.
+We use a public dataset obtained from a Kaggle online competition focusing on estimating unit sales of Walmart retail goods \cite{kaggle_data} for individual items for specific stores on specific days.
 For each demand forecast, Cyclic Boosting is used to predict the full probability distribution of the expected demand at a granularity of (item, store, day) and use the methods described in sec.
 \ref{sec:pdfEvaluation} to evaluate the quality of the individual forecasts.
-Each data record corresponding to an observed sales record is described by the following fields: the identifier of an individual store store (\texttt{store\_id}), the product identifier (\texttt{item\_id}) as well as the date.
+Each data record corresponding to an observed sales record is described by the following fields: the identifier of an individual store (\texttt{store\_id}), the product identifier (\texttt{item\_id}), and the date(\texttt{date}).
 The target $y$, that we need to predict, is the number of sales of a given product in a given store on a specific day, denoted by \texttt{sales}.
 
-For our experiments, we use data from 2013-01-01 to 2016-05-22, that describe the sales of 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3} in 10 stores.
-All data before 2016 are used as the training data and the data from 2016 are used as an independent test or validation set.
-Besides the fields used to identify an individual sales record and the corresponding observed sales value, namely \texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}, we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price} in the forecasts and multiple features are built from these variables that are then used in the machine learning models.
+For our experiments, we use data from 2013-01-01 to 2016-05-22, that describe the sales of 100 different products (\texttt{FOODS\_3\_500}, ..., \texttt{FOODS\_3\_599}) of the department \texttt{FOODS\_3} in all 10 available stores.
+All data before 2016 are used as the training data, and the data from 2016 are used as an independent test set.
+Besides the fields used to identify an individual sales record and the corresponding observed sales value, namely \texttt{item\_id}, \texttt{store\_id}, \texttt{date}, \texttt{sales}, we also use the fields \texttt{event\_name\_1}, \texttt{event\_type\_1}, \texttt{snap\_CA}, \texttt{snap\_TX}, \texttt{snap\_WI}, \texttt{sell\_price}, and \texttt{list\_price}, and multiple features built from these variables are then included in the machine learning models.
 
@@
 
-As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume a negative binomial distribution to model the individual probability distribution of each sales event.
-As a first step, we use Cyclic Boosting as the supervised machine learning method (as described in sec.
+As discussed earlier, we assume that each individual sale can be described by a Poisson-like process and we assume an NBD to model the individual probability distribution of each sales event.
+As a first step, we use the supervised machine learning method Cyclic Boosting (as described in sec.
 \ref{sec:CB}) to predict the mean of the distribution.
 
-\noindent
 This model uses the following variables as features: categorical variables for \texttt{store\_id} and \texttt{item\_id}, several derived variables that are constructed from the time series of the sales records describing trend and seasonality (days since beginning of 2013 as linear trend as well as day of week, day of year, month, and week of month), time windows around the events given in the data set (7 days before until 3 days after for Christmas and Easter, and 3 days before until 1 day after for all other events like New Year or Thanksgiving), a flag denoting a promotion, and the ratio of reduced (\texttt{sell\_price}) and normal price (\texttt{list\_price}).
@@
 As discussed in sec.
-\ref{sec:ts_confounding}, we do not use any lagged target information in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with $\alpha=0.15$) of the predictions and targets of each product-location combination over the corresponding past.
+\ref{sec:ts_confounding}, we do not use any lagged target information, for example via the inclusion of target EWMA features, in the machine learning model, but apply an individual residual correction on each of its predictions, which accounts for deviations between the EWMA (with $\alpha=0.15$) of the predictions and targets of each product-location combination over the corresponding past.
 Empirically, a value of $\alpha=0.15$ gives good results here, though in a practical application, this parameter would have to be optimized using methods such as cross-validation according to a metric that reflects the objective of the business strategy.
@@
 
-\noindent
-We use the mean absolute deviation (MAD) as well as the mean squared error (MSE) as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean.
+We use the MAD as well as the MSE as two common metrics for the evaluation of point estimates to give a rough estimate of the accuracy of the predicted mean.
 These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales.
@@
 \begin{center}
-\caption{Accuracy metrics for the predicted mean values for three different model setups: a) ML model without lagged target information but subsequent residual correction, b) ML model including lagged target information by means of two EMWA features without residual correction, and c) ML model from b) with residual correction.}
+\caption{Accuracy metrics for the predicted mean values for three different model setups: ML model without lagged target information but subsequent residual correction (a), ML model including lagged target information by means of two EMWA features without residual correction (b), and ML model from b) with residual correction (c).}
 \label{tab:mad_mse}
@@
 
-\noindent
-To visualize the operation principle of the residual correction method, \fig \ref{fig:ts_res} shows the time series of sales and machine learning (before residual correction) as well as final predictions (after residual correction) for item \texttt{FOODS\_3\_516} in store \texttt{CA\_3} from beginning of November 2015 to end of April 2016, including all days in the test period in 2016 as well as the last two months of the training period before.
+To visualize the operation principle of the residual correction method, \fig \ref{fig:ts_res} shows the time series of sales and machine learning (before residual correction) as well as final predictions (after residual correction) for item \texttt{FOODS\_3\_516} in store \texttt{CA\_3} from beginning of November 2015 to end of April 2016, including all days of the test period in 2016 as well as the last two months of the training period before.
 Hereby, the machine learning predictions in the training period are in-sample predictions, which are just used to calculate the residual correction.
 A sudden change of the sales time series at the turn of the year is visible, which is seemingly not anticipated by the machine learning model, as its predictions are clearly off in the test period in 2016.
-However, the residual correction is able to account for the resulting deviations between sales and machine learning predictions, which can be seen by the much better agreement of the final predictions.
+However, the residual correction is able to quickly account for the resulting deviations between sales and machine learning predictions, which can be seen by the much better agreement of the final predictions.
 
@@
 To give an indication of the effectiveness of the two-step approach, i.e., a machine learning model without lagged target information and residual correction as outlined in sec.
-\ref{sec:ts_confounding}, we compare its results (for the same training and test setup) with a model following the common practice of including the lagged target information as a variety of EMWA variables used as features in \tab \ref{tab:mad_mse}.
+\ref{sec:ts_confounding}, we compare its results with a model following the common practice of including the lagged target information as a variety of EMWA variables used as features in \tab \ref{tab:mad_mse} (for the same training and test setup).
 For this, we include two EWMA features in the machine learning model, one grouped by \texttt{store\_id} and \texttt{item\_id} (with a lag of two days and $\alpha=0.25$), and one grouped by \texttt{store\_id}, \texttt{item\_id}, and the day of the week (with a lag of one week and $\alpha=0.05$), where the values for $\alpha$ are again found empirically.
@@
 The second model, based on Cyclic Boosting in its negative binomial width mode described in sec.
-\ref{sec:cb_width}, is used to estimate the variance of the negative binomial distribution.
-The data are split into training and test set as above and in addition the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}.
-This effectively means that the mean predictions are created in-sample for the training period using the fully trained and validated model for the mean discussed above.
+\ref{sec:cb_width}, is used to estimate the variance of the NBD.
+The data are split into training and test sets as above, and in addition, the mean predictions for each individual product-location-date combination are fixed in the variance model as stated by \eqn \eqref{eqn:loss_likelihood}.
+This effectively means that the needed mean predictions are created in-sample for the training period using the fully trained and validated model for the mean discussed above.
 
@@
 Using the same mean predictions for both negative binomial and Poisson model, the negative binomial PDF predictions are much closer to the uniform distribution, which we expect for optimal PDF predictions, than the Poisson PDF predictions, showing the effectiveness of our variance estimation.
-
-\noindent
 While the Poisson histogram shows a clear pattern of too narrow PDF predictions (see \fig \ref{fig:cdf_histos}), the most significant deviations of the negative binomial histogram from the uniform distribution can be found in the first bins of CDF values close to $0$ (overprediction) as well as in the last bins of CDF values close to $1$ (underprediction).
@@
 
-Using the inverse quantile profile plots introduced earlier, the prediction quality of the full predicted PDF can be assessed in more detail.
+Using the inverse quantile profile plots introduced in sec.
+\ref{sec:invquant_plot}, the prediction quality of the full predicted PDF can be assessed in more detail.
 In \fig \ref{fig:invquant_dayofweek}, each column corresponds to a day of the week, using 0 for Mondays, 1 for Tuesdays, etc., and the considered quantiles are 0.1, 0.3, 0.5, 0.7, 0.9, and 0.97.
@@
 \includegraphics[width=4cm]{figs/invquant_dayofweek_poisson}
-\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$-axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
+\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$ axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
 \end{center}
@@
 
-\fig \ref{fig:invquant_mean} shows the inverse quantile profile plot using the predicted mean as the $x$-axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
+\fig \ref{fig:invquant_mean} shows the inverse quantile profile plot using the predicted mean as the $x$ axis of the graph, meaning that the distributions are grouped such that the 15 columns correspond to mean predictions in 12 intervals with widths of 5 from 0.0 to 60.0 and 3 remaining intervals $[60.0, 70.0]$, $(70.0, 80.0]$, and $(80.0, 100.0]$ (with mean predictions higher than 100.0 included in the highest interval), while we aggregate over all locations, items, and sales dates.
 The considered quantiles are chosen to be the same as in \fig \ref{fig:invquant_dayofweek}.
@@
 However, it should be noted that due to this segregation, the statistics in each part of the relevant test sample becomes a limiting factor as well.
-This plot also illustrates the benefits of using the quantile profile plot, since we have seen that using more conventional approaches, such as \fig \ref{fig:mean_prediction} or even the lower part of \fig \ref{fig:cdf_demand}, the predictions appear reasonable, even when not relying on simple point metrics such as MAD or MSE.
+This plot also illustrates the benefits of using the quantile profile plot, since we have seen that using more conventional approaches, such as \fig \ref{fig:mean_prediction} or even \fig \ref{fig:cdf_demand}, the predictions appear reasonable, even when not relying on simple point metrics such as MAD or MSE.
 It is therefore paramount both to predict the full PDF instead of just a point estimate, such as the mean, as well as verifying a number of quantiles for each prediction to assess the quality of the prediction thoroughly.
@@
 \includegraphics[width=4cm]{figs/invquant_yhat_mean_poisson}
-\caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$-axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
+\caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$ axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
 \end{center}
@@
 Both from practical and theoretical perspectives, disentangling the forecast of demand from the operational decision making regarding order quantities has potentially significant advantages over an integrated approach using data to predict the resulting order quantities directly.
-When predicting future demand in the form of individual probability distributions, both model-free and a model-based approaches can be used, where the model-based approach is generally more robust, as it is rooted in a theoretical understanding of the sales process.
+When predicting future demand in the form of individual probability distributions, both model-free and model-based approaches can be used, where the model-based approach is generally more robust, as it is rooted in a theoretical understanding of the sales process.
 Compared to a model-free approach like quantile regression, the distributional assumption can drastically reduce the uncertainty of the resulting predictions.
@@
 
-\section{Optimal Point Estimator for Quadratic Cost Function}
+\section{Optimal Point Estimator for Given Cost Function}
 \label{sec:CostQuantile}
 
-Operational decisions ultimately need to be based on a single number, for example, if we want to order new goods, in the end we need a single number representing the expected demand, include operational constraints such as lot sizes and delivery schedules from the manufacturer or wholesaler to arrive at the optimal order quantity for the next delivery period.
-If we predict the future demand as a complete probability density distribution, we need to determine the optimal point estimator that ``reduces'' the PDF to the single best quantity.
-In an ideal case, the full cost function relating point estimator $p_i$ to observed (true) value $t_i$: $C(p_i,t_i)$.
-If the cost function is fully specified, we can obtain the optimal point estimator $p$ by minimizing the cost function with respect to the predicted probability distribution $f(t)$, i.e., we require that the first derivative of the cost function vanishes.
+Operational decisions ultimately need to be based on a single number.
+For example, if we want to order new goods, we need a single number representing the expected demand, and then include operational constraints such as lot sizes and delivery schedules from the manufacturer or wholesaler to arrive at the optimal order quantity for the next delivery period.
+If we predict the future demand as a complete PDF, we need to determine the optimal point estimator that ``reduces'' the PDF to the single best quantity.
+In an ideal case, the full cost function $C(p_i,t_i)$ relating point estimator $p_i$ to observed (true) value $t_i$ is available.
+If the cost function is fully specified, we can obtain the optimal point estimator $p$ by minimizing the cost function with respect to the predicted probability distribution $f(t)$, i.e., we require that the first derivative of the cost function vanishes:
 \begin{equation}
@@
 \end{equation}
-Since the variables are random variables, we need to include the expectation value $E[\cdot]$.
-A common metric to evaluate point estimators is the mean squared error (MSE) which is a quadratic cost function given by $C(p,t) = (p-t)^2$.
-We can use this to derive the optimal point estimator associated with this particular cost function:
+Since we are dealing with random variables, we need to include the expectation value $E[\cdot]$.
+A common metric to evaluate point estimators is the squared error, i.e., a quadratic cost function given by $C(p,t) = (p-t)^2$.
+We can then derive the optimal point estimator associated with this particular cost function:
 \begin{align*}
@@
 \end{equation}
-Therefore, the optimal point estimator associated with a quadratic cost function is the mean of the predicted probability distribution for the demand.
+Therefore, the optimal point estimator associated with a quadratic cost function is the mean of the predicted probability distribution.
 
-In the case of the original newsvendor problem, underage and overage cost are linear. In the special case that they are identical, the cost function is given by the MAD, i.e., $C(p,t) = |p-t|$.
+In the case of the original newsvendor problem, underage and overage cost are linear.
+In the special case that they are identical, the cost function is given by the absolute deviation, i.e., $C(p,t) = |p-t|$.
 We can derive the optimal point estimator in the following way, using the shorthand notation $\partial_p X$ for $\partial/\partial p X$
@@
 This allows us to resolve the absolute value if we choose $x_q$ such that $t$ is always smaller than $p$ in the first integral and the other way around in the second.
-Hence,
+Hence:
 \begin{equation}
@@
 \begin{equation}
-0 = -  \int_{-\infty}^{x_p} f(t) dt + \int_{x_p}^\infty f(t)  dt
+0 = -  \int_{-\infty}^{x_p} f(t) dt + \int_{x_p}^\infty f(t)  dt,
 \end{equation}
@@
 Therefore, the integral from $-\infty$ to the point $x_p$ is the same as from $x_p$ to $\infty$, which means that the point $x_p$ splits the integral in the interval $(-\infty, \infty)$ into two halves.
-This is the definition of the median and therefore a linear cost function with equal over- and underage costs are linked to the median of the predicted probability density function as optimal point estimator.
-If the underage and overage cost differ, effectively the slope of the two linear segments are no longer the same.
-This means that the resulting optimal point estimator is no longer the median but a different quantile.
+This is the definition of the median, and therefore a linear cost function with equal over- and underage costs is linked to the median of the predicted PDF as optimal point estimator.
+If the underage and overage costs differ, the slopes of the two linear segments are no longer the same.
+This means that the resulting optimal point estimator is no longer the median, but a different quantile.
 
@@
 In many cases, scatter plots are used to study the behavior of two distributions or sets of data points visually.
-However, even for moderate amount of data, this approach quickly becomes difficult.
-To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: The distribution of $x$ values was obtained by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and the $y$ values are obtained via $Y \sim X +  {\cal N}(2.0,1.5)$.
+However, even for a moderate amount of data, this approach quickly becomes difficult.
+To illustrate this, a sample of $(x,y)$ data points was obtained in the following way: $x$ values by generating 5,000 samples of Gaussian distributed random numbers $X \sim {\cal N}(0.0,2.0)$ and $y$ values via $Y \sim X +  {\cal N}(2.0,1.5)$.
 \fig \ref{fig:scatter} shows the marginal distributions for $x$ and $y$ as well as a scatter plot of $x$ vs.
@@
 \includegraphics[scale=0.5]{figs/marginal} \includegraphics[scale=0.5]{figs/scatter}
-\caption{\label{fig:scatter} Marginal distribution and scatter plot of variables $X$ and $Y$.}
+\caption{\label{fig:scatter} Marginal distributions and scatter plot of variables $X$ and $Y$.}
 \end{center}
@@
 
-Although the simple linear correlation between $X$ and $Y$ is apparent in the scatter plot, finer details are not visible and it is easy to imagine that a more complex relationship is difficult to discern.
+Although the simple linear correlation between $X$ and $Y$ is apparent in the scatter plot, finer details are not visible, and it is easy to imagine that a more complex relationship is difficult to discern.
 Profile histograms are specifically designed to address this shortcoming.
@@
 One can for example choose equidistant bins in the range of the $x$ axis or non-equidistant bins such that each bin contains the same number of observations.
-Then within each bin of the variable $X$, the a location and dispersion metric is calculated for the variable $Y$.
-This means that the bin-borders on the $X$ axis are used as constraints on the variable $Y$ and with these conditions applied, for example the sample mean of the selected $y$ values as well as the standard deviation are calculated.
+Then within each bin of the variable $X$, a location and a dispersion metric are calculated for the variable $Y$.
+This means that the bin-borders on the $x$ axis are used as constraints on the variable $Y$, and with these conditions applied, for example the sample mean of the selected $y$ values as well as the corresponding standard deviation are calculated.
 These location and dispersion metrics in each bin of $X$ are used to illustrate the behavior of the variable $Y$ as the values of the variable $X$ change from bin to bin.
@@
 This one-dimensional representation allows to understand even a complex relationship between two variables visually.
-Note that due to few data points at the edges of the distributions the profile histogram is expected to show visual artifacts in the corresponding regions.
+Note that due to few data points at the edges of the distributions, the profile histogram is expected to show visual artifacts in the corresponding regions.
 

### 2021-01-01 LaTex/paper.tex
some more changes before resubmit to SNORF

@@
 Furthermore, the cost function may vary between individual stock-keeping units (SKU) and may also change with time.
-However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF, what is demonstrated in Appendix \ref{sec:CostQuantile}.
-For the case of a quadratic cost function, such as the mean squared error (MSE), or the linear cost function used in the newsvendor setting, corresponding to the mean absolute deviation (MAD), the calculations are straight-forward.
+However, the general principle remains: If we have access to the relevant cost function, we can determine the optimal point estimator of the future demand from the associated PDF, see Appendix \ref{sec:CostQuantile}.
+For the case of a relatively simple cost function, such as the mean squared error (MSE) or the mean absolute deviation (MAD), the calculations are straight-forward.
 If the cost function becomes more complex, we may no longer be able to perform the calculations analytically, but will have to resort to numerical approaches.
-Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different points, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change when choosing a different quantile.
-This can be done using, for example, a business oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business relevant KPIs for each setting, taking operational constraints into account.
+Furthermore, if we can predict more than one quantile, we can investigate the effect of choosing different quantiles, and determine, for example, how sensitive operational metrics, such as stockout or waste rate of perishable goods, change.
+This can be done using, for example, a business-oriented simulation that starts from the quantiles of the predicted demand for each SKU to simulate the value of the business-relevant KPIs for each setting, taking operational constraints into account.
 This can then be used, for example, to align the business strategy.
@@
 % PDF predictions
-Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated PDF.
+Due to its stochastic nature, demand is difficult to forecast: It depends on many influencing factors and, as already stated above, we can interpret it as a random variable with associated probability distribution.
 And as such, demand is neither identically and independently distributed (i.i.d.) between different products and sales locations, nor between different times.
@@
 Furthermore, this distribution depends on a wide range of influencing factors, such as seasonal effects, sales price, local weather, etc.
-Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual PDFs rather than globally adjusted parameters for groups of SKUs or locations or sales days.
+Therefore, in order to predict the expected demand for each SKU at each sales location and sales day, we need to forecast individual probability distributions rather than globally adjusted parameters for groups of SKUs, locations, or sales days.
 
@@
 It is therefore tempting to use established time series regression models to predict the future values of the expected demand.
-However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed time series variable $Y$, say $Y_{t}$ and $Y_{t-\mathrm{lag}}$, where lag denotes a specific time difference between the two steps.
-In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product which we can set as part of our business strategy or change according to some promotion.
+However, if we wish to use the predicted values for anything but the prediction itself, we have to be careful to avoid, or at least limit as much as possible, temporal confounding, which emerges if some external variable $X$ causally influences several time steps $t$ of the observed time series variable $Y$, say $Y_{t}$ and $Y_{t-\mathrm{lag}}$, where ``lag'' denotes a specific time difference between the two steps.
+In the case of a retailer, $Y_{t}$ is the demand for a specific product at a given sales location for a specific day, and $X$ can be, for example, the price of the product, which we can set as part of our business strategy or change according to some promotion.
 This dependency creates a backdoor path between $Y_{t-\mathrm{lag}}$ and $Y_{t}$, that in turn leads to a spurious autocorrelation between the values of $Y$ at different times $t$, if we cannot control for the confounder $X$.
-This is particularly important for the case of demand forecasting, since customer demand is not a mere natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc.
+This is particularly important for the case of demand forecasting, since customer demand is not a natural phenomenon we observe as an outsider, but retailers actively influence the customer demand  through a wide range of interventions, such as a given assortment choice, promotions, advertisements, coupons, etc.
 If we cannot disentangle the fundamental causal dependencies from the spurious autocorrelation, understanding the influences that lead to a specific prediction becomes very challenging, even if the underlying prediction method itself is fully explainable.
@@
 Univariate time series forecasting methods operate on each of these many individual time series separately.
-On the other hand, supervised machine learning (ML) algorithms optimize all individual time series together, which can enhance the generalizability of the method by exploiting common aspects between the various time series.
+On the other hand, supervised machine learning (ML) algorithms optimize all individual time series simultaneously, which can enhance the generalizability of the method by exploiting common aspects between the various time series.
 This can lead to a drastic reduction in variance of the model, and in turn to an improvement of the forecast quality of each of the individual time series.
-Furthermore, machine learning provides a straight-forward way to include the many exogenous variables influencing demand as features (or covariates) of the model.
+Furthermore, machine learning provides a straight-forward way to include the many exogenous variables that influence demand as features (or covariates) of the model.
 This can also help to improve the forecast quality compared to methods mainly relying on autocorrelation of the historic sales.
@@
 
-Furthermore, we show how predictions of full individual probability density functions can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form as well as specific quantiles are predicted correctly.
+Furthermore, we show how predictions of full individual probability distributions, no matter if created by an approach with or without assumption of a specific distribution class, can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form (in the case of a parametric approach) as well as specific quantiles are predicted correctly.
 This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.
@@
 % ML
-Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached using supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.
+Alternatively to the methods discussed above, time series forecasts in general and demand predictions in particular can be approached by means of supervised machine learning methods, using the values of the time series to be predicted as target, i.e., the numbers the machine learning algorithm is trained to predict.
 For example, see \cite{Zhang2012,remus2001neural} for applications of neural networks for time series forecasting.
@@
 For the case of individual samples for different time steps, in order to combine the endogenous information from the target autocorrelation with all other exogenous features, the most prominent approach is to include lagged target information explicitly via stacking of univariate time series predictions for the target as features in the machine learning model.
-For example, we could add exponentially smoothed moving averages with a range of dampening factors as features into the model.
+For example, we could add exponentially smoothed moving averages with a range of damping factors as features into the model.
 An example for the estimation of future demand in the context of price optimization, using regression trees \cite{breiman1984}, can be found in \cite{ferreira2016analytics}.
 For the case of sequence samples, deep learning methods such as recurrent neural networks (RNN) \cite{rnn}, especially in form of ``long short-term memory'' (LSTM) networks \cite{hochreiter1997long}, or transformers \cite{transformer}, based on the self-attention mechanism, are used.
-These methods use information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events.
-Both concepts, LSTM cells and self-attention, aim at learning from the data the elements of the individual time series that are deemed important for the prediction of the future values by the algorithm, and implicitly rely on the autocorrelation between the observed events at different times as well.
+These methods employ information about the prediction target autocorrelation implicitly via the input sequences of the time-ordered training events.
+Both concepts, LSTM cells and self-attention, aim at learning  the elements of the individual time series from the data that are deemed important for the prediction of the future values by the algorithm.
 The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}.
@@
 The main idea of this algorithm is the following: First, each feature, denoted by index $j$, is discretized appropriately into $k$ bins to reflect the specific behavior of the feature.
-The global mean $\mu$ is determined from all values $y$ of the  target variable $Y \in [0,\infty)$ observed in the data.
+The global mean $c$ is determined from all values $y$ of the target variable $Y \in [0,\infty)$ observed in the data.
 Single data records, for example, the sales corresponding to a specific product-location-date combination (meaning the sales record of a specific item sold on a specific day at a specific sales location) along with all relevant features, are indexed by $i$.
-The individual predictions $\hat{y_i}$  can then be calculated as:
+The individual predictions for the NBD mean, denoted by $\hat{\mu_i}$, can then be calculated as:
 \begin{equation} \label{eqn:cb}
-\hat{y}_i = \mu \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
+\hat{\mu}_i = c \cdot \prod \limits_{j=1}^p f^k_j \quad \text{with}\; k=\{ x_{j,i} \in b^k_j\}
 \end{equation}
@@
 \begin{enumerate}
-\item{Calculate the global average $\mu$ from all observed $y$ across all bins $k$ and features $j$.}
+\item{Calculate the global average $c$ from all observed $y$ across all bins $k$ and features $j$.}
 \item{Initialize the factors $f^k_j \leftarrow 1$}
 \item{Cyclically iterate through features $j = 1,...,p $ and calculate in turn for each bin $k$ the partial factors $g$ and corresponding aggregated factors $f$, where indices $t$ (current iteration) and $\tau$ (current or preceding iteration) refer to iterations of full feature cycles as the training of the algorithm progresses:
-\begin{equation} \label{factors}
-g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{y}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
+\begin{equation} \label{eqn:factors}
+g^k_{j,t} = \frac{\sum \limits_{x_{j,i} \in b^k_j} y_i}{\sum \limits_{x_{j,i} \in b^k_j} \hat{\mu}_{i,\tau}}\;\; \mathrm{where} \; \; f^k_{j,t} = \prod \limits_{s=1}^t g^k_{j,s}
 \end{equation}
 Here,  $g$ is a factor that is multiplied to the corresponding $f_{t-1}$ in each iteration.
-The current prediction, $\hat{y}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
-\begin{equation} \label{factors3}
-\hat{y}_{i,\tau} = \mu \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
+The current prediction, $\hat{\mu}_\tau$, is calculated according to \eqn \eqref{eqn:cb} with the current values of the aggregated factors $f$:
+\begin{equation}
+\hat{\mu}_{i,\tau} = c \cdot \prod \limits_{j=1}^p f^k_{j,\tau}
 \end{equation}
-To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{y}$.
+To be precise, the determination of $g^k_{j,t}$ for a specific feature $j$ uses $f^k_{j,t-1}$ in the calculation of $\hat{\mu}$.
 For the factors of all other features, the newest available values are used, i.e., depending on the sequence of features in the algorithm, either from the current ($\tau=t$) or the preceding iteration ($\tau=t-1$).}
@@
 \begin{equation} \label{eqn:nbinom}
-\mathrm{NB}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
+\mathrm{NBD}(y; \mu, r) = \frac{\Gamma(r + y)}{y! \cdot \Gamma(r)} \cdot \left(\frac{r}{r + \mu}\right)^r \cdot \left(\frac{\mu}{r + \mu}\right)^y,
 \end{equation}
@@
 The estimate of the dispersion parameter $\hat{r}$ can then be calculated by minimizing the loss function defined in \eqn \eqref{eqn:loss_likelihood}, which is expressed as negative log-likelihood function of an NBD.
-The minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained in the previous step (described in sec.
-\ref{sec:CB}).
+The minimization over all input samples $i$ is performed with respect to the Cyclic Boosting parameters $f^k_j$, constituting the model of $\hat{r_i}$, according to \eqn \eqref{eqn:r}, where the estimates for the mean $\hat{\mu_i}$ are fixed to the values obtained from the mean model described in sec.
+\ref{sec:CB}.
 
 \begin{equation} \label{eqn:loss_likelihood}
-L(r) = -\mathcal{L}(r) = -\ln \sum_i \mathrm{NB}(y_i; \hat{\mu_i}, \hat{r_i})
+L(r) = -\mathcal{L}(r) = -\ln \sum_i \mathrm{NBD}(y_i; \hat{\mu_i}, \hat{r_i})
 \end{equation}
@@
 
-The Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached.
-Unlike in the basic multiplicative regression mode of Cyclic Boosting described in sec.
-\ref{sec:CB}, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example, using a random search.
+In the same way as described in sec.
+\ref{sec:CB} for its basic multiplicative regression mode, the Cyclic Boosting algorithm is trained iteratively using cyclic coordinate descent, processing one feature with all its bins at a time until convergence is reached.
+However, unlike in \eqn \eqref{eqn:factors} of the basic multiplicative regression mode, the minimization of the loss function in \eqn \eqref{eqn:loss_likelihood} cannot be solved analytically and has to be done numerically, for example, using a random search.
 All other advantages of Cyclic Boosting, like individual explainability of predictions, remain valid for its negative binomial width mode.
@@
 Finally, the variance $\hat{\sigma}^2_i$ can be estimated from the dispersion parameter $\hat{r_i}$ using \eqn \eqref{eqn:variance_r}.
-Using  the individual predicted mean $\hat{\mu_i}$ from the first step, the NBD model is fully specified for each individual prediction $i$.
+And together with the individual predicted mean $\hat{\mu_i}$ from the first step, the NBD model is then fully specified for each individual prediction $i$.
 
@@
 Many statistical and most machine learning methods do not provide a full PDF as result.
-Instead, these methods typically predict a single numerical quantity (denoted by $\hat{y}$) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the MSE, MAD, or others.
-In the setting of a retailer, the observed quantity is the sales of individual products and most machine learning approaches would then predict a single number as a direct estimate of the sales.
+Instead, these methods typically predict a single numerical quantity (usually denoted by $\hat{y}$, in this work denoted by $\hat{\mu}$ to make evident that it is the prediction of the mean of an underlying probability distribution) that is then compared to the observed concrete realization of the random variable (denoted by $y$) using metrics such as the MSE, MAD, or others.
+In the setting of a retailer, the observed quantity is the sales of individual products and most approaches would then predict a single number as a direct estimate of the sales.
 However, reducing the prediction to a single number does not allow to account for the uncertainty of the prediction or the dynamics of the system.
 Instead, it is imperative to predict the full PDF for each prediction to be able to optimize the subsequent operational decision.
-Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or even qualitative evaluation methods to assess whether the full distribution has been forecasted correctly, in particular in the tails of the distribution.
+Unfortunately, most statistical or machine learning methods that predict full individual probability functions lack quantitative or even qualitative evaluation methods to assess whether the full distribution has been forecasted correctly, in particular in its tails.
 
@@
 
-In practical applications however, we have a large number of effective prediction models, since although we always use the same model parameterization, such as the NBD, its parameters have to be determined at the required level of granularity.
-For example, for daily orders, we need to predict the parameters of the NBD for each location, sales day, and product.
+In practical applications however, we have a large number of effective prediction models, since although we may always use the same model parameterization, such as the NBD, its parameters have to be determined at the required level of granularity.
+For example, for daily orders, we need to predict the parameters of the NBD for each product, location, and sales day.
 Unlike the simple case discussed above, where we had many observations to compare the prediction model to, we now have just a single observation per prediction, meaning that we cannot use statistical tests directly.
@@
 In the following, we call this method {\em inverse quantile profile plot}.
-Profile plots are akin to scatter plots and described in more detail in appendix \ref{sec:profile}.
+Profile plots are akin to scatter plots and described in more detail in Appendix \ref{sec:profile}.
 
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/temporal_confounding}
+\includegraphics[width=5cm]{figs/temporal_confounding}
 \caption{\label{fig:temporal_confounding} Temporal confounding by an exogenous variable $X_{t-1, t}$ stable over the time steps $t-1$ and $t$, creating a spurious autocorrelation between $Y_{t-1}$ and $Y_t$.}
@@
 Any remaining autocorrelation between different lags of the time series of the residuals formed from the original time series and the time series of the forecasted values from the machine learning model discussed above can therefore originate from two reasons: Either there is a genuine autocorrelation between different lags $Y_{t-\mathrm{lag}}$ and $Y_t$ of the variable we wish to forecast, or we have not included all relevant causal variables in the machine learning model.
-To make use of this remaining autocorrelation, we suggest to apply a residual correction on each of the predictions of the machine learning model, based on the deviations in the recent past between the target and these predictions.
+To make use of this remaining autocorrelation, we suggest to apply a residual correction on each of the predictions of the machine learning model $\hat{\mu}_{\mathrm{ML}}$ (assuming predictions of the mean of an underlying PDF as discussed in sec.
+\ref{sec:pdfEstimation}), based on the deviations in the recent past between the target and these predictions.
 Such a residual correction heuristically captures the recent trends without modeling them explicitly, and can be seen as an empirical correction to improve the final forecast quality.
@@
 \begin{equation} \label{eqn:emov_correction}
-\hat{y} = \frac{\mathrm{EWMA}(y)}{\mathrm{EWMA}(\hat{y}_{\mathrm{ML}})} \cdot \hat{y}_{\mathrm{ML}}
+\hat{\mu} = \frac{\mathrm{EWMA}(y)}{\mathrm{EWMA}(\hat{\mu}_{\mathrm{ML}})} \cdot \hat{\mu}_{\mathrm{ML}}
 \end{equation}
@@
 These metrics do not take the shape of the underlying PDF into account, but only compare the predicted mean to the observed number of sales.
+As shown in Appendix \ref{sec:CostQuantile}, the mean of a predicted PDF is the optimal point estimator for the MSE and the median for the MAD.
+However, as we are only interested in relative numbers for the sake of our comparisons here, we just use our mean prediction in both metrics.
 Using the independent test data, we obtain the metrics as stated in the top line of \tab \ref{tab:mad_mse}.
@@
 \includegraphics[width=4cm]{figs/invquant_dayofweek_poisson}
-\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of week on the $x$ axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
+\caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of the week on the $x$ axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
 \end{center}
@@
 An additional challenge for demand forecasting in specific and time series forecasting in general is the handling of the detrimental effects of temporal confounding, especially on explainability and generalizability of the employed models.
-Using a two-step model consisting of a machine learning model, including all available exogenous variables as features but excluding any lagged target information, to control for all known temporal confounders, and a subsequent empiric residual correction to capture any information on recent trends missed in the ML model, temporal confounding can be avoided, or at least mitigated.
+Using a two-step model consisting of a machine learning model, including all available exogenous variables as features but excluding any lagged target information, to control for all known temporal confounders, and a subsequent empiric residual correction to capture any information on recent trends missed in the machine learning model, temporal confounding can be avoided, or at least mitigated.
 
@@
 %\subsection*{Authors' contributions}
-%F. Wick: Code examples and manuscript text, U. Kerzel: manuscript text, M. Hahn, M. Wolf, T. Singhal, M. Feindt, D. Stemmer, J. Ernst: revision
+%F. Wick: code examples and manuscript text, U. Kerzel: manuscript text, M. Hahn, M. Wolf, T. Singhal, M. Feindt, D. Stemmer, J. Ernst: revision
 
@@
 
-In the case of the original newsvendor problem, underage and overage cost are linear.
+In the case of the original newsvendor problem, underage and overage costs are linear.
 In the special case that they are identical, the cost function is given by the absolute deviation, i.e., $C(p,t) = |p-t|$.
@@
 \begin{center}
-\includegraphics[scale=0.5]{figs/marginal} \includegraphics[scale=0.5]{figs/scatter}
+\includegraphics[width=4cm]{figs/marginal}
+\includegraphics[width=4cm]{figs/scatter}
 \caption{\label{fig:scatter} Marginal distributions and scatter plot of variables $X$ and $Y$.}
@@
 \begin{center}
-\includegraphics[scale=0.5]{figs/profile}
+\includegraphics[width=8cm]{figs/profile}
 \caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}

### 2021-01-01 code/walmart_demand.py

@@ -309,7 +309,7 @@ def plotting(df, suffix=''):
 
     predictions = df[(df['item_id'] == 16) & (df['store_id'] == 2) & (df['date'] < '2016-05-01')]
     ts_data = predictions.groupby(['date'])['y', 'yhat_mean', 'yhat_var', 'yhat_mean_causal'].sum().reset_index()
-    plot_timeseries(ts_data, 'item_16_store_2_res' + suffix, textbox='a)', errorband=False, split_line=True)
+    plot_timeseries(ts_data, 'item_16_store_2_res' + suffix, errorband=False, split_line=True)
 
 
 def eval_results(yhat_mean, y):

### 2021-01-02 LaTex/paper.tex
moved figs from subfolder up

@@
 \begin{center}
-\includegraphics[width=8cm]{figs/PdfCdf}
+\includegraphics[width=8cm]{PdfCdf}
 \caption{\label{fig:PdfCdf} PDF $f(x)$ and CDF $F(x)$ of a normal distribution.}
@@
 \begin{center}
-\includegraphics[scale=0.25]{figs/cdf_truth_broad}
-\includegraphics[scale=0.25]{figs/cdf_truth_narrow}
-\includegraphics[scale=0.25]{figs/cdf_truth_over}
-\includegraphics[scale=0.25]{figs/cdf_truth_under}
+\includegraphics[scale=0.25]{cdf_truth_broad}
+\includegraphics[scale=0.25]{cdf_truth_narrow}
+\includegraphics[scale=0.25]{cdf_truth_over}
+\includegraphics[scale=0.25]{cdf_truth_under}
 \caption{\label{fig:cdf_histos} Histogram of actually observed CDF values for different cases of estimating the full PDF compared to the expected uniform distribution (dashed horizontal line in each of the plots): variance overestimated (a: ``broad''), variance underestimated (b: ``narrow''), mean overestimated (c: ``over''), mean underestimated (d: ``under'')}
@@
 \begin{center}
-\includegraphics[width=8cm]{figs/invquant_example}
+\includegraphics[width=8cm]{invquant_example}
 \caption{\label{fig:invquant_example} Five different inverse quantile profile plots (separated into columns by dotted vertical lines), each comparing all its predicted PDFs to the corresponding observed events.
@@
 \begin{center}
-\includegraphics[width=5cm]{figs/temporal_confounding}
+\includegraphics[width=5cm]{temporal_confounding}
 \caption{\label{fig:temporal_confounding} Temporal confounding by an exogenous variable $X_{t-1, t}$ stable over the time steps $t-1$ and $t$, creating a spurious autocorrelation between $Y_{t-1}$ and $Y_t$.}
@@
 \begin{center}
-\includegraphics[width=8cm]{figs/ts_full}
+\includegraphics[width=8cm]{ts_full}
 \caption{\label{fig:mean_prediction} Time series of final mean predictions as well as sales in test period summed over all products and stores.}
@@
 \begin{center}
-\includegraphics[width=8cm]{figs/ts_item_16_store_2_res}
+\includegraphics[width=8cm]{ts_item_16_store_2_res}
 \caption{\label{fig:ts_res} Time series of mean predictions, before (machine learning prediction) and after (final prediction) residual correction, as well as sales for a specific product-location combination.
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/pdf}
-\includegraphics[width=4cm]{figs/cdf}
+\includegraphics[width=4cm]{pdf}
+\includegraphics[width=4cm]{cdf}
 \caption{\label{fig:pdf_example} Predicted PDF (a) and CDF (b) for a specific product-location-date combination.
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/ts_item_16_store_6}
-\includegraphics[width=4cm]{figs/factors_ts}
-\includegraphics[width=4cm]{figs/factors_ts_width}
+\includegraphics[width=4cm]{ts_item_16_store_6}
+\includegraphics[width=4cm]{factors_ts}
+\includegraphics[width=4cm]{factors_ts_width}
 \caption{\label{fig:factor_plots} Time series of mean predictions (with the surrounding transparent band simplistically indicating a standard deviation as square root of the corresponding variance predictions), before (machine learning prediction) and after residual correction (final prediction), as well as sales (a), constituting factors of the mean predictions (including the factor of the residual correction) (b), and constituting factors of the variance predictions (c), for a specific product-location combination from beginning of February to end of April 2016.
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/cdf_truth_nbinom}
-\includegraphics[width=4cm]{figs/cdf_truth_poisson}
-\includegraphics[width=4cm]{figs/cdf_truth_nbinom_larger1}
-\includegraphics[width=4cm]{figs/cdf_truth_poisson_larger1}
+\includegraphics[width=4cm]{cdf_truth_nbinom}
+\includegraphics[width=4cm]{cdf_truth_poisson}
+\includegraphics[width=4cm]{cdf_truth_nbinom_larger1}
+\includegraphics[width=4cm]{cdf_truth_poisson_larger1}
 \caption{\label{fig:cdf_demand} Histograms of ex post observed target CDF values of the corresponding individual PDF predictions (to be compared to a uniform distribution) for our negative binomial model (a) and a simpler Poisson model for comparison (b), using all product-location-day combinations in the test period.
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/invquant_dayofweek_nbinom}
-\includegraphics[width=4cm]{figs/invquant_dayofweek_poisson}
+\includegraphics[width=4cm]{invquant_dayofweek_nbinom}
+\includegraphics[width=4cm]{invquant_dayofweek_poisson}
 \caption{\label{fig:invquant_dayofweek} Inverse quantile profile plot for the different days of the week on the $x$ axis (from Monday to Sunday) aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/invquant_yhat_mean_nbinom}
-\includegraphics[width=4cm]{figs/invquant_yhat_mean_poisson}
+\includegraphics[width=4cm]{invquant_yhat_mean_nbinom}
+\includegraphics[width=4cm]{invquant_yhat_mean_poisson}
 \caption{\label{fig:invquant_mean} Inverse quantile profile plot for mean predictions on the $x$ axis aggregated over all product-location-day combinations in the test period, for our negative binomial model (a) and a simpler Poisson model for comparison (b).}
@@
 \begin{center}
-\includegraphics[width=4cm]{figs/marginal}
-\includegraphics[width=4cm]{figs/scatter}
+\includegraphics[width=4cm]{marginal}
+\includegraphics[width=4cm]{scatter}
 \caption{\label{fig:scatter} Marginal distributions and scatter plot of variables $X$ and $Y$.}
@@
 \begin{center}
-\includegraphics[width=8cm]{figs/profile}
+\includegraphics[width=8cm]{profile}
 \caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}

### 2021-06-21 LaTex/paper.tex
few additional sentences for SNORF resubmit

@@
 \affil[2]{\small Blue Yonder India Private Limited (Bengaluru, India)}
-\affil[3]{\small IUBH Internationale Hochschule (Erfurt, Germany)}
+\affil[3]{\small IU Internationale Hochschule (Erfurt, Germany)}
 
@@
 
+This work brings together elements from three different fields: operations research, machine learning, and causal inference.
+The combination of these allows to build state-of-the-art demand forecasting models that are fully explainable and incorporate the causal structure of demand.
 The contribution of this paper is two-fold:
@@
 The general methods developed in this paper can be applied to any setting that requires the prediction and evaluation of probability distributions from observational data, not just demand forecasting.
-However, to make the impact and benefits more tangible in a practical setting, we limit the discussion to demand forecasting in retail.
+However, to make the impact and benefits more tangible in a practical setting, as well as describe a general methodology how to build such systems from a practitioner's perspective, we limit the discussion to demand forecasting in retail.
 
@@
 
-As forecasting relies upon statistical dependencies, no matter if spurious or from true causal effects, this does not seem to be a problem for the predictive power at first glance.
+As traditional statistical forecasting or machine learning methods generally rely upon statistical dependencies, no matter if spurious or from true causal effects, this does not seem to be a problem for the predictive power at first glance.
 However, it can have severe impacts on the explainability and generalizability of the model, because the autocorrelation, being a composed effect of many causes, is less stable over time than the actual causal effects.

### 2021-07-11 LaTex/paper.tex
some fixes from SNORF proof reads

@@
 {Keywords: \textbf{explainable machine learning, retail demand forecasting, probability distribution, temporal confounding}}
+
+{DOI: 10.1007/s43069-021-00079-8}
+
+{Published in Springer Operations Research Forum.}
 
@@
 The combination of these allows to build state-of-the-art demand forecasting models that are fully explainable and incorporate the causal structure of demand.
-The contribution of this paper is two-fold:
+The contribution of this paper is twofold:
 
@@
 
-Furthermore, we show how predictions of full individual probability distributions, no matter if created by an approach with or without assumption of a specific distribution class, can be evaluated qualitatively and quantitatively, such that we can assess if both overall functional form (in the case of a parametric approach) as well as specific quantiles are predicted correctly.
+Furthermore, we show how predictions of full individual probability distributions, no matter if created by an approach with or without assumption of a specific distribution class, can be evaluated qualitatively and quantitatively, such that we can assess whether both overall functional form (in the case of a parametric approach) as well as specific quantiles are predicted correctly.
 This allows to verify that the predicted demand distribution appropriately reflects the observed data in terms of model choice and forecast accuracy.
@@
 Demand forecasting is one of the core operational activities of any customer-facing business and accurate demand predictions are vital to the optimization of the business operations.
-This is of particular concern for example for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}.
+This is of particular concern, for example, for supermarkets, as their profit margin tend to be very low, e.g., 1\% and lower for European discounters \cite{MarginLidlSweden} to about 4\% for regular supermarkets \cite{MarginWillysSweden,MarginHemkopSweden} to about 8\% in the USA \cite{MarginPublixUSA}.
 It can be approached in a variety of ways, see, e.g., \cite{beheshti2015survey} for a survey of methods used in retail sales forecasting with a particular focus on fashion retail.
@@
 The interest in deep learning approaches for demand forecasting has increased significantly recently, see, e.g., \cite{bandara2019sales,yu2017application, goyal2018solution,helmini2019sales, golkabek2020demand}.
-A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}.
+A review of deep learning methods for time series forecasting in a wide range of applications can also be found in \cite{langkvist2014review}, and the use of RNNs with focus on industrial applications is discussed in \cite{dixon2020industrial}.
 A further approach is to explore the use of generative adversarial networks \cite{goodfellow2014generative} in the context of time series forecasting \cite{haas2020statistical, ramponi2018t, smith2020conditional}.
@@
 In the simplest case, we only have one model with one set of model parameters to cover all predictions.
-In this case, the evaluation of the full PDF is straight-forward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as an NBD with predicted parameters, that is used for all observations.
+In this case, the evaluation of the full PDF is straightforward: We would fill a histogram of all observed values, such as sales records, and overlay this with the single model, such as an NBD with predicted parameters, that is used for all observations.
 Then, we compare the model curve directly with the observations, using statistical tests such as the Kolmogorov-Smirnov test.
@@
 
-For a first qualitative assessment, we make use of the probability integral transform, see e.g., \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1.
+For a first qualitative assessment, we make use of the probability integral transform, see, e.g., \cite{Angus1994,casella2002statistical}, which states that a random variable distributed according to the CDF of another random variable is uniformly distributed between 0 and 1.
 We therefore expect that the distribution of the actually observed CDF values of the corresponding individual PDF predictions is uniform, if the predicted PDF is calibrated correctly, {\em regardless} of the shape of the predicted distribution.
@@
 In order to obtain a uniform distribution for discrete PDFs that is comparable to the case of continuous distributions, the histogram holding the values of the CDF is filled using random numbers according to the intervals of the CDF.
-For example, if the sales of zero items accounts for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
+For example, if the sales of zero items account for 75 percent of the observed sales distribution for this item, the value of the CDF function that is used to fill the histogram in case of zero actual sales is randomly chosen in the interval $[0, 0.75]$.
 Proceeding similarly for all other observed values, with the intervals from which to randomly choose values to fill in the histogram defined by the CDF values of the corresponding discrete sales value and the one below (e.g., for 3 actual sales: random pick between discrete CDF values for 2 and 3), the resulting histogram of CDF values is again uniform, as in the case of a continuous PDF.
@@
 \fig \ref{fig:invquant_example} illustrates five different collections of inverse quantile profile plots (each collection comparing to 6 specified quantile values, namely $q = 0.1$, $q = 0.3$, $q = 0.5$, $q = 0.7$, $q = 0.9$, and $q = 0.97$), for separate sets of exemplary PDF estimations and observed data combinations.
-The dashed horizontal lines indicates the fraction we expect, i.e., the specified quantile value, if the predictions are correct.
+The dashed horizontal lines indicate the fraction we expect, i.e., the specified quantile value, if the predictions are correct.
 For example, for the median, the line at $q = 0.5$ indicates that  50 percent of all PDF prediction and observed data combinations in a given data set should fall above the line, and 50 percent should fall below the line.
@@
 For example, if we were to forecast three days ahead into the future, the lag of the feature time series with respect to the target time series has to be three as well.
-By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past, and use these as features for the machine learning model.
+By creating many variants of the lagged target information contained in the original time series, for example by varying the value of the hyperparameter $\alpha$ or calculating distinct moving averages for the different days of the week, we can build a wide range of variables capturing the dynamics of the recent past and use these as features for the machine learning model.
 In addition, we can include further covariates in the model, such as pricing information, promotion details, characteristics of the products we wish to forecast, or characteristics of the location at which the product is sold.
@@
 \noindent
-The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the inverse dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends towards lower values.
+The different \texttt{item\_store} values for mean (higher than $1$) and variance model factors (lower than $1$), including the static one-dimensional features \texttt{store\_id} and \texttt{item\_id} and their two-dimensional combination, represent the fact that this product-location combination sells more than the average of all product-location combinations, and the inverse dispersion parameter $1/r$ in \eqn \eqref{eqn:variance_r} in turn tends toward lower values.
 The explanation of the model for the peak in sales and mean prediction on February 7, 2016 is the event Super Bowl.
@@
 The latter case primarily stems from slow moving articles with less than one item sold per store per day on average, what can be seen in plots c and d of \fig \ref{fig:cdf_demand}, where the spike at CDF values close to $1$ is reduced by excluding all samples with mean predictions lower than $1.0$.
-The resulting distribution for the negative binomial model indicates a slight tendency towards too broad PDF predictions, at least for some observations.
+The resulting distribution for the negative binomial model indicates a slight tendency toward too broad PDF predictions, at least for some observations.
 
@@
 While the Poisson model again shows a significant pattern of too narrow PDF predictions across the full range of mean prediction values, we can also see several deviations from the expected uniform behavior for the negative binomial model.
-For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing towards too broad PDF predictions (as also observed in the CDF histogram above).
+For mean predictions around $30$, the negative binomial model deviates significantly from the expected behavior, with the shape of the deviations pointing toward too broad PDF predictions (as also observed in the CDF histogram above).
 And for mean predictions higher than $80$, the pattern of the deviations indicate an underprediction of the mean parameter.
@@
 Therefore, the integral from $-\infty$ to the point $x_p$ is the same as from $x_p$ to $\infty$, which means that the point $x_p$ splits the integral in the interval $(-\infty, \infty)$ into two halves.
-This is the definition of the median, and therefore a linear cost function with equal over- and underage costs is linked to the median of the predicted PDF as optimal point estimator.
+This is the definition of the median, and therefore, a linear cost function with equal over- and underage costs is linked to the median of the predicted PDF as optimal point estimator.
 If the underage and overage costs differ, the slopes of the two linear segments are no longer the same.
@@
 \begin{center}
-\includegraphics[width=8cm]{profile}
+\includegraphics[width=4cm]{profile}
 \caption{\label{fig:profile} Profile histogram of variables $X$ and $Y$.}

### 2021-07-11 code/gaussian.py

@@ -1,0 +1,15 @@
+import numpy as np
+from scipy.stats import norm
+import matplotlib.pyplot as plt
+
+fig, axs = plt.subplots(2, 1)
+
+x = np.linspace(norm.ppf(0.0001), norm.ppf(0.9999), 10000)
+axs[0].fill_between(x, norm.pdf(x))
+axs[0].set_title('f(x)')
+axs[0].axes.xaxis.set_ticklabels([])
+axs[1].plot(x, norm.cdf(x), 'k--')
+axs[1].set_title('F(x)')
+
+fig.savefig('PdfCdf.pdf')
+
